### 2015-03-24 leaflifespan.Rmd
Move functions into utils file, move packages into remake.yml

---
output: html_document
---
#file names
- leaf_counts is file with calculated leaf counts and leaf lifespan data
- LLSummary is file with means of LL - and other variables - by species and age; LMA data is merged to this file
- subsets of LLSummary are created to omit babies or HATE for various analyses

# general results

- strong correlation between LMA and LL when all species*site combinations included
- LL variable across species, as expected; same for LMA
- LL does not change as plants age; LMA does not change as plants age
- leaf count along the leader decreases in older plants for many species - which is why LL appears to decrease

#next to do

- check how overall plant leaf numbers changes with age, diameter

```{r, echo=FALSE,results='hide', message=FALSE}
# load packages

library(remake)
create_bindings()

# InvestmentSummary <- InvestmentSummary
# names(InvestmentSummary)[1] <- "tag_ID"

LMA <- read.csv("data/LMA_short.csv",header=TRUE,stringsAsFactors=FALSE)%>%
  filter(species!="" & species!=" ")

wood <- read.csv("data/woodDensity.csv", header=TRUE,stringsAsFactors=FALSE)%>%
  filter(use=="use")

seedsize <- read.csv("data/seedsize.csv", header=TRUE,stringsAsFactors=FALSE)

# leaf lifespan data
raw_leaf_lifespan <- read.csv("data/leaderGrowthLL.csv",header=TRUE,stringsAsFactors=FALSE) %>%
  filter(dont_use!="dead" & dont_use!="dont_use")

# Create new dataset taking only count measures, bc these are ultimately what we want
leaf_lifespan <- raw_leaf_lifespan %>%
  select(species,age,site,tag_ID,d_start,d_end,length_start,new_length,segment,lvs_start_count,lvs_end_count,lvs_new_count,new_but_shed_count)

# where counts did not previously exists, set as 0; CHECK WITH DANIEL THAT "V" IS A SET "NAME" IN THESE FUNCTIONS
for(v in c("lvs_start_count","lvs_end_count","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leaf_lifespan[[v]])
  leaf_lifespan[[v]][i] <- 0
}

# For individuals with "length" measures, translate into count

# get average leaves per length for spp
leaves_per_length_spp <- read.csv("data/leaves_per_length.csv", header=TRUE, stringsAsFactors=FALSE) %>%
  filter(use=="yes") %>%
  mutate(count_per_length = leaf_count/length_mm) %>%
  group_by(species) %>%
  summarise(count_per_length = mean(count_per_length, na.rm=TRUE))

# get spp average for all individuals 
raw_leaf_lifespan$count_per_length <- leaves_per_length_spp$count_per_length[match(raw_leaf_lifespan$species, leaves_per_length_spp$species)]

# if possible use data for that individual, if it exists; this code substitutes in species count per length #s for individuals where this data is available
i <- !is.na(raw_leaf_lifespan$mm_lvs)
raw_leaf_lifespan$count_per_length[i] <- (raw_leaf_lifespan$count_lvs_spec/raw_leaf_lifespan$mm_lvs)[i]

#this is the end of the section Daniel helped me rewrite - next section is mine

#get total leaf counts
for(v in c("lvs_start_length","lvs_start_count","lvs_end_length","lvs_end_count","lvs_new_length","lvs_new_count","new_but_shed_count")) {
  i <- is.na(raw_leaf_lifespan[[v]])
  raw_leaf_lifespan[[v]][i] <- 0
  }
raw_leaf_lifespan$lvs_start <- raw_leaf_lifespan$lvs_start_count + (raw_leaf_lifespan$lvs_start_length * raw_leaf_lifespan$count_per_length)
raw_leaf_lifespan$lvs_end <- raw_leaf_lifespan$lvs_end_count + (raw_leaf_lifespan$lvs_end_length * raw_leaf_lifespan$count_per_length)
raw_leaf_lifespan$lvs_new <- raw_leaf_lifespan$lvs_new_count + (raw_leaf_lifespan$lvs_new_length * raw_leaf_lifespan$count_per_length)

#calculating leaf lifespan
raw_leaf_lifespan$LL_bd <- (raw_leaf_lifespan$lvs_start)/(((raw_leaf_lifespan$lvs_start)-(raw_leaf_lifespan$lvs_end))+(raw_leaf_lifespan$lvs_new)/2)
raw_leaf_lifespan$LL_death <- (raw_leaf_lifespan$lvs_start)/((raw_leaf_lifespan$lvs_start)-(raw_leaf_lifespan$lvs_end))
raw_leaf_lifespan$LL_birth <- (raw_leaf_lifespan$lvs_start)/(raw_leaf_lifespan$lvs_new)

raw_leaf_lifespan$LL_death[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA
raw_leaf_lifespan$LL_bd[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA
raw_leaf_lifespan$LL_birth[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA

#write.csv(raw_leaf_lifespan, file = "raw_leaf_lifespan.csv")
```

#TODO: Check outlier for HATE

```{r, echo=FALSE,results='hide', message=FALSE}
#adding leaf, stem weight data from InvestmentSummary to raw_leaf_lifespan dataframe; can't RUN LINES 95-108 UNTIL INVESTMENT SUMMARY OPEN - THEN PROBLEM WITH PLYR
#raw_leaf_lifespan <- left_join(raw_leaf_lifespan, InvestmentSummary, by = "tag_ID")

#adding columns to raw_leaf_lifespan to look at change across year
raw_leaf_lifespan$change_shoot_diam <- raw_leaf_lifespan$d_end - raw_leaf_lifespan$d_start
#raw_leaf_lifespan$change_basal_diam <- raw_leaf_lifespan$FinalBasalDiam - raw_leaf_lifespan$StartBasalDiam
raw_leaf_lifespan$change_shoot_area <- (3.14*((raw_leaf_lifespan$d_end/2)^2)) - (3.14*((raw_leaf_lifespan$d_start/2)^2))
#raw_leaf_lifespan$change_basal_area <- (3.14*((raw_leaf_lifespan$FinalBasalDiam/2)^2)) - (3.14*((raw_leaf_lifespan$StartBasalDiam/2)^2))

#removng crazy numbers; probably better to do this for each analysis than overall, because removing indiviiduals from all three at once
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_bd>0)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_bd<6)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_birth<8)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_death<8)

#creating summary files
#leaf lifespan summary by species, age
LLSummary <- raw_leaf_lifespan %>%
  filter(LL_bd>0 & LL_bd<6 & LL_birth<8 & LL_death<8) %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length, lvs_end, lvs_end_total)

#LMA summary by species, age
LMA_summary <- LMA %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LMA)
names(LMA_summary)<-c("species","age","LMA_mean","LMA_se","LMA_length")

#wood density summary by species
Wood_summary <- wood %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), density)
names(Wood_summary)<-c("species","WD_mean","WD_se","WD_length")

#leaf lifespan summary by species
LL_spp <- raw_leaf_lifespan %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

#LMA summary by species
LMA_spp <- LMA %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LMA)
names(LMA_spp)<-c("species", "LMA_mean","LMA_se","LMA_length")

#merge LMA, wood density, seed size data into LLSummary dataframe
LLSummary <- merge(LLSummary, LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
LLSummary <- merge(LLSummary, Wood_summary, by.x=c("species"), by.y=c("species"))
LLSummary <- merge(LLSummary, seedsize, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, LMA_spp, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, Wood_summary, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, seedsize, by.x=c("species"), by.y=c("species"))
#still to add, max H from Investment data, absolute and relative growth measures

#plotting and stats
#creating lists of information for plotting symbols, colors, labels
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","dark green")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU") 
labels.age <- c("1.4","2.4","5","7","9","32")

#LL vs LMA
#remove HATE, babies
#instead of separate dataframes, can probably incorporate this information directly into plotting and analysis commands
#LL_nobabies <- subset(raw_leaf_lifespan, age>2)
#LMA_nobabies <- subset(LMA, age>2)
#LL_noHate <- subset(LLSummary, species != "HATE")
#LL_noHatebabies <- subset(LL_nobabies, species != "HATE")
```

# relationship between LMA, LL by age, species

For 7 year old plants
```{r, echo=FALSE}
plot(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7), xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA", las=1, pch=16)
text(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7), labels=species, pos=1, offset=-0.9, cex=0.9,col="blue")
text(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7 & species=="HATE"), labels=species, pos=1, offset=-0.9, cex=0.9,col="red")
mod <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,species!="HATE" & age==7))
summary(mod)
abline(mod)
```

# Plotting final leaf number on the leader by diameter increase of the entire plant (basal diameter) and the leader (shoot diameter).

Both show a strong correlation and species effect as expected. For the model with basal diameter there is a strong age effect, but not for shoot diameter. This is presumably because a shoot of a given species should always deploy the same number of leaves for the same diameter increase. In contrast, the whole plant will have many differing shoot numbers with age, so differing effects of a single shoot's increased diameter. Strangely, the model with basal diameter shows a stronger effect between diameter increase and leaf number.

```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_basal_diam ~ leaf_new_total, raw_leaf_lifespan, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in basal diameter")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_basal_diam+1) + species +age, raw_leaf_lifespan)
summary(mod)

plot(change_shoot_diam ~ leaf_new_total, raw_leaf_lifespan, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in diameter of leader")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_shoot_diam+10) + species +age, data= subset(raw_leaf_lifespan, !is.na(change_shoot_diam)))
summary(mod)
```

#LL vs LMA using species by site means

Overall plot shows a lack of relationship when HATE included. This is because it has such a high LMA. The youngest and oldest plants have no pattern, presumably because LL data are inaccurate

```{r, echo=FALSE,results='hide', message=FALSE}

col.LMA <-c("black","purple", "pink", "light green", "orange","blue", "light blue","red")
labels.LMA <-c("all except HATE","age 1.4", "age 2.4", "age 5", "age 7", "age 9", "age 32", "including HATE")
plot(LL_birth_mean ~ LMA_mean, LLSummary, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(LLSummary, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(LLSummary, species=="HATE"), pch=17)
#need log transformation to get these to plot onto axes
mod <- lm(LL_birth_mean ~ LMA_mean, LLSummary)
abline(mod, col="red", lty=2)
mod_no_HATE <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary, species!="HATE"))
abline(mod_no_HATE, lwd=2)
mod_no_HATE5 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==5&species!="HATE"))
abline(mod_no_HATE5, col="light green", lwd=1.5)
mod_no_HATE7 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7&species!="HATE"))
abline(mod_no_HATE7, col="orange", lwd=1.5)
mod_no_HATE9 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==9&species!="HATE"))
abline(mod_no_HATE9, col="blue", lwd=1.5)
mod_no_HATE2.4 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==2.4&species!="HATE"))
abline(mod_no_HATE2.4, col="pink", lwd=1.5)
mod_no_HATE1 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age<2.4&species!="HATE"))
abline(mod_no_HATE1, col="purple", lwd=1.5)
mod_no_HATE32 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==32&species!="HATE"))
abline(mod_no_HATE32, col="light blue", lwd=1.5)

legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1.4)
summary(mod)
```

#Leaf lifespan patterns
First a model showing data by age and species, omitting the youngest plants. Uses calculations for "leaf births" only. Plots of leaf lifespan by age for each of the species

Age is not significant, but some significant interactions (COER, HATE, HEPU)
And many species have sigificantly different LL from one another

```{r, echo=FALSE,results='hide', message=FALSE}
mod <- lm(LL_birth ~ age + species +age*species, data=subset(raw_leaf_lifespan,age>3 & LL_birth!="Inf"))
summary(mod)

#WHY ARE THE FOLLOWING TWO LINES OF CODE NEEDED?
raw_leaf_lifespan %>% group_by(species) %>%
    summarise(mean(LL_birth))

#where does the [spp] comes from? This is the first place in the script it is used like this - I assume it could be any name?
plot_LLvAge <- function(spp) {
  plot(LL_birth ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf lifespan (leaf births)", cex.axis=.8, las=1, 
     main=labels.spp[spp], xlim=c(1,35), ylim=c(0,8), log="x", pch=16)  
  points(LL_birth_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((LL_birth_mean+LL_birth_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((LL_birth_mean-LL_birth_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

for(n in names(labels.spp)) {
  plot_LLvAge(n)
}

```

#LMA by age, species, without youngest plants

There is also no overall affect of age on LMA, although a few species would show an effect if analyzed independently. (Again, this model omits 1-year old plants.)

```{r, echo=FALSE, message=FALSE}
#plot(LMA_mean ~ as.factor(species), data=subset(LLSummary,age>3), xlab="species", ylab="LMA", cex.axis=.8, las=1)
#mod <- lm(LMA ~ age +species + age*species, LMA)
#summary(mod)

LMA %>% group_by(species) %>%
  summarise(mean(LMA))

plot_LMAvAge <- function(spp) {
  plot(LMA ~ age, data=subset(LMA,species==spp), xlab="age", ylab="LMA", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(LMA_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((LMA_mean+LMA_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((LMA_mean-LMA_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
  #I want to add a line of code that plots over the means for each age. I did this above by calling in another file, but wondering about on-the-spot calculations
}

mod_LMAvAge <- function(spp) {
  lm(LMA ~ age, data=subset(LMA,species==spp))
}

for(n in names(labels.spp)) {
  plot_LMAvAge(n)
  mod_LMAvAge(n)
  summary(mod_LMAvAge(n))
}

```

#models with leaf counts per age
The number of leaves on a leader is slightly affected by age however, with a decrease in leaf count along the leader with increasing age. This is what gives the perception of fewer leaves in some older plants. I'm not sure why this pattern looks stronger than leaf lifespan.

```{r, echo=FALSE,results='hide', message=FALSE}

plot_lfcountvAge <- function(spp) {
  plot(lvs_end_total ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf count along shoot", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(lvs_end_total_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((lvs_end_total_mean+lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((lvs_end_total_mean-lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

mod_lfcountvAge <- function(spp) {
  lm(lvs_end ~ age, data=subset(raw_leaf_lifespan,age>2.4 & species==spp))
}

for(n in names(labels.spp)) {
  plot_lfcountvAge(n)
  mod_lfcountvAge(n)
  summary(mod_lfcountvAge(n))
}

```

#models with shoot extenstion by age
These models show that for some species, shoot extension decreases enormously in the older plants. This effect is larger than with number of leaves.

```{r, echo=FALSE,results='hide', message=FALSE}

plot_newlengthvAge <- function(spp) {
  plot(new_length ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leader shoot extention", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(new_length_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((new_length_mean+new_length_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((new_length_mean-new_length_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

mod_newlengthvAge <- function(spp) {
  lm(new_length ~ age, data=subset(raw_leaf_lifespan,age>2.4 & species==spp))
}

for(n in names(labels.spp)) {
  plot_newlengthvAge(n)
  mod_newlengthvAge(n)
  summary(mod_newlengthvAge(n))
}
```

#plots regressing variables against one another
```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_shoot_area~new_length,data=subset(raw_leaf_lifespan,change_shoot_area<1000000), log="xy",col=col.age[as.factor(age)], pch=16, cex=.8, xlab="shoot extension (mm)", ylab="increased shoot cross sectional area")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(change_shoot_area~new_length +age,data=subset(raw_leaf_lifespan,change_shoot_area<1000000))
summary(mod)

raw_leaf_lifespan$lvs_end_total <- raw_leaf_lifespan$lvs_end + raw_leaf_lifespan$lvs_new
plot(lvs_end_total~lvs_start,raw_leaf_lifespan, log="xy",col=col.age[as.factor(age)], pch=16, cex=.8, xlab="starting leaf count", ylab="ending leaf count")

legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(lvs_start~lvs_end_total+age +age*lvs_end_total, raw_leaf_lifespan)
summary(mod)
#IS THERE AN EASY WAY TO ADD A 1:1 LINE TO THIS PLOT?

plot(LL_birth_mean~LMA_mean,data=subset(LLSummary,species!="HATE"),col=col.age[as.factor(age)], pch=16, cex=.8, log="xy",xlab="LMA", ylab="leaf lifespan (calculated from leaf births")
points(LL_birth_mean~LMA_mean,data=subset(LLSummary,species=="HATE"),col="red", pch=19, cex=.8)