### 2015-03-26 talk.Rmd
build talk in remake

---
title: "Untitled"
output:
     pdf_document:
       toc: true
       highlight: zenburn
     html_document:
       toc: true
       theme: united
---

# file names

- leaf_counts is file with calculated leaf counts and leaf lifespan data
- SummarySppAge is file with means of LL - and other variables - by species and age; LMA data is merged to this file
- subsets of SummarySppAge are created to omit babies or HATE for various analyses