### 2015-03-26 talk.Rmd

---
title: "Untitled"
output:
     pdf_document:
       toc: true
       highlight: zenburn
     html_document:
       toc: true
       theme: united
---
```{r, echo=FALSE,results='hide', message=FALSE}

# load packages and "open" dataframes to use
library(remake)
create_bindings()
```

```{r, echo=FALSE,results='hide', message=FALSE}
LMA <- filter(LMA_raw, species!="" & species!=" ") %>%
  select(species,age,LMA,branch_age)

wood <- filter(woodDensity_raw, use=="use") %>%
  select(species,density)

seedsize <- seedSize_raw %>%
  select(species,seed_size)

harvest <- filter(HarvestData, use_status=="use" & segment==1 & start_end=="end") %>%
  select(species,individual,age,age2,height,total_leaf_weight,site)

investment <- ReproductionAllocation_all
investment$age2 <- investment$age
investment$age <- round(investment$age, digits=1)

accessory <- AccessoryCosts_all
accessory$age2 <- accessory$age
accessory$age <- round(accessory$age, digits=1)
#accessory$PerAccessory <- {
#  select(accessory, Total >0,
#  1 - (accessory$Prop/accessory$Total)
#}

#accessory$PerPropagule = 1 - accessory$PerAccessory

leafLifespan <- filter(leafLifespan_raw, dont_use!="dead" & dont_use!="dont_use") %>%
  select(species,age,individual,d_start,d_end,length_start,new_length,lvs_start_length,lvs_start_count,lvs_end_length,lvs_end_count,lvs_new_length,lvs_new_count,new_but_shed_count,mm_lvs_spec,count_lvs_spec)

leafLifespan$individual <- as.factor(leafLifespan$individual)
leafLifespan$species <- as.factor(leafLifespan$species)

#loading this column as character, so using this for now

# Calculating Leaf Lifespan

# where counts did not previously exists, set as 0; CHECK WITH DANIEL THAT "V" IS A SET "NAME" IN THESE FUNCTIONS

leaf_lifespan_temp <- leafLifespan %>%
              select(species,age,individual,d_start,d_end,length_start,new_length,lvs_start_count,lvs_end_count,lvs_new_count,new_but_shed_count)

for(v in c("lvs_start_count","lvs_end_count","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leaf_lifespan_temp[[v]])
  leaf_lifespan_temp[[v]][i] <- 0
}

# For individuals with "length" measures, translate into count

# get average leaves per length for spp
leaves_per_length_spp <-  filter(leavesPerLength_raw,use=="yes") %>%
  mutate(count_per_length = leaf_count/length_mm) %>%
  group_by(species) %>%
  summarise(count_per_length = mean(count_per_length, na.rm=TRUE))

# get spp average for all individuals
leafLifespan$count_per_length <- leaves_per_length_spp$count_per_length[match(leafLifespan$species, leaves_per_length_spp$species)]

# if possible use data for that individual, if it exists; this code substitutes in species count per length #s for individuals where this data is available
i <- !is.na(leafLifespan$mm_lvs)
leafLifespan$count_per_length[i] <- (leafLifespan$count_lvs_spec/leafLifespan$mm_lvs)[i]

#this is the end of the section Daniel helped me rewrite - next section is mine

#get total leaf counts
for(v in c("lvs_start_length","lvs_start_count","lvs_end_length","lvs_end_count","lvs_new_length","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leafLifespan[[v]])
  leafLifespan[[v]][i] <- 0
  }
leafLifespan$lvs_start <- leafLifespan$lvs_start_count + (leafLifespan$lvs_start_length * leafLifespan$count_per_length)
leafLifespan$lvs_end <- leafLifespan$lvs_end_count + (leafLifespan$lvs_end_length * leafLifespan$count_per_length)
leafLifespan$lvs_new <- leafLifespan$lvs_new_count + (leafLifespan$lvs_new_length * leafLifespan$count_per_length)

#calculating leaf lifespan
leafLifespan$LL_bd <- (leafLifespan$lvs_start)/(((leafLifespan$lvs_start)-(leafLifespan$lvs_end))+(leafLifespan$lvs_new)/2)
leafLifespan$LL_death <- (leafLifespan$lvs_start)/((leafLifespan$lvs_start)-(leafLifespan$lvs_end))
leafLifespan$LL_birth <- (leafLifespan$lvs_start)/(leafLifespan$lvs_new)

leafLifespan$LL_death[which(is.infinite(leafLifespan$LL_death))] <- NA
leafLifespan$LL_bd[which(is.infinite(leafLifespan$LL_death))] <- NA
leafLifespan$LL_birth[which(is.infinite(leafLifespan$LL_death))] <- NA

#adding investment and accessory costs data to leafLifespan dataframe to create a dataframe with all individual level data
SummaryInd <- merge(leafLifespan, select(investment, -species, -age), by="individual")
SummaryInd <- merge(SummaryInd, select(accessory, -species, -age, -age2), by="individual")
SummaryInd <- merge(SummaryInd, select(harvest, -species, -age, -age2), by="individual")

#adding variables to look at change across years
SummaryInd$change_shoot_diam <- SummaryInd$d_end - SummaryInd$d_start
SummaryInd$change_basal_diam <- SummaryInd$FinalBasalDiam - SummaryInd$StartBasalDiam
SummaryInd$change_shoot_area <- (3.14*((SummaryInd$d_end/2)^2)) - (3.14*((SummaryInd$d_start/2)^2))
SummaryInd$change_basal_area <- (3.14*((SummaryInd$FinalBasalDiam/2)^2)) - (3.14*((SummaryInd$StartBasalDiam/2)^2))
SummaryInd$lvs_end_total <- SummaryInd$lvs_end + SummaryInd$lvs_new
```

```{r, echo=FALSE,results='hide', message=FALSE}
#summarizing data by species, age

#leaf lifespan summary by species, age
LL_summary <- SummaryInd %>%
  filter(LL_bd>0 & LL_bd<6 & LL_birth<8 & LL_death<8) %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length, lvs_end, lvs_end_total, change_shoot_diam, change_basal_diam, change_shoot_area, change_basal_area)

#LMA summary by species, age
LMA_summary <- LMA %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LMA)
names(LMA_summary)<-c("species","age","LMA_mean","LMA_se","LMA_length")

#accessory tissue summary by species, age
accessory_summary <- accessory %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), PrePol_A, PrePol_S, PostPol_A, PD, Prop)

#investment summary by species, age
investment_summary <- investment %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), GrowthInv, FinalWeight, FinalBasalDiamAv, TotalInvestment, RA)

#harvest summary by species, age
harvest_summary <- harvest %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), height,total_leaf_weight)
#names(harvest_summary)<-c("species","age","height_mean","height_se","height_length")

#RA for each species/age combination
maxRA <- investment %>%
  group_by(species, age) %>%
  summarise_each(funs(max), RA)
names(maxRA)<-c("species","age","maxRA")

#maximum H for each species/age combination
maxH <- harvest %>%
  group_by(species, age) %>%
  summarise_each(funs(max), height)
names(maxH)<-c("species","age","maxH")

#LMA summary by species, age
LMA_summary <- LMA %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LMA)
names(LMA_summary)<-c("species","age","LMA_mean","LMA_se","LMA_length")

#wood density summary by species
wood_summary <- wood %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), density)
names(wood_summary)<-c("species","WD_mean","WD_se","WD_length")

#leaf lifespan summary by species
LL_spp <- leafLifespan %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

#LMA summary by species
LMA_spp <- LMA %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LMA)
  names(LMA_spp)<-c("species", "LMA_mean","LMA_se","LMA_length")

#maximum height for each species
maxH_spp <- harvest %>%
  group_by(species) %>%
  summarise_each(funs(max), height)
names(maxH_spp)<-c("species", "maxH_spp")

SummaryInd <- merge(SummaryInd,wood_summary, by=c("species"))
SummaryInd <- merge(SummaryInd,LMA_summary, by=c("species","age"))
SummaryInd <- merge(SummaryInd,maxH_spp, by=c("species"))
SummaryInd <- merge(SummaryInd,seedsize, by=c("species"))

SummaryInd$leafArea <- SummaryInd$total_leaf_weight * SummaryInd$LMA_mean

#merge various data summaries into SummarySppAge dataframe
SummarySppAge <- merge(LL_summary, LMA_summary, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, wood_summary, by=c("species"))
SummarySppAge <- merge(SummarySppAge, seedsize, by=c("species"))
SummarySppAge <- merge(SummarySppAge, maxH_spp, by=c("species"))
SummarySppAge <- merge(SummarySppAge, maxH, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, investment_summary, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, accessory_summary, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, harvest_summary, by=c("species","age"))

#merge various data summaries into SummarySpp dataframe
SummarySpp <- merge(LL_spp, LMA_spp, by=c("species"))
SummarySpp <- merge(SummarySpp, wood_summary, by=c("species"))
SummarySpp <- merge(SummarySpp, seedsize, by=c("species"))
SummarySpp <- merge(SummarySpp, select(maxH, -age), by=c("species"))
```

#plotting and stats
```{r, echo=FALSE,results='hide', message=FALSE}
#creating lists of information for plotting symbols, colors, labels
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","dark green")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")
labels.age <- c("1.4","2.4","5","7","9","32")
names(labels.age) <- c("1.4","2.4","5","7","9","32")
labels.age2 <- c("2.4","5","7","9","32")
col.age2 <-c("pink","green","orange","blue","dark green")
col.age3 <-c("green","orange","blue","dark green")
```

# file names

- leaf_counts is file with calculated leaf counts and leaf lifespan data
- SummarySppAge is file with means of LL - and other variables - by species and age; LMA data is merged to this file
- subsets of SummarySppAge are created to omit babies or HATE for various analyses

#What causes RA to shift with age in species?
Theoretical models indicate that there need to be size (or age) related shifts in *some* trade-off to create graded RA schedules.

For instance, shifts in a variable such as leaf lifespan, fecundity per RA, photosynthetic rates with size (or other variables) will change the RA schedule.

I have both RA schedules and lots of other data across multiple ages:
- leaf lifespa
- LMA
- accessory tissue costs

#Shifts in RA with age

Showing just a few of the RA schedules from my last lab meeting.

They show that species differ greatly in RA, that there is a general increase with age, but also that the data are noisy.

```{r, echo=FALSE,results='hide', message=FALSE}
plot_RAbyAge <- function(spp) {
  plot(RA~age, data=subset(SummaryInd,species==spp), xlab="age (years)",ylab="reproductive allocation (%)", log="x", main=labels.spp[spp],xlim=c(1,35),ylim=c(0,1))
  points(RA_mean~age, data=subset(SummarySppAge, species==spp),pch=16,cex=1.4)
}

for(n in names(labels.spp)) {
  plot_RAbyAge(n)
}
```
#Individual shoots looked decidedly "less leafy" in many older plants
This led me to wonder how leaf lifespan - or some other variable - was shifting with age

The number of leaves along a leader is overall uneffected by age, although some species show a marked decline
```{r, echo=FALSE,results='hide', message=FALSE}
plot(lvs_end_total~age,SummaryInd,log="xy", xlab="age",ylab="final leaf count along shoot",col=col.spp[as.factor(species)])
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

#LMA versus age
The plot below is a bulked LMA sample from each site, of leaves from the second most recent cohort

There is a significant increase with age (p=0.002) and many species differ from one another, despite the very few points
Age is only borderline significant (p=0.04) when the seedlings are removed

```{r, echo=FALSE,results='hide', message=FALSE}
plot(LMA~age,data=subset(LMA,branch_age=="mid"),log="xy", xlab="age",ylab="LMA",col=col.spp[as.factor(species)],pch=16)
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

#Leaf lifespan with age
Leaf lifespan increases from seedlings (1.4 yrs) to saplings (2.4 years) to adults (>4 years), but there are no differences between the different adult populations
```{r, echo=FALSE,results='hide', message=FALSE}
plot(LL_birth~age,SummaryInd,log="xy", xlab="age",ylab="leaf lifespan (based on leaf births)",col=col.spp[as.factor(species)])
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

#Plots of individual species show that no species has a significant change either - although several show slight declines at the end
```{r, echo=FALSE,results='hide', message=FALSE}
#DF: Here is a place where it would be good to run the statistical model on each of the species. I have the function I wrote below, ut hashed out

plot_LLvAge <- function(spp) {
  plot(LL_birth ~ age, data=subset(SummaryInd,species==spp), xlab="age", ylab="leaf lifespan (leaf births)", cex.axis=.8, las=1,
     main=labels.spp[spp], xlim=c(1,35), ylim=c(0,8), log="x", pch=16)
  points(LL_birth_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8)
  points((LL_birth_mean+LL_birth_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=25)
  points((LL_birth_mean-LL_birth_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=24)
}

#mod_LLvAge <- function(spp) {
#  lm(LL_birth ~ age, data=subset(SummaryInd,species==spp))
#}

for(n in names(labels.spp)) {
  plot_LLvAge(n)
#  mod_LMAvAge(n)
#  summary(mod_LMAvAge(n))
}
```

#Several species show significant declines in shoot extension with age
```{r, echo=FALSE,results='hide', message=FALSE}
plot_newlengthvAge <- function(spp) {
  plot(new_length ~ age, data=subset(SummaryInd,species==spp), xlab="age", ylab="leader shoot extention", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp])
  points(new_length_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8)
  points((new_length_mean+new_length_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=25)
  points((new_length_mean-new_length_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=24)
}

mod_newlengthvAge <- function(spp) {
  lm(new_length ~ age, data=subset(SummaryInd,age>2.4 & species==spp))
}

for(n in names(labels.spp)) {
  plot_newlengthvAge(n)
  mod_newlengthvAge(n)
  summary(mod_newlengthvAge(n))
}
```

#Had expected a stronger pattern either in leaf lifespan or leaf count on a shoot
Apparently it is simply that older plants have more stem - plotting total leaf weight against final weight, shows bigger plants have relatively fewer leaves
(Yes, I know the 2.4 yr olds are missing...)

```{r, echo=FALSE,results='hide', message=FALSE}
plot(total_leaf_weight~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
```
#Next steps?
As I begin to analyze RA patterns, think more specifically about the species that do show a distinct change with age in one of hte "leaf" parameters.

For instance, is there some way to capture "change with age" as a variabe of its own that is worth comparing across species.

Afterall, the energy allocation models suggest that different RA schedules may be due to different amounts of "shifting" across species - some species might be predicted to shift more than others.

#LMA vs Leaf lifespan
This correlation is as expected, except for the babies

```{r, echo=FALSE,results='hide', message=FALSE}
col.LMA <-c("purple", "pink", "light green", "orange","blue", "light blue","red")
labels.LMA <-c("age 1.4", "age 2.4", "age 5", "age 7", "age 9", "age 32", "HATE")
plot(LL_birth_mean ~ LMA_mean, SummarySppAge, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(SummarySppAge, species=="HATE"), pch=17)
legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1.4)
```

#LMA vs LL without young plants
The pattern becomes more obvious when young plants (and Hakea) are omitted

```{r, echo=FALSE,results='hide', message=FALSE}
plot(LL_bd_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE" & age>1.5), pch=16, cex=1.2, col=col.age2[as.factor(age)])
legend("bottomright",col=col.age2,labels.age2,pch=16)
```

#Growth rates increase with increasing LMA in adult plants
But there is no relationship in seedlings (age=1.4) or saplings (age=2.4)
This is not predicted by Daniel's model or Anais' review
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age==1.4),log="xy",col=col.age[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Seedlings")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==1.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")

plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age==2.4),log="xy",col="pink",pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Saplings")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==2.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")

plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Adults")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age>3), labels=species, pos=1, offset=-1.1, cex=0.6,col="blue")
legend("bottomright",col=col.age,labels.age,pch=16)
```

#Species with greater wood density have higher growth rates
Again there is no relationship in younger plants
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ WD_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1,xlab="wood density",ylab="growth investment (mg per year)",xlim=c(0.57,0.9),main="Relationship between wood density and growth rates")
text(GrowthInv_mean ~ WD_mean, data=subset(SummarySppAge,age==5), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
```

#Species with larger seeds have higher growth rates
The same basic relationship exists across all ages, so combined all data here
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ seed_size,data=subset(SummarySppAge,age>.5),log="xy",col=col.age[as.factor(age)],pch=16,cex=1,xlab="seed size (mg)",ylab="growth investment (mg per year)",main="Effect of seed size on growth rates")
legend("topleft",col=col.age, labels.age, pch=16, cex=1)
text(GrowthInv_mean ~ seed_size, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

#Species with greater maximum height have higher growth rates
For younger plants there is no relationship with *SPECIES* overall maximum H, but there is a positive correlation with the maximum height the species reaches at a given age -i.e. fast growing young plants are tall when young.
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ maxH_spp,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="maximum height (mm)",ylab="growth investment (mg per year)",xlim=c(500,3200),main="Relationship between growth rates on maximum H")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
text(GrowthInv_mean ~ maxH_spp, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

#Correlations between different growth measures
Increased cross-sectional area of a leader and the entire plant are fairly correlated and uneffected by age (once not baby)
```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_shoot_area~change_basal_area,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=.8,xlab="change in basal stem area",ylab="change in shoot area")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
```