### 2015-04-17 results.Rmd
Add stub for results file

```{r, echo=FALSE,results='hide', message=FALSE}
# create_bindings()

LMA <- filter(LMA_raw, species!="" & species!=" ") %>%
  select(species,age,LMA,branch_age)

wood <- filter(woodDensity_raw, use=="use") %>%
  select(species,density)

seedsize <- seedSize_raw %>%
  select(species,seed_size)

individuals <- IndividualsList

harvest <- filter(HarvestData,
                  individuals$use_for_allocation_calculations[match(HarvestData$individual, individuals$individual)] & segment==1 & start_end=="end") %>%
  select(species,individual,age,height,total_leaf_weight,site)

investment <- ReproductionAllocation_all

investment$age <- round(investment$age, digits=1)

accessory <- AccessoryCosts_all

```