### 2015-05-20 results_leaf_lifespan.rmd

```{r, echo=FALSE,results='hide', message=FALSE}
library(remake)
create_bindings()

source("scripts/summary_dataframes.R")
```

#Leaf lifespan data
General results:
  - Leaf lifespan increases from seedlings (1.4 yrs) to saplings (2.4 years) to adults (>4 years), but there are no differences between the different adult populations
  - Species differ en

#Generic functions for most plotting below
```{r, echo=FALSE,results='hide', message=FALSE}
plot_yvar_vs_xvar <- function(data, yvar = "y", xvar = "x", ...) {
  Y <- data[[yvar]]
  X <- data[[xvar]]
  plot(Y ~ X, data=data,  cex.axis=.8, las=1, pch=16, ...)
}

summarise_fit <- function(x) {
  data.frame(
    select(glance(x), r.squared, p.value),
    n= length(resid(x)),
    a = coef(x)[1],
    b = coef(x)[2]
  )
}

summarise_all_fits <- function(x) {
  tmp <- lapply(x, summarise_fit) %>%
    bind_rows
  data.frame(species=names(x), tmp)
}
```

#First issue is how to calculate leaf lifespan
A plot of leaf lifespan calculated using birth of leaves, death of leaves, and the mean of the two shows they are poorly correlated, even when the youngest plants are omitted

```{r, echo=FALSE, message=FALSE}
yvar <- "LL_birth"
xvar <- "LL_death"
data <- filter(SummaryInd, age>2, LL_birth<10&LL_death<10)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp,pch=16,col=col.spp,bty="n",cex=.8)
mod <- lm(LL_birth~LL_death + age,data=data)
anova(mod)

yvar <- "LL_birth"
xvar <- "LL_bd"
data <- filter(SummaryInd, age>2, LL_birth<10&LL_death<10)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp,pch=16,col=col.spp,bty="n",cex=.8)
mod <- lm(LL_birth~LL_bd,data=data)
summary(mod)

yvar <- "LL_death"
xvar <- "LL_bd"
data <- filter(SummaryInd, age>2, LL_birth<10&LL_death<10)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp,pch=16,col=col.spp,bty="n",cex=.8)
mod <- lm(LL_death~LL_bd,data=data)
summary(mod)
```

###Starting with the basics: Are leaf lifespan and LMA correlated?  
- Plotting species*site means  
- Figures shown using all three measures of leaf lifespan  
```{r, echo=FALSE,message=FALSE}
yvar <- "LMA_mean"
xvar <- "LL_death_mean"
data <- filter(SummarySppAge)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp,pch=16,col=col.spp,bty="n",cex=.8)
mod <- lm(LL_death_mean~LMA_mean,data=data)
summary(mod)

yvar <- "LMA_mean"
xvar <- "LL_birth_mean"
data <- filter(SummarySppAge)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp,pch=16,col=col.spp,bty="n",cex=.8)
mod <- lm(LL_birth_mean~LMA_mean,data=data)
summary(mod)

yvar <- "LMA_mean"
xvar <- "LL_bd_mean"
data <- filter(SummarySppAge)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp,pch=16,col=col.spp,bty="n",cex=.8)
mod <- lm(LL_bd_mean~LMA_mean,data=data)
summary(mod)
```

###Same plots as above, but omitting HATE, which has much higher LMA than any other species due to its round "leaves"  
- Using leaf deaths gives the poorest fit...
```{r, echo=FALSE,message=FALSE}
yvar <- "LMA_mean"
xvar <- "LL_death_mean"
data <- filter(SummarySppAge, species!="HATE")

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp.noHATE,pch=16,col=col.spp.noHATE,bty="n",cex=.8)
mod <- lm(LL_death_mean~LMA_mean+species,data=data)
summary(mod)
anova(mod)

plot(LL_birth_mean ~ LMA_mean, SummarySppAge, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(SummarySppAge, species=="HATE"), pch=15)
legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1,bty="n")

yvar <- "LMA_mean"
xvar <- "LL_birth_mean"
data <- filter(SummarySppAge, species!="HATE")

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp.noHATE,pch=16,col=col.spp.noHATE,bty="n",cex=.8)
mod <- lm(LL_birth_mean~LMA_mean+species,data=data)
summary(mod)
anova(mod)

yvar <- "LMA_mean"
xvar <- "LL_bd_mean"
data <- filter(SummarySppAge, species!="HATE")

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar,col=col.spp[as.factor(data$species)])
legend("bottomright",legend=labels.spp.noHATE,pch=16,col=col.spp.noHATE,bty="n",cex=.8)
mod <- lm(LL_bd_mean~LMA_mean+species,data=data)
summary(mod)
anova(mod)

yvar <- "LMA_mean"
xvar <- "LL_death_mean"
data <- filter(SummarySpp)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar)
text(LMA_mean ~LL_death_mean, data=data, labels=SummarySpp$species, pos=1, offset=-1.1, cex=0.9,col="blue")
mod <- lm(LL_death_mean~LMA_mean,data=subset(data,species!="HATE"))
summary(mod)
```

#How does leaf lifespan change with age and size across all species and within each species
Using leaf deaths for the following plots and removing the youngest individuals

Age is log transformed.
BAER, HEPU, PUTU show a significant DECREASE in leaf lifespan with increasing age (p<0.05)
LEES, PEPU shows a marginally significant DECREASE in accessory costs with increasing age (0.05<p<0.10)
```{r, echo=FALSE,message=FALSE}
yvar <- "LL_death"
xvar <- "age"
data <- filter(SummaryInd,LL_death<10&age>2)

data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(x[[yvar]] ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=3)
results$b <- round(results$b,digits=3)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar, 
                    main=spp, log="x", xlab=xvar, ylab=yvar)

  abline(lmfits[[spp]])
}
```

And using size...
Size is log transformed.
BAER, COER, EPMI show a significant DECREASE in leaf lifespan with increasing size (p<0.05)
PEPU, PHPH show a marginally significant DECREASE in accessory costs with increasing size (0.05<p<0.10)

If leaf lifespan based on leaf births is used instead, there are a number of significant INCREASES in leaf lifespan with increasing size
```{r, echo=FALSE,message=FALSE}
yvar <- "LL_death"
xvar <- "total_weight"
data <- filter(SummaryInd,LL_death<8&age>2)

data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(x[[yvar]] ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=3)
results$b <- round(results$b,digits=3)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,
                    main=spp, log="x", xlab=xvar, ylab=yvar)

  abline(lmfits[[spp]])
}
```

#Higher growth rates correlated with lower leaf lifespan in a handful of species
BAER, EPMI, PEPU show declining LL and COER shows marginally significant decline with increasing absolute growth investment.

```{r, echo=FALSE,message=FALSE}
yvar <- "LL_death"
xvar <- "GrowthInv"
data <- filter(SummaryInd,LL_death<8&age>2)

data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(x[[yvar]] ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=3)
results$b <- round(results$b,digits=3)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,
                    main=spp, log="x", xlab=xvar, ylab=yvar)

  abline(lmfits[[spp]])
}
```

#Putting age and speices together in a model shows, not surprisingly, that both age and species contribute to the variation and that species matters much more.
```{r, echo=FALSE,message=FALSE}
plot(LL_death ~ age, data=SummaryInd, cex.axis=.8, las=1, xlim=c(1,35), ylim=c(.95,10), log="xy", pch=16,col=col.spp[as.factor(species)])
legend(14,8,legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mod <- lm(LL_death~age*species, data=SummaryInd)
summary(mod)
anova(mod)
```

#Plotting leaf lifespan against the number of leaves on a branch shows that for the babies, "LL_birth" is negatively correlated with the number of leaves - i.e. the shoots grow a lot during the year and the more they grow the lower the LL, because there are so few leaves to begin with. Indicates that for "LL_birth", but not "LL_death", the youngest site should be omitted

However for LL_death most species show a positive correlation between number of leaves on the leader and leaf lifespan. For some species this is driven by a few individuals with high LL_death values, but 6 species still significant when LL > 6 removed (and 4 more species near-significant).

This trend worries me - is leaf count very variable across years?

```{r, echo=FALSE,message=FALSE}
plot(LL_death ~ lvs_end_total, data=subset(SummaryInd), log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(LL_birth ~ lvs_end_total, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(LL_death ~ lvs_end_total, SummaryInd, log="xy",col=col.age[as.factor(age)],pch=16)
legend("topleft",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")

plot(LL_birth ~ lvs_end_total, data=subset(SummaryInd), log="xy",col=col.age[as.factor(age)],pch=16)
legend("topleft",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")

mod <- lm(LL_death ~ lvs_end_total*species*age, SummaryInd)
summary(mod)
anova(mod)
step(mod)

mod2 <- lm(formula = LL_death ~ lvs_end_total + species + age + lvs_end_total:species + 
    species:age, data = SummaryInd)
summary(mod2)
anova(mod2)

yvar <- "LL_death"
xvar <- "lvs_end_total"
data <- filter(SummaryInd,LL_death<6)

data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(x[[yvar]] ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=3)
results$b <- round(results$b,digits=3)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,
                    main=spp, log="x", xlab=xvar, ylab=yvar)

  abline(lmfits[[spp]])
}
```

#LMA vs. age
- older plants have marginally higher LMA, but the effect mostly disappears when the youngest plants are excluded  
- there is enormous variation in LMA across species  
- also had plot of just branches from 2nd-last leaf cohort - no real difference, so using mean LMA of up to 3 leaf cohorts

```{r, echo=FALSE,message=FALSE}

plot(LMA_mean~age, SummarySppAge,log="xy", col=col.spp[as.factor(species)], pch=16)
mod <- lm(LMA_mean~species*age, SummarySppAge)
summary(mod)
anova(mod)

mod <- lm(LMA_mean~species*age, data=subset(SummarySppAge,age>2))
summary(mod)
anova(mod)

plot(LMA_mean ~age, SummarySppAge, log="xy", col=col.spp[as.factor(species)], pch=16)
#text(LMA_mean~age, data=subset(SummarySppAge,age==5), labels=SummarySpp$species, pos=4, offset=.5, cex=0.7,col="blue")
legend(12,.04,col=col.spp, pch=16,legend=labels.spp,cex=.8,bty="n")
```

```{r, echo=FALSE,results='hide', message=FALSE}
plot(LMA~age,data=subset(LMA,branch_age=="mid"),log="xy", xlab="age",ylab="LMA",col=col.spp[as.factor(species)],pch=16)
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

#Plotting LMA vs. weight
There is a stronger trend with respect to weight, even after the effect of species is taken into account

The interactions terms weren't significant and were removed from the model
```{r, echo=FALSE,message=FALSE}
mod <- lm(LMA_mean~species+log(total_weight_mean), SummarySppAge)
summary(mod)
anova(mod)

mod <- lm(LMA_mean~species+log(total_weight_mean), data=subset(SummarySppAge,age>2))
summary(mod)
anova(mod)

plot(LMA_mean ~total_weight_mean, SummarySppAge, log="xy", col=col.spp[as.factor(species)], pch=16)
legend("topleft",col=col.spp, pch=16,legend=labels.spp,cex=.8,bty="n")

plot(LMA_mean ~total_weight_mean, data=subset(SummarySppAge), log="xy", col=col.age[as.factor(age)], pch=16, cex=1)
legend("topleft",col=col.age, pch=16,legend=labels.age,cex=.8,bty="n")

```

#PropH vs LL_death
- Have just plotted overall data, because no relationships across or within species; only GRSP has significant negative relationship and BOLE has significant positive relationship
```{r, echo=FALSE,message=FALSE}
plot((LL_death~PropH), data=subset(SummaryInd,LL_death<10), col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot((PropH~log(age)), data=subset(SummaryInd), col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
```

```{r, echo=FALSE,message=FALSE}

```