### 2015-05-26 results_accessory_shorter.Rmd
lots of work on growth data today, bit of continued cleaning on accessory costs

```{r, echo=FALSE,results='hide', message=FALSE}
library(remake)
create_bindings()

source("scripts/summary_dataframes.R")
source("scripts/functions.R")
```

#Accessory costs
##Basic questions
* Total reproductive investment can be divided into energy invested in propagules vs. other reproductive tissues (accessory costs)
* Reproductive investment can be considered on an energy or "count" (fecundity) basis

    1. What reproduction "pool" should one measure?
      * Is fecundity good enough?
      * Is fruit or propagule weight good enough?
    2. What might cause shifts in accessory costs across species?
      * Previous studies have shown smaller-seeded species have lower accessory costs.
    3. How does alotment to different accessory tissue types shift within a species as it ages, grows, or increases RA? 
      * If accessory costs decrease with increased RA, increasing RA yields compounding returns.

##Summary of outcomes
1. Total reproductive investment is by far the best correlate with total plant size (or growth investment)
2. Plants with smaller seeds, smaller adult sizes, and higher seedset have lower accessory costs. It is mathematically plausable that the concept of "accessory costs" is not all that different from "seed set"
  * In other words, plants with higher seed set, have lower accessory costs due to calculations
  * It is nice to have a data set that shows, across 14 species, that larger plants have lower seed set.
3. Bigger plants have higher RA. 
  * Plants that produce more buds, invest more absolute energy into either pre-pollination tissues or propagules also have higher RA.
  * As plants get bigger, there is a slight shift of energy from pre-pollination tissues to propagules, but this is a small effect 
  *(i.e. accessory costs decline slightly with increasing plant size)
4. Among reproducing individuals, bigger or older plants DO NOT have higher seedset, although more individuals become reproductively mature over time. 
  * Proportional pre-pollination costs also decrease with plant age in most species
5. Species differ enormously in their relative investment in different reproductive tissue types

##Still thinking about
1. Most analyses look at propagules and pre-pollination tissues, but allocation can be further divided - it would be interesting to look at proportion energy to aborted buds vs. flowers (I do know aborted buds decrease)
###Which measure of reproductive output best correlates with plant size or growth (growth plotted here)
* Measuring total reproductive output has a much stronger correlation than any of the other measures
* Note that each of these figures (and models) only includes individuals with non-zero reproductive investment

```{r, echo=FALSE,message=FALSE,,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(.6,0.3,.3,0.3), mai=c(.8,1,.3,0.3))
mod <- lm (log(total_weight)~log(ReproInv), data=subset(SummaryInd,ReproInv>0))
plot(ReproInv~total_weight,data=subset(SummaryInd,ReproInv>0), log="xy",xlab="", col=col.spp[as.factor(species)], pch=16, ylab="Total reproductive investment (mg)",xlim=c(100,2000000))
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(fruit_weight), data=subset(SummaryInd,fruit_weight>0))
plot(fruit_weight~total_weight,data=subset(SummaryInd,fruit_weight>0), log="xy",col=col.spp[as.factor(species)], pch=16, ylab="Fruit weight (mg)",xlim=c(100,2000000),xlab="")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(propagule_inv), data=subset(SummaryInd,propagule_inv>0))
plot(propagule_inv~total_weight,data=subset(SummaryInd,propagule_inv>0), log="xy",col=col.spp[as.factor(species)], pch=16, xlab="", ylab="Propagule investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(seedset), data=subset(SummaryInd,seedset>0))
plot(seedset~total_weight,data=subset(SummaryInd,prop_propagule>0), log="xy",col=col.spp[as.factor(species)], pch=16, ylab="seedset",xlim=c(100,2000000),xlab="")
words.bottom.right.logxy(mod)
mtext("Total size (mg)", 1, outer=TRUE,cex=1.2)
```

### How do accessory costs shift with seed size?
- Larger-seeded species have higher accessory costs, as has been shown in other studies
- However, it is also true that "larger species" have higher accessory costs and than species with higher seed set have lower accessory costs
- In contrast, there is a relatively low correlation between accessory costs and proportion reproductive energy invested in pre-pollination tissues DESPITE this being the biggest pool of energy for most species
- Taken together these data (and various other plots, not shown) suggest that accessory costs (as a proportion) are most strongly driven by seedset
- In other words, if flowers are pollinated and develop into mature seeds, the plants ends up investing more energy in reproduction and that extra energy is being invested into propagules.
- Therefore, seed set goes up and by default, a greater proportion of energy goes into propagules and less into "accessory costs"

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(0.4,0.6,0.3,0.3), mai=c(1.1,0.8,0.3,0.3))
mod <- lm(log(seed_size)~asin(sqrt(prop_accessory_mean)),SummarySpp)
plot(asin(sqrt(prop_accessory_mean))~seed_size,SummarySpp, log="x", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="seed size (mg)",cex=1.3)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logx(mod)

mod <- lm(log(total_weight_mean)~asin(sqrt(prop_accessory_mean)),SummarySpp)
plot(asin(sqrt(prop_accessory_mean))~total_weight_mean,SummarySpp, log="x", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="Total weight (mg)",cex=1.3)
words.bottom.right.logx(mod)

mod <- lm(log(seedset_mean)~asin(sqrt(prop_accessory_mean)),SummarySpp)
plot(asin(sqrt(prop_accessory_mean))~seedset_mean,SummarySpp, log="x", pch=16,col=col.spp[as.factor(species)],ylab="",cex=1.3)
words.top.right.logx(mod)

mod <- lm(asin(sqrt(prop_prepollen_all_mean))~asin(sqrt(prop_accessory_mean)),SummarySpp)
plot(asin(sqrt(prop_accessory_mean))~asin(sqrt(prop_prepollen_all_mean)),SummarySpp, log="", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="pre-pollination costs (proportional)",cex=1.3)
words.top.right(mod)
mtext("Proportion reproductive investment to accessory tissues (asin transformed)",2, outer=TRUE,cex=1.2)
```

* investment vs. costs per successful seed
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(0.4,0.6,0.3,0.3), mai=c(1.1,0.8,0.3,0.3)) 
mod <- lm(log(seed_size)~(accessory_per_seed_mean),SummarySpp)
plot(accessory_per_seed_mean~seed_size,SummarySpp, log="xy", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="seed size (mg)",cex=1.3)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm(log(total_weight_mean)~log(accessory_per_seed_mean),SummarySpp)
plot(accessory_per_seed_mean~total_weight_mean,SummarySpp, log="xy", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="Total weight (mg)",cex=1.3)
words.bottom.right.logxy(mod)

mod <- lm(log(seedset_mean)~log(accessory_per_seed_mean),SummarySpp)
plot((accessory_per_seed_mean)~seedset_mean,SummarySpp, log="xy", pch=16,col=col.spp[as.factor(species)],ylab="",cex=1.3)
words.top.right.logxy(mod)

mod <- lm(log(prepollen_all_per_seed_mean)~log(accessory_per_seed_mean),SummarySpp)
plot(log(accessory_per_seed_mean)~log(prepollen_all_per_seed_mean),SummarySpp, log="", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="pre-pollination costs (proportional)",cex=1.3)
words.top.right(mod)
mtext("Accessory tissue investment per successful seed (mg)",2, outer=TRUE,cex=1.2)
```

### Plotting accessory costs that go into making a single seed - ignoring aborted tissues
* Bigger seeds have bigger fixed costs of making that seed (1st plot)
* But other than LEES having much lower costs, no shifts across species in their relative costs to make a seed (2nd plot)
* In other words, patterns of accessory costs vs. seed size, plant size, etc come about due to the differential effect of bud, flower abortion across species  
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2))
plot(seed_size~seedcosts,data=subset(SummarySpp),log="xy",pch=16,col=col.spp[as.factor(species)],xlab="costs to make 1 seed (mg)",cex=1.2)
mod <- lm(seedcosts~seed_size, SummarySpp)
words.bottom.right.logxy(mod)
text(seed_size~seedcosts,data=subset(SummarySpp),labels=species, pos=1,offset=.6,cex=0.8,col="black")

plot(seed_size~costs_per_seed,data=subset(SummarySpp),log="y",pch=16,col=col.spp[as.factor(species)],ylab="",xlab="proportional accessory costs, scaled by seed weight",cex=1.2)
text(seed_size~costs_per_seed,data=subset(SummarySpp),labels=species, pos=1,offset=.6,cex=0.8,col="black")
mod <- lm(costs_per_seed~seed_size, SummarySpp)
words.bottom.left.logy(mod)
```

* Here is the plot just of seed set by propagule investment (for all individuals), showing the effect is true within and across species
* This is the complement to the plot of accessory costs  
* Replotting the same data by age, shows that the pattern is not age-driven
* At most, there are most "old" plants with very low seedset 
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(0.6,0.3,0.3,0.3), mai=c(0.8,1,0.3,0.3))  
mod <- lm(log(seedset)~log(prop_propagule),data=subset(SummaryInd,seedset>0&seedset<1))
plot(prop_propagule~seedset,data=subset(SummaryInd,seedset>0&seedset<1), log="xy", pch=16,col=col.spp[as.factor(species)],xlab="",ylab="proportion reproduction to propagules (1-accessory)")
words.bottom.right.logxy(mod)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mod <- lm(log(seedset)~log(prop_propagule),data=subset(SummaryInd,seedset>0&seedset<1))
plot(prop_propagule~seedset,data=subset(SummaryInd,seedset>0&seedset<1), log="xy", pch=16,col=col.age[as.factor(age)],ylab="proportion reproduction to propagules (1-accessory)",xlab="")
words.bottom.right.logxy(mod)
legend("topleft",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")
mtext("Seed set",1, outer=TRUE,cex=1.1)
```

*Comparing proportion energy to propagules vs accessory cost per seed
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(0.6,0.3,0.3,0.3), mai=c(0.8,1,0.3,0.3))  
mod <- lm(log(seedset)~log(prop_propagule),data=subset(SummaryInd,seedset>0&seedset<1))
plot(prop_propagule~seedset,data=subset(SummaryInd,seedset>0&seedset<1), log="xy", pch=16,col=col.spp[as.factor(species)],xlab="",ylab="proportion reproduction to propagules (1-accessory)")
words.bottom.right.logxy(mod)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mod <- lm(log(seedset)~log(accessory_per_seed),data=subset(SummaryInd,seedset>0&seedset<1&accessory_per_seed>0))
plot(accessory_per_seed~seedset,data=subset(SummaryInd,seedset>0&seedset<1), log="xy", pch=16,col=col.spp[as.factor(species)],xlab="",ylab="accessory costs per seed produced (mg)")
words.bottom.left.logxy(mod)
mtext("Seed set",1, outer=TRUE,cex=1.1)
```

###Relationship between tissue allocation and reproductive investment
* these data also tell the story of seedset being an important "component" and being lower in bigger species
* Proption pre-pollination costs completely unrelated to reproductive investment
* However, proportional investment in propagules decreases with increasing reproductive investment, because the bigger species investment more overall in reproduction and less in propagules
* There is also no pattern with respect to proportional investment in packaging and dispersal (R^2=0, p-value=0.959) and a slight decrease in aborted post-pollination tissues with increasing reproductive investment (R^2=0.0661, p-value=0.0002)
```{r, echo=FALSE,message=FALSE, ,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(0.6,0.3,0.1,0.3), mai=c(0.8,1,0.3,0.3))  
mod <- lm(log(ReproInv)~(prop_prepollen_all),data=subset(SummaryInd,prop_prepollen_all>0&prop_prepollen_all<1))
plot((prop_prepollen_all)~ReproInv,data=subset(SummaryInd,prop_prepollen_all>0&prop_prepollen_all<1), log="x", pch=16,col=col.spp[as.factor(species)],xlab="",ylab="Pre-pollination tissues (as proportion repro inv)")
words.top.right.logx(mod)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mod <- lm(log(ReproInv)~log(prop_propagule),data=subset(SummaryInd,prop_propagule>0))
plot(prop_propagule~ReproInv,data=subset(SummaryInd,prop_propagule>0), log="xy", pch=16,col=col.spp[as.factor(species)],xlab="",ylab="Propagules (as proportion repro inv)")
words.top.right.logxy(mod)
mtext("Reproductive investment (mg)",1, outer=TRUE,cex=1.1)
```

###What happens against RA
* Overall RA is correlated with reproductive investment and age
* These plots only include individuals that have some reproductive investment
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(0.3,0.3,0.1,0.3), mai=c(1.1,1,0.3,0.3)) 
mod <- lm(log(RA)~log((ReproInv)),data=subset(SummaryInd,RA>0))
plot(RA~(ReproInv),data=subset(SummaryInd,RA>0), pch=16,col=col.spp[as.factor(species)],log="xy")
words.bottom.right.logxy(mod)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mod <- lm(log(RA)~species*age*log((ReproInv)),data=subset(SummaryInd,RA>0))
anova(mod)

mod <- lm(log(RA)~log((age)),data=subset(SummaryInd,RA>0))
plot(RA~(age),data=subset(SummaryInd,RA>0), pch=16,col=col.spp[as.factor(species)],log="xy",ylab="")
words.bottom.right.logxy(mod)
```

* RA is also correlated with count of floral units - the total number of buds produced (i.e. final development stage can be buds, flowers, aborted fruit or fruit)
* This is true for many individual species as well as across species
* RA and seed count are also correlated, but mainly that no low RA individuals have high seed counts
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(0.6,0.3,0.1,0.3), mai=c(0.8,1,0.3,0.3))  
mod <- lm(log(RA)~log(repro_all_count),data=subset(SummaryInd,repro_all_count>0))
plot(repro_all_count~RA,data=subset(SummaryInd,repro_all_count>0),col=col.spp[as.factor(species)],pch=16,log="xy",ylab="# of buds produced",xlab="")
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm(log(RA)~log(seed_count+.1),data=subset(SummaryInd,repro_all_count>0))
plot((.1+seed_count)~RA,data=subset(SummaryInd,repro_all_count>0),col=col.spp[as.factor(species)],pch=16,log="xy",ylab="# of seeds produced",xlab="")
words.bottom.right.logxy(mod)
mtext("RA",1, outer=TRUE,cex=1.1)

yvar <- "repro_all_count"
xvar <- "RA"

data <- subset(SummaryInd, repro_all_count>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm((log(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

* This same pattern is true just looking  at energy invested in pre-pollination tissues and at propagule allocation
* Plants that invest more in some reproductive tissues, on average, invest more in other reproductive tissues, so have a higher RA

* Looking at proportions, the pattern is much less informative
* Plants with higher RA have lower allocation to pre-pollination tissues, probably because of their higher seedset
* Relationship between RA and proportion energy to propagules simply shows a species effect - different species have different RA values and different proportional propagule investment
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2))  
mod <- lm(log(RA)~log((prepollen_all_inv)),data=subset(SummaryInd,RA>0&prepollen_all_inv>0))
plot(prepollen_all_inv~(RA),data=subset(SummaryInd,RA<1&RA>0), pch=16,col=col.spp[as.factor(species)],log="xy",ylab="pre-pollination tissues (mg)")
words.bottom.right.logxy(mod)

legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
mod <- lm(log(RA)~log((propagule_inv)),data=subset(SummaryInd,RA>0&propagule_inv>0))
plot(propagule_inv~(RA),data=subset(SummaryInd,RA>0&propagule_inv>0), pch=16,col=col.spp[as.factor(species)],log="xy", ylab="energy to propagules (mg)")
words.bottom.right.logxy(mod)

mod <- lm(asin(sqrt(RA))~asin(sqrt(prop_prepollen_all)),data=subset(SummaryInd,RA>0&prop_prepollen_all>0))
plot(asin(sqrt(prop_prepollen_all))~asin(sqrt(RA)),data=subset(SummaryInd,RA>0&prop_prepollen_all>0), pch=16,col=col.spp[as.factor(species)], ylab="pre-pollination tissues (as proportion repro inv) (asin)")
words.top.right(mod)

mod <- lm(log(RA)~log((prop_propagule)),data=subset(SummaryInd,RA>0&prop_propagule>0))
plot((prop_propagule)~(RA),data=subset(SummaryInd,RA>0&prop_propagule>0), log="xy", pch=16,col=col.spp[as.factor(species)],ylab="propagules (as proportion repro inv)")
words.bottom.left.logxy(mod)
```

###Shifts across age
* Seedset shows a slight decrease with proportional plant height (the best proxy for "relative size")  
* this pattern likely exists because species that begin reproducing at shorter proportional heights have higher seedset  
* only LEES shows a significant decline with age, although most species have a segative slope  
* there is a similar pattern for Propagule investment (not shown)  
* Table shows results, by species, for seedset by proportional height  
```{r, echo=FALSE,message=FALSE}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2))  
mod <- lm((prop_maxH)~asin(sqrt(seedset)),data=subset(SummaryInd,seedset<1&seedset>0))
plot(asin(sqrt(seedset))~prop_maxH,data=subset(SummaryInd,seedset<1&seedset>0), pch=16,col=col.spp[as.factor(species)])
words.top.right(mod)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

yvar <- "seedset"
xvar <- "prop_maxH"

data <- subset(SummaryInd, seedset>0&seedset<1)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

* Results, by species, for seedset by age  
* Only HEPU, LEES show significant decline, but again mostly negative slopes
```{r, echo=FALSE,message=FALSE}
yvar <- "seedset"
xvar <- "age"

data <- subset(SummaryInd, seedset>0&seedset<1)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

* A number of species show declines in accessory costs with age, but this effect disappears once you include only individuals that set *any* seed
```{r, echo=FALSE,message=FALSE}
yvar <- "prop_accessory"
xvar <- "age"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

* 8 of the 14 species show declines in allocation to pre-pollination tissues with increasing age
```{r, echo=FALSE,message=FALSE}
yvar <- "prop_prepollen_all"
xvar <- "age"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

##Shifts in proportion to tissue types across species  
- Reasonably idiosyncratic patterns
- COER, EPMI, LEES, PILI invest the most energy in propagules  
- HATE invests by far the least into propagules  
```{r, echo=FALSE, message=FALSE, fig.width=14, fig.height=28}
par(mfrow=c(4,2), cex=1, omi=c(1,.1,.7,.1), mai=c(.5,1.1,.2,0.2))  
mean1 <- SummarySpp$prop_propagule_mean
se1 <- SummarySpp$prop_propagule_se
plot(mean1~species,SummarySpp,cex.axis=.5,pch=16,log="y",ylim=c(.002,.25),ylab="propagules",xlab="")
se_function(mean1,se1)

mean1 <- SummarySpp$propagule_per_seed_mean
se1 <- SummarySpp$propagule_per_seed_se
plot(mean1~species,SummarySpp,cex.axis=.5,pch=16,log="y",ylab="propagules",xlab="")
se_function(mean1,se1)

mean1 <- SummarySpp$prop_prepollen_all_mean
se1 <- SummarySpp$prop_prepollen_all_se
plot(mean1~species,data=subset(SummarySpp),cex.axis=.5,pch=16,ylim=c(.1,1),ylab="pre-pollination tissues",xlab="")
se_function(mean1,se1)

mean1 <- SummarySpp$prepollen_all_per_seed_mean
se1 <- SummarySpp$prepollen_all_per_seed_se
plot((.01+prepollen_all_per_seed_mean)~species,data=subset(SummarySpp),cex.axis=.5,pch=16,ylab="pre-pollination tissues",xlab="",log="y")
se_function(mean1,se1)

mean1 <- SummarySpp$prop_packaging_dispersal_mean
se1 <- SummarySpp$prop_packaging_dispersal_se
plot(mean1~species,SummarySpp,cex.axis=.5,pch=16,log="y",ylim=c(.02,1),ylab="packaging and dispersal")
se_function(mean1,se1)

mean1 <- SummarySpp$packaging_dispersal_per_seed_mean
se1 <- SummarySpp$packaging_dispersal_per_seed_se
plot((packaging_dispersal_per_seed_mean+.1)~species,SummarySpp,cex.axis=.5,pch=16,ylab="packaging and dispersal",log="y")
se_function(mean1,se1)

mean1 <- SummarySpp$prop_postpollen_aborted_mean
se1 <- SummarySpp$prop_postpollen_aborted_se
plot(mean1~species,SummarySpp,cex.axis=.5,pch=16,log="y",ylim=c(.02,1),ylab="aborted post-pollination tissues")
se_function(mean1,se1)

mean1 <- SummarySpp$postpollen_aborted_per_seed_mean
se1 <- SummarySpp$postpollen_aborted_per_seed_se
plot((postpollen_aborted_per_seed_mean+.1)~species,SummarySpp,cex.axis=.5,pch=16,ylab="aborted post-pollination tissues",log="y")
se_function(mean1,se1)

mtext("proportion of reproductive investment                                                                       mg investment per seed",3,outer=TRUE,cex=1.1)
mtext("species",1, outer=TRUE,cex=1.1)
```

- just looking among plants >3 years old  
```{r, echo=FALSE, message=FALSE, fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(1,.1,.7,.1), mai=c(.5,1.1,.2,0.2)) 
tmp <- SummaryInd %>%
  filter(ReproInv!=0&age>3) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se), prop_prepollen_all)

plot(mean~species,
    tmp,
     cex.axis=.5,pch=16,log="y",ylim=c(.02,1),
     ylab="Prop reproductive investment put into pre-pollination tissues")

mean1 <- tmp$mean
se1 <- tmp$se
se_function(mean1,se1)

tmp2 <- SummaryInd %>%
  filter(ReproInv!=0&age>3) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se), prop_packaging_dispersal)

plot(mean~species,
    tmp2,
     cex.axis=.5,pch=16,log="y",ylim=c(0.08,.8),
     ylab="Prop reproductive investment put into packaging & dispersal tissues")

mean1 <- tmp2$mean
se1 <- tmp2$se

se_function(mean1,se1)
```

#Shifts in accessory costs across species and ages
##Overall results
- Accessory costs higher for bigger seeds

- Accessory costs higher for bigger species; these are the same species that invest less in RA

- Of course there is also a correlation between seed size and plant size, but actually not as strong as the accessory cost one, presumably because older plants have lower accessory costs as well as species that are inherently bigger

- A number of species show fairly consistent declines in accessory costs with various measures of "bigger", including size, age, RA. Other species show no patterns  
- Lower pre-pollination costs in plants with higher RA: 9 (of 14) species decrease the proportion of reproductive investment in "pre-pollination" tissues as their RA increases. in other words, they get better returns on investment the more they invest. (This is just a pattern within, not across, species)  
- Species very different in which accessory tissues they invest the most in - which reflects the differences in basic structure (i.e. Banksia cones), but also more subtlely that some species abort lots of buds (pre-pollination aborted), while other species abort more fully formed flowers  
- Haven't actually run all the tests, but assume all reproductive tissue types will differ in proportional allocation across species
- For most of the statistical analyses, the youngest site has been removed, since it included individuals that were harvested while flowering (and therefore never "got to" set fruit) and automatically have no "successful" reproduction. There are of course some individuals like this at other ages as well, but far fewer