### 2015-07-30 error_checking.rmd
mostly addition of "error_checking.rmd" which should show problems with leaf weight vs diameter

```{r, echo=FALSE,results='hide', message=FALSE}
library(remake)
create_bindings()

source("scripts/summary_dataframes.R")
source("scripts/functions.R")
```

* you can see that for all plants, even old ones, if just the upper segments are used, the leaf weight vs. diameter still plots on the main line, but if the lower stem segment(s), where leaves are few or lacking (even on side branches), there is increased diameter with little increase in leaf weight. 
* this means that for the plots, segment 1 (which are mostly the labeled dots) fall "above" the main line of points

```{r, echo=FALSE, message=FALSE, fig.width=7, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2))
check <- BOLE_HarvestData

plot(dia~stem_weight,data=subset(check), pch=16,log="xy",col=col.lots[as.factor(segment)])
text(dia~stem_weight,data=subset(check,segment==1),labels=individual,cex=.5,pos=3,offset=-.7)

plot(dia~leaf_weight,data=subset(check), pch=16,log="xy",col=col.lots[as.factor(segment)])
text(dia~leaf_weight,data=subset(check,segment==1),labels=individual,cex=.5,pos=3,offset=-.7)

#colors represent segments, with red the entire plant, then purple, gray, green
#for stems, linear no matter what portion plant included
#for leaves, with older plants (ages 5,7,9 for BOLE, older for other species) have continued diameter increase with little increase in leaf area

plot(dia~stem_weight,data=subset(check,age==7), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~stem_weight,data=subset(check,segment==1&age==7),labels=individual,cex=.5,pos=3,offset=-.7)

plot(dia~leaf_weight,data=subset(check,age==7), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~leaf_weight,data=subset(check,segment==1&age==7),labels=individual,cex=.5,pos=3,offset=-.7)

#second set of plots just fpr 7 year old plants
#colors now by individual, with labels on "whole plant" measures

check <- PELA_HarvestData

plot(dia~stem_weight,data=subset(check,age==32), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~stem_weight,data=subset(check,segment==1&age==32),labels=individual,cex=.5,pos=3,offset=-.7)

plot(dia~leaf_weight,data=subset(check,age==32), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~leaf_weight,data=subset(check,segment==1&age==32),labels=individual,cex=.5,pos=3,offset=-.7)

#at the other extreme, with PELA, effect only seed among age 32 plants, and only among 5 of the 7 reps (colors here represent individuals)

check <- HATE_HarvestData

plot(dia~stem_weight,data=subset(check,age==32), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~stem_weight,data=subset(check,segment==1&age==32),labels=individual,cex=.5,pos=3,offset=-.7)

plot(dia~leaf_weight,data=subset(check,age==32), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~leaf_weight,data=subset(check,segment==1&age==32),labels=individual,cex=.5,pos=3,offset=-.7)

#and no asymptoting at all for HATE

check <- COER_HarvestData

plot(dia~stem_weight,data=subset(check,age>5), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~stem_weight,data=subset(check,segment==1&age>5),labels=individual,cex=.5,pos=3,offset=-.7)

plot(dia~leaf_weight,data=subset(check,age>5), pch=16,log="xy",col=col.spp[as.factor(individual)])
text(dia~leaf_weight,data=subset(check,segment==1&age>5),labels=individual,cex=.5,pos=3,offset=-.7)

#COER here

```