### 2015-08-18 results_growth_short.Rmd

 ```{r, echo=FALSE,results='hide', message=FALSE}
library(remake)
create_bindings()

source("scripts/summary_dataframes.R")
source("scripts/functions.R")

```
##Growth data, general patterns
1. Growth-reproduction trade-off illustrated by plots of RA vs RGR
    -  As expected, a cost of reproduction is reduced growth
    -  Meanwhile there is the strong positive relationship between growth investment and reproductive investment, because bigger plants grow and reproduce more.  
    -  It is good to have a big enough data set to capture both trends across a single dataset - and to capture them consistently across many species

2. Some species have more obvious asymptoting in growth with increasing relative H ("prop_maxH") than others
    -  The 4 "canopy" species all tend to have high correlations between "prop maxH" and growth and reproductive investment (both relative and absolute)  (HATE, BAER, PELA, PEPU)
    -  The lower-growing species are more variable, with "prop maxH"" mattering more for some than others - not surprising since some species like BOLE, PILI show inconsistent     increases in H with increasing age, due to breakage, shedding

##Growth-reproduction trade-off
###Bigger plants reproduce more
- There is a positive correlation between reproductive investment vs growth investment among individuals that are reproducing: plants that grow more also allocate more to reproduction.  
- It is also true that bigger plants allocate more to reproduction.

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7.5}
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(log(GrowthInv)~log(ReproInv),data=subset(SummaryInd,ReproInv>0))
plot((1+ReproInv)~GrowthInv, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
words.top.left.logxy(mod)
legend("bottomright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5")
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75")
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9")
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95")
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1")
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01")

plot((1+ReproInv)~GrowthInv, SummaryInd, log="xy", col=col.age[as.factor(age)],pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
legend("bottomright",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5")
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75")
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9")
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95")
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1")
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01")
```

###Plants that allocate more to reproduction (higher RA) have lower growth
- When looked at individually, this pattern holds for most species, although there is lots of variation in the exact trade-off pattern displayed by different species  
    * when plotted again stem diameter increment, most species show a distinctly linear pattern - decline in stem diameter increment fairly exactly offset by increases in RA across all individuals of the species   
    * for plots of RGR vs RA some species have a distinct "knickpoint" where growth impacted -i.e. BOLE, EPMI, HATE

```{r, echo=FALSE,message=FALSE, warning=FALSE}
par(mfrow=c(2,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(RGR~asin(sqrt(RA)),data=subset(SummaryInd,RA>0))
plot((RA+.01)~RGR, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logxy(mod)

mod <- lm(log(growth_stem_diam)~asin(sqrt(RA)),data=subset(SummaryInd,RA>0))
plot((RA+.01)~growth_stem_diam, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logxy(mod)

mod <- lm(log(prop_growth_stem_diam)~asin(sqrt(RA)),data=subset(SummaryInd,RA>0))
plot((RA+.01)~prop_growth_stem_diam, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logxy(mod)

SummaryInd$prop_growth_height <- SummaryInd$growth_height/SummaryInd$maxH_spp

mod <- lm(log(prop_growth_height)~asin(sqrt(RA)),data=subset(SummaryInd,growth_height>0&species!="PEPU"&species!="PELA"&species!="COER"))
plot((RA+.01)~(prop_growth_height), data=subset(SummaryInd,species!="PEPU"&species!="PELA"), log="xy", col=col.spp[as.factor(species)],pch=16)
```
```{r, echo=FALSE,message=FALSE, warning=FALSE}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "growth_stem_diam"
xvar <- "RA"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  asin(sqrt(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="y",
                    main=spp)
}

mtext("RA", 1, outer=TRUE,cex=1.2)
mtext("Yearly increase in stem diameter (mm)", 2, outer=TRUE,cex=1.2)
```

```{r, echo=FALSE,message=FALSE, warning=FALSE}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "RA"
xvar <- "age"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",
                    main=spp,xlim=c(1,35))
}

mtext("RA", 1, outer=TRUE,cex=1.2)
mtext("age", 2, outer=TRUE,cex=1.2)
```

```{r, echo=FALSE,message=FALSE, warning=FALSE}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "RA"
xvar <- "diameter"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",
                    main=spp)
}

mtext("RA", 1, outer=TRUE,cex=1.2)
mtext("basal diameter", 2, outer=TRUE,cex=1.2)
```

###PCA analysis indicates the same two patterns
* This combination of variables shows that RGR is approximately opposite a collection of variables that indicate a plant is "bigger" - so smaller plants have higher RGR
* Orthoganal to these in WD vs. maxH, LMA_mean, seed_size, and shoot_growth - species that are taller as adults have bigger seeds, higher LMA, and greater shoot growth (both stem growth & leaf count), independent of their current (age dependent) size
* I interpret this to mean that the three traits (Hmax, LMA, WD) strongly influence each other - presumably, Hmax a function of WD and LMA), but do not particularly 
* this analysis doesn't include accessory proportions, because then I have to use subset of reproducing individuals, greatly reducing sample size. Including these variables, doesn't change the basic story.
* fine tuning:  
  ** remove RGR leaf (similar to RGR)
  ** remove prop_max_weight and prop_max_repro (similar to prop_maxH)

```{r, echo=FALSE,message=FALSE,fig.width=7, fig.height=7}
par(mfrow=c(1,1), cex=1, omi=c(.5,0.3,.1,0.2), mai=c(1.1,1,.6,0.2))
SummaryIndPCA <- SummaryInd %>%
  filter(lvs_end_total>0&shoot_leaf_area>0&shoot_leaf_area_growth!=0) %>%
  select(prop_maxH,height,maxH_spp,WD_mean,LMA_mean,seed_size,growth_leaf,
         growth_stem_diam,leaf_area,growth_shoot_area,growth_stem_area,lvs_end_total,total_weight,
         RA,seedset,repro_all_count,accessory_inv,propagule_inv,shoot_leaf_area,shoot_leaf_area_growth,RGR)

SummaryIndPCA$leaf_area <- log((SummaryIndPCA$leaf_area)+.001)
SummaryIndPCA$total_weight <- log(SummaryIndPCA$total_weight)
SummaryIndPCA$height <- log(SummaryIndPCA$height)
SummaryIndPCA$maxH_spp <- log(SummaryIndPCA$maxH_spp)
SummaryIndPCA$seed_size <- log(SummaryIndPCA$seed_size)
SummaryIndPCA$accessory_inv <- log(SummaryIndPCA$accessory_inv+.001)
SummaryIndPCA$propagule_inv <- log(SummaryIndPCA$propagule_inv+.001)
SummaryIndPCA$growth_shoot_area <- log(SummaryIndPCA$growth_shoot_area)
SummaryIndPCA$growth_leaf <- log(SummaryIndPCA$growth_leaf+10000)
SummaryIndPCA$repro_all_count <- log(SummaryIndPCA$repro_all_count+1)
SummaryIndPCA$lvs_end_total <- log(SummaryIndPCA$lvs_end_total)
SummaryIndPCA$growth_stem_area <- log(SummaryIndPCA$growth_stem_area)
SummaryIndPCA$growth_stem_diam <- log(SummaryIndPCA$growth_stem_diam)
SummaryIndPCA$shoot_leaf_area <- log(SummaryIndPCA$shoot_leaf_area)
SummaryIndPCA$shoot_leaf_area_growth <- log(SummaryIndPCA$shoot_leaf_area_growth)

PCA_mod <- prcomp(SummaryIndPCA, scale=TRUE)
summary(PCA_mod)
biplot(PCA_mod,xlim=c(-.1,.15),ylim=c(-.1,.1),cex=.8)

PCA_loadings <- as.data.frame(PCA_mod$rotation)
PCA_loadings$variables <- row.names(PCA_loadings)

PCA_scaling <- as.data.frame(PCA_mod$scale)
PCA_scaling$variables <- row.names(PCA_scaling)
names(PCA_scaling) <- c("scale","variables")
PCA_loadings <- merge(PCA_loadings,PCA_scaling,by="variables")

PCA_lambda <- as.data.frame(PCA_mod$sdev)

PCA_scores <- as.data.frame(PCA_mod$x)
PCA_scores$individual <- subset(SummaryInd$individual, SummaryInd$lvs_end_total>0&SummaryInd$shoot_leaf_area_growth!=0)
SummaryInd_PCA_results <- merge(SummaryInd,PCA_scores, by="individual", all.x=TRUE)

n <- NROW(PCA_scores)
scale =1

PCA_lambda <- PCA_lambda*sqrt(n)
PCA_lambda <- PCA_lambda^scale

SummaryInd_PCA_results$PC1_scale <- SummaryInd_PCA_results$PC1/PCA_lambda[1]
SummaryInd_PCA_results$PC2_scale <- SummaryInd_PCA_results$PC2/PCA_lambda[2]
SummaryInd_PCA_results$PC3_scale <- SummaryInd_PCA_results$PC3/PCA_lambda[3]
SummaryInd_PCA_results$PC4_scale <- SummaryInd_PCA_results$PC4/PCA_lambda[4]
PCA_loadings$PC1_scale <- PCA_loadings$PC1 / 4.5
PCA_loadings$PC2_scale <- PCA_loadings$PC2 / 4.5
PCA_loadings$PC3_scale <- PCA_loadings$PC3 / 4.5
PCA_loadings$PC4_scale <- PCA_loadings$PC4 / 4.5
```

#Looking at PCA by species
* Main axes are (1) across species affects in trait values (PC1) and (2) within species shifts in RGR vs RA with plant size/age (PC2)
* PC1 is WD vs a whole collection of traits that have higher "values" in plants that are larger as adults, including measures of size, height, and growth increment; PELA is a rare exception in shifting along this axis. 
* PC2 most strongly separates RGR from RA (and a collection of strongly correlated traits, such as propagule investment, prop maxH). 
* PC3 primarily separates LMA and maxH from WD; high wood density is also somewhat associated with higher RGR and various shoot growth and leafiness measures, reflecting that younger individuals have more shoot growth and higher RGR 
* PC4 is effectively leaf count on shoots ("lvs_end_total") and seed set vs. seed size, shoot leaf area and shoot leaf area growth - to first order, some plants have lots of small leaves and others have fewer leaves but greater leaf area, Not sure how seed size fits in

* amazing that RA and RGR almost completely orthoganal to PC1 and PC4; To reiterate, RA and RGR are 2 variables that shift across each species' lifetime and much less across species. I suspect the inter-specific difference are effectively what proportion of individuals in this study are reproductively mature - with early maturing species having overall higher RA and lower RGR (i.e. BOLE) than do late maturing species (i.e. PELA).

* Continuing, RGR and RA are measures that give little insight to a species trait values - or vice versa, trait values do not provide much information about RA...

* "growth_stem_area" and "growth leaf", two whole plant measures of size increase are orthoganal to RGR, while some shoot-level variables (i.e. "growth_shoot_area") and other whole plant variables (e.g. "leaf_area_growth"), are somewhat correlated with both RGR (PC2) and the "maxH" direction (PC1). I suppose this suggests that for plants with high

* For PC2, interesting that "growth_shoot_area"" and other variables associated with shoot growth and species max H correlate somewhat with RGR, but "growth_stem_area" and other whole plant "size" variables do not
* these plots are another way of affirming that RA (and RGR) is a variable that does a good job of comparing trends across species * in other words there is a trade-off between RA and RGR in all species. And even though the RA trends are messy, individuals that are proportionally taller (prop_maxH) or invest more in reproduction (investment, count, or seedset) have a higher RA.

* this last set of variables are all ones that do not vary with the PC1 axis

PCA related code to keep:  
* plot(PCA_mod)
* pairs(SummaryIndPCA,log="xy")
* yy <- t(t(PCA_loadings[,2:3])*PCA_lambda[1:2])
* xx <- t(t(SummaryInd_PCA_results[,105:106])/PCA_lambda[1:2])
* biplot(xx,yy)
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(.5,0.3,.1,0.2), mai=c(1.1,1,.6,0.2))

PCA_by_species <- SummaryInd_PCA_results %>%
  filter(lvs_end_total>0) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se), PC1_scale, PC2_scale, PC3_scale, PC4_scale)

plot(PC1_scale~(PC2_scale),SummaryInd_PCA_results,pch=16,col=col.spp[as.factor(species)],xlim=c(-.1,.1))
arrows(0, 0, (.9*PCA_loadings$PC2_scale), (.9*PCA_loadings$PC1_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC1_scale ~ PC2_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mean1 <- PCA_by_species$PC1_scale_mean
se1 <- PCA_by_species$PC1_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 1")
se_function(mean1,se1)

plot(PC2_scale~(PC1_scale),SummaryInd_PCA_results,pch=16,col=col.spp[as.factor(species)])
arrows(0, 0, (.9*PCA_loadings$PC1_scale), (.9*PCA_loadings$PC2_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC2_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")

mean1 <- PCA_by_species$PC2_scale_mean
se1 <- PCA_by_species$PC2_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 2")
se_function(mean1,se1)

plot(PC3_scale~PC1_scale,SummaryInd_PCA_results,pch=16,col=col.spp[as.factor(species)])
arrows(0, 0, (.9*PCA_loadings$PC1_scale), (.9*PCA_loadings$PC3_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC3_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")

mean1 <- PCA_by_species$PC3_scale_mean
se1 <- PCA_by_species$PC3_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 3")
se_function(mean1,se1)

plot(PC4_scale~PC1_scale,SummaryInd_PCA_results,pch=16,col=col.spp[as.factor(species)],ylim=c(-.15,.15))
arrows(0, 0, (.9*PCA_loadings$PC1_scale), (.9*PCA_loadings$PC4_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC4_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mean1 <- PCA_by_species$PC4_scale_mean
se1 <- PCA_by_species$PC4_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 4",ylim=c(-.07,.04))
se_function(mean1,se1)
```

##How do growth and reproduction shift was Hmax is approached?

###Growth Investment versus proportion max H
- For the canopy species, PELA, HATE, PEPU, BAER, investment in growth plateaus (i.e. growth relative to size decreases) at ~40% maxH
- I think this suggests that relative height is a good predictor of growth investment in these species, but not in some of the lower-growing species

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(.8,0.3,.5,0.2), mai=c(.7,1,.1,0.2))
plot((GrowthInv~prop_maxH), SummaryInd, log="y", col=col.spp[as.factor(species)],pch=16,xlab="")
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(GrowthInv~prop_maxH, SummaryInd, log="y", col=col.age[as.factor(age)],pch=16,ylab="",xlab="")
legend("topleft",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")
mtext("proportion Hmax", 1, outer=TRUE,cex=1.2)
```

- For the shorter and shorter-lived species there is much less of a relationship, or a complete lack of relationship (i.e. BOLE, PHPH)  
- Quite interesting, there is a distinctly negative relationship for GRBU; its absolute growth investment actually declines as it increases in size  
- For many of the species with a good relationship, it seems that there are two breakpoints in the curve - wondering what these mean with respect to reproductive onset and RA 
- **Need to fit proper curves to identify "knickpoints**

###Looking at species individually
```{r, echo=FALSE, message=FALSE, fig.width=24, fig.height=24}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "GrowthInv"
xvar <- "prop_maxH"
data <- SummaryInd
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar, log="y",
                    main=spp, xlab="",ylab="")
}

mtext("proportion Hmax", 1, outer=TRUE,cex=1.2)
mtext("Growth investment", 2, outer=TRUE,cex=1.2)

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm((x[[yvar]]) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

### RGR by proportion Hmax
- Ignoring a few outliers, RGR declines with increasing height is 12 species  
- Only in HEPU and PILI is there no significant trend  
- The species with a poorer correlation are mostly shorter ones that are presumably often broken off   
- These are continued ways to look at the question of how important is height (relative to the species maxH) in understanding its growth and reproduction

```{r, echo=FALSE, message=FALSE, fig.width=7, fig.height=7}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,1,0.6)) 
plot(RGR~prop_maxH,SummaryInd,col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
```

```{r, echo=FALSE, message=FALSE, fig.width=24, fig.height=24}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.7,.5,0.2)) 
yvar <- "RGR"
xvar <- "prop_maxH"
data <- SummaryInd
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,
                    main=spp, xlab="",ylab="")
}

mtext("proportion Hmax", 1, outer=TRUE,cex=1.2)
mtext("RGR", 2, outer=TRUE,cex=1.2)

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm((x[[yvar]]) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

###RA by proportion Hmax
```{r, echo=FALSE, message=FALSE, fig.width=7, fig.height=7}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,1,0.6)) 
plot(RA~prop_maxH,SummaryInd,col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
```

```{r, echo=FALSE, message=FALSE, fig.width=24, fig.height=24}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.7,.5,0.2)) 
yvar <- "RA"
xvar <- "prop_maxH"
data <- SummaryInd
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,
                    main=spp, xlab="",ylab="")
}

mtext("proportion Hmax", 1, outer=TRUE,cex=1.2)
mtext("RA", 2, outer=TRUE,cex=1.2)

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm((x[[yvar]]) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```