### 2015-08-19 results_RAschedules_shorter.Rmd

```{r, echo=FALSE,results='hide', message=FALSE}
library(remake)
create_bindings()

source("scripts/summary_dataframes.R")
source("scripts/functions.R")
```

#RA schedule paper
* I have a dataset with enough resolution across species and ages to show patterns of growth and reproductive investment across ages  
* This includes both RA schedules and RV curves  
* Below are curves plotted relative to "prop_maxH", which is a better fit for most species - but other species should be plotted against total plant weight  
* Plants that grow more also reproduce more, but plants that allocate more to reproduction have lower RGR 
* There is a disinct decline in RGR (or Growth Investment) at the exact "prop_maxH" that plants begin to reproduce 
* Plotting reproductive investment (repro_inv) against RGR yields a much lower r^2 - for every single species. Only RA captures the actual trade-off incurred by plants that allocate more to reproduction, showing one way RA schedules are important  
- As discussed in the review, RA schedule differentiate between greater reproductive output because you are bigger (a positive correlation with plant growth) and greater reproductive output because you are allocating more to reproduction (a negative correlation with plant growth)  
* RV curves continue to be the more relevent metric to determine the actual amount of reproductive output produced by a plant of a given size or height

##RV curves (by proportion max H)
- In a sense, by using proportion max H, these figures are showing the proportion of total growth the plant undergoes before starting reproduction
- The plot across all species of reproductive investment vs. Prop H is much less tight than reproductive investment vs. total weight, but within a species not as obviously different  
- The plot that includes all species shows that *once* a given species has ramped up reproduction, the log-log plot of reproductive investment vs final weight is nearly linear across all species; in other words, to first order, all species allocate the same amount to reproduction relative to their weight.  
- This is presumably the basic allometric scaling relationship discussed by others and wondering if slope I get is same as theirs    
- However, it is important to note that species early in their reproductive trajectory do not follow this pattern

###Plots of size against reproductive investment
* Plots of total reproductive investment against size or height are both highly significant
```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
mod <- lm(log(total_weight)~log(repro_inv), data=subset(SummaryInd,repro_inv>0))
plot(repro_inv~total_weight, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm(log(height)~log(repro_inv), data=subset(SummaryInd,repro_inv>0))
plot(repro_inv~height, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm((prop_maxH)~log(repro_inv), data=subset(SummaryInd,repro_inv>0))
plot((repro_inv)~(prop_maxH), SummaryInd, log="y",col=col.spp[as.factor(species)],pch=16,xlim=c(4e-03,1))
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
```

###Plots of size against RA
* RA not significantly correlated with total height or weight, but of course across and within species effects opposite - bigger individuals of a species have higher RA, but species that are overall bigger have lower RA
* Therefore, there is a negative correlation between prop_maxH (proportion of a species' maximum height) and RA; highly significant, but very low R^2

```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(log(total_weight)~asin(sqrt(RA)), data=subset(SummaryInd,repro_inv>0))
plot(asin(sqrt(RA))~total_weight, SummaryInd, log="x",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logx(mod)

mod <- lm(log(height)~asin(sqrt(RA)), data=subset(SummaryInd,repro_inv>0))
plot(asin(sqrt(RA))~height, SummaryInd, log="x",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logx(mod)

mod <- lm((prop_maxH)~asin(sqrt(RA)), data=subset(SummaryInd,repro_inv>0))
plot((asin(sqrt(RA)))~(prop_maxH), SummaryInd,col=col.spp[as.factor(species)],pch=16,xlim=c(4e-03,1))
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right(mod)
```

##RA schedules (vs. age and height)  
- RA schedules still look "best" plotted against age  
- RA schedules plotted against height are particularly good for a number of species such as EPMI, COER, PELA  
  1. they also show that for PILI that among individuals that reproduce, the "potential" RA decreases with increasing prop_maxH  
  2. when all data points are included, all species except PEPU, PHPH, and PULU show increasing RA with increasing height  
  3. when RA =0 is excluded, the following species still have a significant positive correlation: BOLE, COER, EPMI, GRBU, GRSP, HATE. LEES and PELA have a p-value between 0.05-0.10   
  4. PEPU has the lowest slope - matches Mark's suggestion that it is "hanging on"  
- There are species - such as BOLE - which have a much "cleaner" plot against age instead of height (or any other size measure). This indicates that these species are programmed to reproduce at a certain "rate" regardless of their size

NOTE: Again, a linear line is not necessarily the most appropriate

```{r, echo=FALSE,message=FALSE, fig.width=14, fig.height=7}
yvar <- "RA"
xvar <- "age"

yvar2 <- "RA"
xvar2 <- "height"

data <- SummaryInd
data <- split(data, data$species)

for(spp in names(data)) {
  par(mfrow=c(1,2), cex=1, omi=c(.1,0.3,.5,0.2), mai=c(1.2,1,.5,0.2))  
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar, 
                    main=spp, xlab=xvar, ylab=yvar)
  plot_yvar2_vs_xvar2(data[[spp]], yvar2, xvar2, 
                      main=spp, log="x", xlab=xvar2, ylab="")

}  
```

* RA vs. height by species  
```{r, echo=FALSE, message=FALSE}
yvar <- "RA"
xvar <- "height"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

* RA vs. age by species  
```{r, echo=FALSE, message=FALSE}
yvar <- "RA"
xvar <- "age"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

###RV curves, by age, weight and height
```{r, echo=FALSE,message=FALSE, warning=FALSE, fig.width=21, fig.height=7}
yvar <- "repro_inv"
xvar <- "age"

yvar2 <- "repro_inv"
xvar2 <- "total_weight"

yvar3 <- "repro_inv"
xvar3 <- "height"

data <- SummaryInd
data <- split(data, data$species)

for(spp in names(data)) {
  par(mfrow=c(1,3), cex=1, omi=c(.1,0.3,.5,0.2), mai=c(1.2,1,.5,0.2))  
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar, 
                    main=spp, ,log="xy", xlab=xvar, ylab=yvar)
  plot_yvar2_vs_xvar2(data[[spp]], yvar2, xvar2, 
                      main=spp, log="xy", xlab=xvar2, ylab="")
  plot_yvar3_vs_xvar3(data[[spp]], yvar3, xvar3, 
                      main=spp, log="xy", xlab=xvar3, ylab="")
}  
```

* Reproductive investment vs. height for each species
```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
yvar <- "repro_inv"
xvar <- "height"

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]+1) ~  log(x[[xvar]]+1)))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

* Reproductive investment vs. weight for each species
```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
yvar <- "repro_inv"
xvar <- "total_weight"

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]+1) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results
```

##Growth-reproduction trade-off
###Bigger plants reproduce more
- There is a positive correlation between reproductive investment vs growth investment among individuals that are reproducing: plants that grow more also allocate more to reproduction.  
- It is also true that bigger plants allocate more to reproduction.

```{r, echo=FALSE,message=FALSE}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(log(growth_inv)~log(repro_inv),data=subset(SummaryInd,repro_inv>0))
plot((1+repro_inv)~growth_inv, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)
```

###Plants that allocate more to reproduction (higher RA) have a lower RGR
- When looked at individually, this pattern holds for all species except PEPU and PELA, although there is lots of variation in the exact trade-off pattern displayed by different species  
- You can also look at the plots of RGR vs RA and identify the RA at which RGR is impacted. Some plots show a very linear decline throught, whereas for others there is a distinct "knickpoint" -i.e. BOLE, EPMI, HATE

```{r, echo=FALSE,message=FALSE}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(RGR~asin(sqrt(RA))*species,data=subset(SummaryInd,RA>0))
plot((RA+.01)~RGR, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logxy(mod)
```

###To think about regarding RA schedules and RV curves
- What does it mean that prop_maxH is the better correlate that plant weight?  
1. That "plant weight" is a mis-leading number, since it includes dead "heartwood"  
2. That this is a community where height is the most important variable? Because height = light access or because of some relationship between height and leaf area.  
** Once I have leaf weights, I'll be curious how leaf weight plots against prop_maxH, this might be more of the story

###Looking jointly at effects of RA and prop_maxH on growth inveestment
- Model analyzed is: lm(log([[growth_inv]]) ~  asin(sqrt([[RA]]))*asin(sqrt([[prop_maxH]]))))  
- The canopy species have, not suprisingly, a very obvious maxH, with Growth Investment or RGR tapering markedly at the maxH is approached. In these species growth investment is poorly correlated with RA, but well correlated with height.  
- In some of the mid-height species (GRSP, BOLE, GRBU), growth investment is very poorly correlated with height, but better correlated with RA - in other words, increased RA = less growth.  
- One of the difficult features interpreting this, is that there are species with overall higher correlations and overall lower correlations, and this pattern dominates, not the relative strength  
- My current analyses still lack the ability to tease apart the most obvious pattern - because I don't actually want to show that growth and height are correlated, so much as that the   asymptoting of growth is correlated with height - i.e. the change in slope of the relationship.  
```{r, echo=FALSE, message=FALSE}
yvar <- "growth_inv"
xvar <- "RA"
x2var <- "prop_maxH"

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  asin(sqrt(x[[xvar]]))*asin(sqrt(x[[x2var]]))))
results <- summarise_all_fits_2var(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
names(results) <- c("species","r,squared","p-value","n","intercept","coef_RA","coef_prop_maxH","coef_interaction","SS_RA","SS_prop_maxH","SS_interaction")
results
```

###Plotting RA species means against various traits
* species with higher wood density have higher RA
* taller species have lower RA
* species with higher LMA have lower RA
* no relationship between RA and seed size
* these are mean RA values across all plants, including babies and other non-reproducing individuals
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(.6,0.3,.1,0.2), mai=c(1.1,1,.5,0.2))
plot(RA_mean~WD_mean,data=subset(SummarySpp,RA_mean>0), col=col.spp[as.factor(species)],pch=16,log="y",ylab="")
mod <- lm(WD_mean~asin(sqrt(RA_mean)),SummarySpp)
words.bottom.left.logy(mod)
plot(RA_mean~seed_size, SummarySpp, col=col.spp[as.factor(species)],pch=16,log="xy",ylab="")
mod <- lm(log(seed_size)~asin(sqrt(RA_mean)), SummarySpp)
words.bottom.left.logxy(mod)
legend(.05,.28,col=col.spp, legend=names(labels.spp), pch=16, cex=.8,bty="n")
plot(RA_mean~LMA_mean, SummarySpp, col=col.spp[as.factor(species)],pch=16,log="xy",ylab="")
mod <- lm(log(LMA_mean)~asin(sqrt(RA_mean)), data=subset(SummarySpp,RA_mean>0))
words.top.right.logxy(mod)
plot(RA_mean~maxH, SummarySpp, col=col.spp[as.factor(species)],pch=16,log="xy",ylab="")
mod <- lm(log(maxH)~asin(sqrt(RA_mean)), data=subset(SummarySpp,RA_mean>0))
words.bottom.left.logxy(mod)
mtext("RA",2,outer=TRUE)
```

*just plotting mean RA of mature plants is similar, although BAER throws off many plots - surprised by how high its RA is, and wondering if something is wrong with these #s
```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(.6,0.3,.1,0.2), mai=c(1.1,1,.5,0.2))
plot(maxH~WD_mean,data=subset(SummarySpp), col=col.spp[as.factor(species)],pch=16,log="y",ylab="")
mod <- lm(mature_RA_mean~asin(sqrt(mature_RA_mean)),data=subset(SummarySpp,species!="BAER"))
words.bottom.left.logy(mod)
plot(mature_RA_mean~seed_size, data=subset(SummarySpp), col=col.spp[as.factor(species)],pch=16,log="xy",ylab="")
mod <- lm(log(seed_size)~asin(sqrt(mature_RA_mean)), data=subset(SummarySpp,species!="BAER"))
words.bottom.left.logxy(mod)
legend(.05,.28,col=col.spp, legend=names(labels.spp), pch=16, cex=.8,bty="n")
plot(mature_RA_mean~LMA_mean, data=subset(SummarySpp), col=col.spp[as.factor(species)],pch=16,log="xy",ylab="")
mod <- lm(log(LMA_mean)~asin(sqrt(mature_RA_mean)), data=subset(SummarySpp,species!="BAER"))
words.top.right.logxy(mod)
plot(mature_RA_mean~maxH, data=subset(SummarySpp), col=col.spp[as.factor(species)],pch=16,log="xy",ylab="")
mod <- lm(log(maxH)~asin(sqrt(mature_RA_mean)), data=subset(SummarySpp,species!="BAER"))
words.bottom.left.logxy(mod)
mtext("RA (mature plants only)",2,outer=TRUE)
```

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14}
mod <- lm(log(growth_stem_diam) ~ age*WD_mean*LMA_mean*maxH*RA, SummaryInd)
anova(mod)

spp_zero_RA <- SummaryInd %>%
  filter(RA==0) %>%
  group_by(species) %>%
  summarise_each(funs(length),RA)
spp_zero_RA$RAcex <- (((spp_zero_RA$RA)/20))

sppage_zero_RA <- SummaryInd %>%
  filter(RA==0) %>%
  group_by(species,age) %>%
  summarise_each(funs(length),RA)
sppage_zero_RA$RAcex <- (((sppage_zero_RA$RA)/10))

par(mfrow=c(2,2), cex=1, omi=c(.6,0.3,.1,0.2), mai=c(1,1,.6,0.2))
plot(RA~WD_mean,data=subset(SummaryInd,RA>0), col=col.spp[as.factor(species)],pch=16,log="y",ylim=c(.0008,1))
points((RA+.001)~WD_mean,data=subset(SummaryInd,RA==0),col=col.spp[as.factor(species)],pch=16, cex=spp_zero_RA$RAcex)
mod <- lm(WD_mean~asin(sqrt(RA)),SummaryInd)
words.bottom.left.logy(mod)
plot(RA~seed_size, SummaryInd, col=col.spp[as.factor(species)],pch=16,log="xy",ylim=c(.0008,1))
points((RA+.001)~seed_size,data=subset(SummaryInd,RA==0),col=col.spp[as.factor(species)],pch=16, cex=spp_zero_RA$RAcex)
mod <- lm(log(seed_size)~asin(sqrt(RA)), SummaryInd)
words.bottom.left.logxy(mod)
plot(RA~LMA_mean, SummaryInd, col=col.spp[as.factor(species)],pch=16,log="xy",ylim=c(.0008,1))
points((RA+.001)~LMA_mean,data=subset(SummaryInd,RA==0),col=col.spp[as.factor(species)],pch=16, cex=sppage_zero_RA$RAcex)
mod <- lm(log(LMA_mean)~asin(sqrt(RA)), data=subset(SummaryInd,RA>0))
words.top.right.logxy(mod)
plot(RA~maxH, SummaryInd, col=col.spp[as.factor(species)],pch=16,log="xy",ylim=c(.0008,1))
points((RA+.001)~maxH,data=subset(SummaryInd,RA==0),col=col.spp[as.factor(species)],pch=16, cex=spp_zero_RA$RAcex)
mod <- lm(log(maxH)~asin(sqrt(RA)), data=subset(SummaryInd,RA>0))
words.bottom.left.logxy(mod)
```

As plants approach maximum height, less of their reproductive investment is going to propagules...This only includes individuals with some propagule investment
```{r, echo=FALSE,message=FALSE,fig.width=6, fig.height=6}
mod <- lm(log(prop_maxH)~asin(sqrt(prop_propagule)),data=subset(SummaryInd,prop_maxH>0))
plot(asin(sqrt(prop_propagule))~prop_maxH,data=subset(SummaryInd,prop_maxH>0&prop_propagule>0), log="x", pch=16,col=col.spp[as.factor(species)],ylab="propagules (as proportion repro inv)")
words.bottom.left.logx(mod)
```