### 2015-08-19 RA_figs.rmd

```{r, echo=FALSE,results='hide', message=FALSE,warning=FALSE}
library(remake)
create_bindings()

source("scripts/functions.R")
```

##Summary  
1. RA shifts across species and, on average, increases with increasing age  
2. Plots of RA and investment against proportion maxH show  
    * COER, EPMI, GRBU, HATE, PELA show decent curves; BOLE, GRSP, HEPU, LEES, PHPH have noisier curves  
    * different species show predictable increase in repro vs. veg investment with size, once reproductively mature  
    * all species show predictable increase in veg investment before reproductive maturity
3. Plants that allocate more to reproduction (higher RA) have lower growth  
    * pattern holds against proportional increase in shoot diameter or stem diameter, both across and within species  
    * pattern also true against RGR, but probably less biologically relevent  
4. PCA indicates variables that explain more across vs. within species shifts  
    * against relative growth vs. RA (and absolute reproduction) important (age related effect)  
    * species separated by size, WD, LMA, seedsize  
    * 3rd and 4th axes finetune parts of these patterns, in particular separating canopy species from others  
5. Overall, data set shows increasing RA with age (despite messay RA schedules) and differences across species in community

##1. RA shifts enormously across species  
* There is also an age effect, with increasing RA with age

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=7.5}
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(log(growth_inv)~log(repro_inv),data=subset(SummaryInd,repro_inv>0))
plot((1+repro_inv)~growth_inv, SummaryInd, log="xy", col=col.spp(species),pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
words.top.left.logxy(mod)
legend("bottomright",legend=labels.spp(),col=col.spp,pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5")
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75")
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9")
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95")
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1")
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01")

plot((1+repro_inv)~growth_inv, SummaryInd, log="xy", col=col.age(age),pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
legend("bottomright",legend=labels.age(),col=col.age,pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5")
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75")
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9")
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95")
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1")
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01")
```

##2. RA, RV, and "Veg investment by size" (VV) curves for each species  
* These plots show a lot of scatter...I think it is real  
* Interestingly, different species have a "good" RV vs VV curve  
    * BOLE, COER, EPMI, GRSP, HATE, LEES, sort-of PHPH show consistent increase in reproductive investment with increasing size
    * BAER, GRSP, HATE, HEPU, PEPU, PHPH, PUTU, and sort-of BOLE, COER, EPMI show a consistent increase in vegetative investment with size
* Not surprisingly, COER, EPMI and HATE which have some of the best regressions for both RV and VV-curves have the cleanest RA schedules
* if VV-curves plotted including non-reproducing individuals, get beautiful curves - wonderfully consistent when young, but becoming scattered once reproducing  
    ***it would be nice to plot the VV-curves, with different colors for reproducing and non-reproducing individuals***

###RA curves by species  
* curves here plotted against prop_maxH, not total weight, because on average "cleaner"  
    * take account for inter-site discrepencies in plant weight, that don't seem to affect plant height or reproductive maturity as much

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "RA"
xvar <- "prop_maxH"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  ((x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="y",xlab="",ylab="",cex=1.4,
                    main=spp,col=col.age(data[[spp]]$age)
}

mtext("RA", 2, outer=TRUE,cex=1.2)
mtext("proportion maximum H", 1, outer=TRUE,cex=1.2)
legend("bottomright",legend=labels.age(),col=col.age,pch=16, cex=1.1,bty="n")
```

###RV curves by species
```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "repro_inv"
xvar <- "total_weight"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.4,
                    main=spp,col=col.age(data[[spp]]$age)
}

mtext("Reproductive investment (mg)", 2, outer=TRUE,cex=1.2)
mtext("Total plant weight (mg)", 1, outer=TRUE,cex=1.2)
legend("bottomright",legend=labels.age(),col=col.age,pch=16, cex=1.1,bty="n")
```

###Vegetative invesment by size curves by species, only including reproducing individuals  
* colored by age, but legend not correct, because "age 1" color used for youngest age with reproducing individuals

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "growth_inv"
xvar <- "total_weight"

data <- subset(SummaryInd, RA>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.3,
                    main=spp,col=col.age(data[[spp]]$age)
}

mtext("Vegetative investment (mg)", 2, outer=TRUE,cex=1.2)
mtext("Total plant weight (mg)", 1, outer=TRUE,cex=1.2)
legend("bottomright",legend=labels.age(),col=col.age,pch=16, cex=1.1,bty="n")
```

###Vegetative invesment by size curves by species, including all individuls  
* at younger ages, species follow very predictable growth curves  
* Once reproductively mature, plants no longer following predictable vegetative investment vs. size  
    ***What does this mean in terms of models?*** 
* If these plots were colored by age, they would show that the 5-year old site has somewhat stunted plants (site used for "young plants" by growth project)

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "growth_inv"
xvar <- "total_weight"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.3,
                    main=spp,col=col.mature(data[[spp]]$mature)
}

mtext("Vegetative investment (mg)", 2, outer=TRUE,cex=1.2)
mtext("Total plant weight (mg)", 1, outer=TRUE,cex=1.2)
legend("topleft",legend=labels.mature(),col=col.mature(),pch=16, cex=1.1,bty="n")
```

###Including all investment vs. growth  
* reproducing individuals now plot closer to the "expected line" (as defined by juvenile individuals), but for most species there is still a considerable discrepency
* fit line of veg investment vs. total plant weight through non-reproductive plants
* derive "predicted" total investment (veg+repro) for reproducing individuals
* determine proportional reproductive cost for reproducing individuals; probably look by age

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "total_inv"
xvar <- "total_weight"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.3,
                    main=spp,col=col.mature(data[[spp]]$mature)
}

mtext("Vegetative + reproductive investment (mg)", 2, outer=TRUE,cex=1.2)
mtext("Total plant weight (mg)", 1, outer=TRUE,cex=1.2)
legend("topleft",legend=labels.mature(),col=col.mature(),pch=16, cex=1.1,bty="n")
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "growth_height"
xvar <- "height"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.3,
                    main=spp,col=col.mature(data[[spp]]$mature)
}

###calc_veg <- merge(SummaryInd, select(results, -species), by="species",all.x=TRUE)
###WORK HERE

mtext("Increase in plant height (mm)", 2, outer=TRUE,cex=1.2)
mtext("Plant height (mm)", 1, outer=TRUE,cex=1.2)
legend("topleft",legend=labels.mature(),col=col.mature(),pch=16, cex=1.1,bty="n")
```

####RA is a population level parameter, so noise expected across individuals
* None of these species is known to mast, but of course environmental variation year-to-year
* Within a population, some individuals invest more or less in repro in a given year, but overall population should change with time
* This indicates it is informative to consider site level means - which would show an increase in RA with age in many species

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(1,.9,.4,0.2)) 
yvar <- "RA"
xvar <- "age"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",xlab="",ylab="", cex=1.3,
                    main=spp,xlim=c(1,35))
}

mtext("RA", 1, outer=TRUE,cex=1.2)
mtext("age", 2, outer=TRUE,cex=1.2)
```

###3. Plants that allocate more to reproduction (higher RA) have lower growth
* When plotted again stem diameter increment, most species show a distinct trade-off 
* The pattern is weaker for growth in shoot diameter
* There is also a very distinct correlation across individuals of all species

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(.7,.7,.1,.1), mai=c(1.1,1.1,.4,0.2)) 
plot(asin(sqrt(RA))~growth_stem_diameter,data=subset(SummaryInd,RA>0),log="x",col=col.spp(species), pch=16)
mod <- lm(log(growth_stem_diameter)~asin(sqrt(RA)),data=subset(SummaryInd,RA>0))
words.top.right.logx(mod)
legend("bottomleft",legend=labels.spp(),col=col.spp,pch=16, cex=0.8,bty="n")

plot(asin(sqrt(RA))~growth_shoot_diameter,data=subset(SummaryInd,RA>0),log="x",col=col.spp(species), pch=16)
mod <- lm(log(growth_shoot_diameter)~asin(sqrt(RA)),data=subset(SummaryInd,RA>0))
words.top.right.logx(mod)
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "RA"
xvar <- "growth_stem_diameter"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",cex=1.4, xlab="",ylab="",
                    main=spp,col=col.age(data[[spp]]$age)
}

mtext("Stem diameter growth", 1, outer=TRUE,cex=1.2)
mtext("RA", 2, outer=TRUE,cex=1.2)  
legend("bottomleft",legend=labels.age(),col=col.age,pch=16, cex=1.1,bty="n")
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "RA"
xvar <- "growth_shoot_diameter"

data <- subset(SummaryInd,RA>0)
data <- split(data, data$species)

for(spp in names(data)) {
plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",cex=1.4, xlab="",ylab="",
                        main=spp,col=col.age(data[[spp]]$age)
}

mtext("Shoot diameter growth", 1, outer=TRUE,cex=1.2)
mtext("RA", 2, outer=TRUE,cex=1.2)  
legend("bottomleft",legend=labels.age(),col=col.age,pch=16, cex=1.1,bty="n")
```

###4. PCA analysis also indicates a decline in growth/increasing age with increasing RA - as well as species level differences
* Axis 2 groups a collection of variables indicating increased reproduction; opposite these is relative diameter growth  
    * I have removed RGR from the PCA and added in a scaled stem diameter measure  
    * Although RGR and "Stem diameter increment" have a similar negative correlation with RA, 
    RGR created a much easer to describe PCA - it fell out wonderfully opposite RGR and orthogonal to 
    WD, seedsize, LMA, total plant size

* Orthoganal to these in WD vs. maxH, LMA_mean, seed_size, and shoot_growth - species that are taller as adults have bigger seeds, higher LMA, and greater shoot growth (both stem growth & leaf count), independent of their current (age dependent) size

* I interpret this to mean that the three traits (Hmax, LMA, WD) are strongly correlated and shift predictably across the species; confirms that I have selected a wide range of species from the community

*** this analysis doesn't include accessory proportions, because then I have to use subset of reproducing individuals, greatly reducing sample size. Including these variables, doesn't change the basic story.***

```{r, echo=FALSE,message=FALSE,fig.width=7, fig.height=7}
par(mfrow=c(1,1), cex=1, omi=c(.5,0.3,.1,0.2), mai=c(1.1,1,.6,0.2))
SummaryIndPCA <- SummaryInd %>%
  filter(shoot_leaf_count>0&shoot_leaf_area>0&shoot_growth_leaf_area!=0) %>%
  select(prop_maxH,height,maxH,wood_density,LMA_mean,seed_size,growth_leaf,
         growth_stem_diameter,leaf_area,growth_shoot_diameter,growth_stem_area,shoot_leaf_count,total_weight,
         RA,seedset,repro_all_count,accessory_inv,propagule_inv,shoot_leaf_area,shoot_growth_leaf_area,scaled_growth_stem_diameter)

SummaryIndPCA$leaf_area <- log((SummaryIndPCA$leaf_area)+.001)
SummaryIndPCA$total_weight <- log(SummaryIndPCA$total_weight)
SummaryIndPCA$height <- log(SummaryIndPCA$height)
SummaryIndPCA$maxH <- log(SummaryIndPCA$maxH)
SummaryIndPCA$seed_size <- log(SummaryIndPCA$seed_size)
SummaryIndPCA$accessory_inv <- log(SummaryIndPCA$accessory_inv+.001)
SummaryIndPCA$propagule_inv <- log(SummaryIndPCA$propagule_inv+.001)
SummaryIndPCA$growth_shoot_diameter <- log(SummaryIndPCA$growth_shoot_diameter+.001)
SummaryIndPCA$growth_leaf <- log(SummaryIndPCA$growth_leaf+10000)
SummaryIndPCA$repro_all_count <- log(SummaryIndPCA$repro_all_count+1)
SummaryIndPCA$shoot_leaf_count <- log(SummaryIndPCA$shoot_leaf_count)
SummaryIndPCA$growth_stem_diameter <- log(SummaryIndPCA$growth_stem_diameter)
SummaryIndPCA$shoot_leaf_area <- log(SummaryIndPCA$shoot_leaf_area)
SummaryIndPCA$shoot_growth_leaf_area <- log(SummaryIndPCA$shoot_growth_leaf_area)
SummaryIndPCA$scaled_growth_stem_diameter <- log(SummaryIndPCA$scaled_growth_stem_diameter)
SummaryIndPCA$RA <- asin(sqrt(SummaryIndPCA$RA))

PCA_mod <- prcomp(SummaryIndPCA, scale=TRUE)
summary(PCA_mod)
biplot(PCA_mod,xlim=c(-.1,.15),ylim=c(-.15,.1),cex=.8)

PCA_loadings <- as.data.frame(PCA_mod$rotation)
PCA_loadings$variables <- row.names(PCA_loadings)

PCA_scaling <- as.data.frame(PCA_mod$scale)
PCA_scaling$variables <- row.names(PCA_scaling)
names(PCA_scaling) <- c("scale","variables")
PCA_loadings <- merge(PCA_loadings,PCA_scaling,by="variables")

PCA_lambda <- as.data.frame(PCA_mod$sdev)

PCA_scores <- as.data.frame(PCA_mod$x)
PCA_scores$individual <- subset(SummaryInd$individual, SummaryInd$shoot_leaf_count>0&SummaryInd$shoot_growth_leaf_area!=0)
SummaryInd_PCA_results <- merge(SummaryInd,PCA_scores, by="individual", all.x=TRUE)

n <- NROW(PCA_scores)
scale =1

PCA_lambda <- PCA_lambda*sqrt(n)
PCA_lambda <- PCA_lambda^scale

SummaryInd_PCA_results$PC1_scale <- SummaryInd_PCA_results$PC1/PCA_lambda[1]
SummaryInd_PCA_results$PC2_scale <- SummaryInd_PCA_results$PC2/PCA_lambda[2]
SummaryInd_PCA_results$PC3_scale <- SummaryInd_PCA_results$PC3/PCA_lambda[3]
SummaryInd_PCA_results$PC4_scale <- SummaryInd_PCA_results$PC4/PCA_lambda[4]
PCA_loadings$PC1_scale <- PCA_loadings$PC1 / 4.5
PCA_loadings$PC2_scale <- PCA_loadings$PC2 / 4.5
PCA_loadings$PC3_scale <- PCA_loadings$PC3 / 4.5
PCA_loadings$PC4_scale <- PCA_loadings$PC4 / 4.5

SummaryInd_PCA_results <-SummaryInd_PCA_results %>%
  filter(complete.cases(SummaryInd_PCA_results$PC1_scale))

SummaryInd_PCA_results <-SummaryInd_PCA_results %>%
  filter(complete.cases(SummaryInd_PCA_results$PC2_scale))

SummaryInd_PCA_results <-SummaryInd_PCA_results %>%
  filter(complete.cases(SummaryInd_PCA_results$PC3_scale))

SummaryInd_PCA_results <-SummaryInd_PCA_results %>%
  filter(complete.cases(SummaryInd_PCA_results$PC4_scale))

PCA_by_species <- SummaryInd_PCA_results %>%
  group_by(species) %>%
summarise_each(funs(mean, se), PC1_scale, PC2_scale, PC3_scale, PC4_scale)

PCA_by_age <- SummaryInd_PCA_results %>%
group_by(age) %>%
summarise_each(funs(mean, se), PC1_scale, PC2_scale, PC3_scale, PC4_scale)
```

####PCA axis 1 is across species  
* it explains 34% of the variance  
* it separates species with high wood density (wood_density), seedset, RA versus species with large total size, high leaf area, high LMA, large seedsize etc
* the age indepentent measures like prop_maxH and scaled_growth_stem_area have near-zero loadings
* BAER, HATE have by far the highest loadings and BOLE, COER, EPMI, HEPU, LEES, PILI the lowest  
* not sure why high wood density species also have high seedset - is this just by chance with my species?

```{r, echo=FALSE,message=FALSE,fig.width=20, fig.height=14}
par(mfrow=c(2,3), cex=1, omi=c(.5,0.3,.1,0.2), mai=c(1.1,1,.6,0.2))

plot(PC1_scale~(PC2_scale),SummaryInd_PCA_results,pch=16,col=col.spp(species),xlim=c(-.1,.1))
arrows(0, 0, (.8*PCA_loadings$PC2_scale), (.8*PCA_loadings$PC1_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC1_scale ~ PC2_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")
legend("topright",legend=labels.spp(),col=col.spp,pch=16, cex=.8,bty="n")

mean1 <- PCA_by_species$PC1_scale_mean
se1 <- PCA_by_species$PC1_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 1")
se_function(mean1,se1)

PCA_loadings <- transform(PCA_loadings,variables=reorder(variables,-PC1_scale))
plot(PC1_scale~as.factor(variables),PCA_loadings,cex.axis=.8,las=3,xlab="")

plot(PC1_scale~(PC2_scale),SummaryInd_PCA_results,pch=16,col=col.age(age)
arrows(0, 0, (.9*PCA_loadings$PC2_scale), (.9*PCA_loadings$PC1_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC1_scale ~ PC2_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")
legend("topright",legend=labels.age(),col=col.age,pch=16, cex=.8,bty="n")

mean1 <- PCA_by_age$PC1_scale_mean
se1 <- PCA_by_age$PC1_scale_se
plot(mean1~as.factor(age),PCA_by_age,cex.axis=.5,pch=16,ylab="PCA axis 1")
se_function(mean1,se1)
```

####PCA axis 2 is about age  
* it explains 23% of the variance 
* top variable loadings: shoot_growth_leaf_area and growth_shoot_diameter vs repro_all_count, RA, prop_maxH - a collection of variables that shift fro each to late in life   
* COER, EPMI,PELA, and PEPU have the higher loadings and BOLE, LEES have the lowest - to first order these differences are about the average age of reproducing individuals of each species
```{r, echo=FALSE,message=FALSE,fig.width=20, fig.height=14}
par(mfrow=c(2,3), cex=1, omi=c(.5,0.3,.1,0.2), mai=c(1.1,1,.6,0.2))

plot(PC2_scale~(PC1_scale),SummaryInd_PCA_results,pch=16,col=col.spp(species))
arrows(0, 0, (.9*PCA_loadings$PC1_scale), (.9*PCA_loadings$PC2_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC2_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")

mean1 <- PCA_by_species$PC2_scale_mean
se1 <- PCA_by_species$PC2_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 2")
se_function(mean1,se1)

PCA_loadings <- transform(PCA_loadings,variables=reorder(variables,-PC2_scale))
plot(PC2_scale~as.factor(variables),PCA_loadings,cex.axis=.8,las=3,xlab="")

plot(PC2_scale~(PC1_scale),SummaryInd_PCA_results,pch=16,col=col.age(age)
arrows(0, 0, (.9*PCA_loadings$PC1_scale), (.9*PCA_loadings$PC2_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC2_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")

mean1 <- PCA_by_age$PC2_scale_mean
se1 <- PCA_by_age$PC2_scale_se
plot(mean1~as.factor(age),PCA_by_age,cex.axis=.5,pch=16,ylab="PCA axis 2")
se_function(mean1,se1)
```

####PCA axis 3 again divides mainly by species
* it explains 9% of the variance (total variance explained now at 68%)  
* It is dividing the three (of 4) canopy species, HATE, PEPU, PELA, (BAER) versus
* top variable loadings: scaled_growth_stem_diameter, wood_density,growth_stem_diameter, shoot_leaf_count vs. maxH, LMA_mean, seed_size, RA  
* not terribly interpretable, since doesn't make sense that scaled_growth_stem_diameter with wood_density

```{r, echo=FALSE,message=FALSE,fig.width=20, fig.height=14}
par(mfrow=c(2,3), cex=1, omi=c(.5,0.3,.1,0.2), mai=c(1.1,1,.6,0.2))

plot(PC3_scale~PC1_scale,SummaryInd_PCA_results,pch=16,col=col.spp(species),ylim=c(-.15,.15))
arrows(0, 0, (.8*PCA_loadings$PC1_scale), (.8*PCA_loadings$PC3_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(1.1*PC3_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=.8,col="black")

mean1 <- PCA_by_species$PC3_scale_mean
se1 <- PCA_by_species$PC3_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 3")
se_function(mean1,se1)

PCA_loadings <- transform(PCA_loadings,variables=reorder(variables,-PC3_scale))
plot(PC3_scale~as.factor(variables),PCA_loadings,cex.axis=.8,las=3,xlab="")

plot(PC3_scale~(PC1_scale),SummaryInd_PCA_results,pch=16,col=col.age(age),ylim=c(-.15,.15))
arrows(0, 0, (.8*PCA_loadings$PC1_scale), (.8*PCA_loadings$PC3_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(1.1*PC3_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")

mean1 <- PCA_by_age$PC3_scale_mean
se1 <- PCA_by_age$PC3_scale_se
plot(mean1~as.factor(age),PCA_by_age,cex.axis=.5,pch=16,ylab="PCA axis 2")
se_function(mean1,se1)
```

####PCA axis 4  
* it explains 7% of the variance (total variance explained now at 74%)
* it divides stem vs shoot growth variables  
* high loadings for growth_stem_area, growth_leaf, LMA_mean, seed_set vs low loadings for shoot_growth_leaf_area, wood_density, prop_maxH, seed_size  
* BAER, EPMI, HATE and HEPU have higher values vs. BOLE, GRBU, GRSP with lower values; in a sense these later species have more shoot growth than stem growth and spend more of their life close to their maxH - which is true  
* there is no significant age component  
```{r, echo=FALSE,message=FALSE,fig.width=20, fig.height=14}
par(mfrow=c(2,3), cex=1, omi=c(.5,0.3,.1,0.2), mai=c(1.1,1,.6,0.2))

plot(PC4_scale~PC1_scale,SummaryInd_PCA_results,pch=16,col=col.spp(species),ylim=c(-.15,.15))
arrows(0, 0, (.9*PCA_loadings$PC1_scale), (.9*PCA_loadings$PC4_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(PC4_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")
legend("topright",legend=labels.spp(),col=col.spp,pch=16, cex=.8,bty="n")

mean1 <- PCA_by_species$PC4_scale_mean
se1 <- PCA_by_species$PC4_scale_se
plot(mean1~species,PCA_by_species,cex.axis=.5,pch=16,ylab="PCA axis 4",ylim=c(-.05,.06))
se_function(mean1,se1)

PCA_loadings <- transform(PCA_loadings,variables=reorder(variables,-PC4_scale))
plot(PC4_scale~as.factor(variables),PCA_loadings,cex.axis=.8,las=3,xlab="")

plot(PC4_scale~(PC1_scale),SummaryInd_PCA_results,pch=16,col=col.age(age),ylim=c(-.15,.15))
arrows(0, 0, (.8*PCA_loadings$PC1_scale), (.8*PCA_loadings$PC4_scale), length = 0.25, angle = 30, code = 2, col = "red",lwd=2)
text(1.1*PC4_scale ~ PC1_scale, PCA_loadings, labels=variables, pos=1, offset=0, cex=1,col="black")

mean1 <- PCA_by_age$PC4_scale_mean
se1 <- PCA_by_age$PC4_scale_se
plot(mean1~as.factor(age),PCA_by_age,cex.axis=.5,pch=16,ylab="PCA axis 2")
se_function(mean1,se1)
```