### 2015-08-26 Accessory_figs.rmd
starting to get scripts to run again - accessory_figs runs with a few small hitches (i.e. not using SE)

```{r}
remake::make("export")

SummaryInd <- readRDS("export/SummaryInd.rds")
SummarySpp <- readRDS("export/SummarySpp.rds")
SummarySppAge <- readRDS("export/SummarySppAge.rds")

source("R/figures.R")
```

#Summary  
1. Total reproductive investment (vs. propagule investment or fruit investment) shows the strongest correlation with total plant size
2. Accessory costs represent the majority of reproductive investment  
3. Accessory costs are higher in large species, species with low seedset, and species with large seeds
4. Accessory costs within a species do not shift with increasing RA, total reproductive investment, or prop_maxH, 
indicating no compounding returns with greater relative or absolute reproductive investment or with aging (not sure what manuscript this goes in)
5. There are ACROSS species shifts in investment going to some of the 5 tissue pools with increasing RA:  
      ***waiting for correct calculations to draw conclusions for 5***
6. Proportion of reproductive investment going to each of the 5 tissue pools (pre-pollen aborted, pre-pollen successful, post-pollen aborted, packaging&dispersal, propagule) 
is idiosyncratic across species

##1. Total reproductive investment is the best correlate with total plant size 
* Measuring total reproductive output has a much stronger correlation than investment in fruit weight or propagules
* The regression of total weight against prepollination investment is as good as total reproductive investment - one could just count buds + flowers!
    * This is likely because prepollination investment is much a large proportion of total reproductive investment
* Note that each of these figures (and models) only includes individuals with non-zero investment in the reproductive pool plotted
* Important for research that wishes to predict reproductive investment from plant size  
* Of course propagule investment is the correlate with seed count

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=14}
options(scipen=10)
par(mfrow=c(2,2), cex=1, omi=c(.6,0.3,.3,0.3), mai=c(.8,1,.3,0.3))
mod <- lm (log(total_weight)~log(repro_inv), data=subset(SummaryInd,repro_inv>0))
plot(repro_inv~total_weight,data=subset(SummaryInd,repro_inv>0), log="xy",xlab="", col=col.spp(species), pch=16, ylab="Total reproductive investment (mg)",xlim=c(100,2000000))
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(fruit_weight), data=subset(SummaryInd,fruit_weight>0))
plot(fruit_weight~total_weight,data=subset(SummaryInd,fruit_weight>0), log="xy",col=col.spp(species), pch=16, ylab="Fruit investment (mg)",xlim=c(100,2000000),xlab="")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(propagule_inv), data=subset(SummaryInd,propagule_inv>0))
plot(propagule_inv~total_weight,data=subset(SummaryInd,propagule_inv>0), log="xy",col=col.spp(species), pch=16, xlab="", ylab="Propagule investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(prepollen_all_inv), data=subset(SummaryInd,prepollen_all_inv>0))
plot(prepollen_all_inv~total_weight,data=subset(SummaryInd,prepollen_all_inv>0), log="xy",col=col.spp(species), pch=16, xlab="", ylab="Prepollen investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mtext("Total plant size (mg)",1, outer=TRUE,cex=1.2)
```

##2. Accessory costs are the majority of reproductive investment for most species  
* Also, most accessory costs are due to pre-pollination investment

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=7.5}
par(mfrow=c(1,2), cex=1, omi=c(.6,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(accessory_inv~repro_inv,data=subset(SummaryInd,accessory_inv>0),log="xy",pch=16,col=col.spp(species),
     xlim=c(1,700000),ylim=c(1,700000),xlab="Total reproductive investment (mg)",ylab="Investment in accessory tissues (mg)")
abline(0,1)
abline(-0.30103,1)
abline(-1,1)
text(120,350,"all RE to accessory tissues",srt=45)
text(12000,720,"10% RE to accessory tissues",srt=45)
text(5,1.5,"50%",srt=45)

plot(prepollen_all_inv~accessory_inv,data=subset(SummaryInd,accessory_inv>0),log="xy",pch=16,col=col.spp(species),
     xlim=c(1,700000),ylim=c(1,700000),ylab="Investment in pre-pollination tissues (mg)",xlab="Investment in all accessory tissues (mg)")
abline(0,1)
abline(-0.30103,1)
abline(-1,1)
text(120,250,"all acc costs pre-pollination",srt=45)
text(3000,180,"10% acc costs pre-pollination",srt=45)
text(10,3,"50%",srt=45)
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
```

##3. What do accessory costs correlate with?
Results:  
    - Larger-seeded species have higher accessory costs, as has been shown in other studies  
    - Also: "larger species" and species with lower seed set have higher accessory costs  
    - And: species with greater prepollination investment have higher accessory costs (since this is the majority of accessory costs for most species)  
    - When all 4 variables put into a single model, depending on how variables expressed, transformed, different factors (seedset, seed size, maxH-species, pre-pollination costs) come out as have the best explanatory power

Interpretation:  
    - Not sure one can separate the effect of plant size, seed set, seed size on what drive accessory costs  
    - the reason seed set is strongly correlated, is because if flowers are pollinated and develop into mature seeds, the individual plant ends up investing more energy in reproduction and that extra energy is being invested into propagules, 
    reducing accessory costs.

* This is true at the species and individual level.

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(0.4,0.6,0.7,0.3), mai=c(.7,1,0.3,0.3))

mod <- lm(log(seed_size)~(accessory_per_seed),SummarySpp[["mean"]])
plot(accessory_per_seed~seed_size,SummaryInd, log="xy", pch=16,col=col.spp(species),ylab="accessory costs per successful seed",xlab="",cex=.9)
points((accessory_per_seed)~seed_size,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
words.bottom.right.logxy(mod)
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")

mod <- lm(log(seed_size)~prop_accessory,SummarySpp[["mean"]])
plot(prop_accessory~seed_size,SummaryInd, log="x", pch=16,col=col.spp(species),ylab="proportional accessory costs",xlab="",cex=.9)
words.bottom.right.logx(mod)
points(prop_accessory~seed_size,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
mtext("seed size (mg)",1, outer=TRUE,cex=1.2)

mod <- lm(log(total_weight)~log(accessory_per_seed),SummarySpp[["mean"]])
plot(accessory_per_seed~total_weight,SummaryInd, log="xy", pch=16,col=col.spp(species),ylab="accessory costs per successful seed",xlab="",cex=.9)
points((accessory_per_seed)~total_weight,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
words.bottom.right.logxy(mod)

mod <- lm(log(total_weight)~prop_accessory,SummarySpp[["mean"]])
plot(prop_accessory~total_weight,SummaryInd, log="x", pch=16,col=col.spp(species),ylab="proportional accessory costs",xlab="",cex=0.9)
points(prop_accessory~total_weight,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
words.bottom.right.logx(mod)
mtext("total plant weight (mg)",1, outer=TRUE,cex=1.2)

mod <- lm(log(seedset)~log(accessory_per_seed),SummarySpp[["mean"]])
plot(accessory_per_seed~seedset,SummaryInd, log="xy", pch=16,col=col.spp(species),ylab="accessory costs per successful seed",xlab="",cex=0.9)
points(accessory_per_seed~seedset,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
words.top.right.logxy(mod)

mod <- lm(log(seedset)~prop_accessory,SummarySpp[["mean"]])
plot(prop_accessory~seedset,SummaryInd, log="x", pch=16,col=col.spp(species),ylab="proportional accessory costs",xlab="",cex=0.9)
legend("bottomleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
points(prop_accessory~seedset,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
words.top.right.logx(mod)
mtext("seedset",1, outer=TRUE,cex=1.2)

par(mfrow=c(1,2), cex=1, omi=c(0.4,0.6,0.7,0.3), mai=c(1.1,1,0.3,0.3))
mod <- lm(log(prepollen_all_per_seed)~log(accessory_per_seed),SummarySpp[["mean"]])
plot(accessory_per_seed~(prepollen_all_per_seed),SummaryInd, log="xy",pch=16,col=col.spp(species),ylab="accessory costs per successful seed",cex=0.9)
points(accessory_per_seed~prepollen_all_per_seed,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
words.bottom.right.logxy(mod)

mod <- lm(prop_prepollen_all~prop_accessory,SummarySpp[["mean"]])
plot(prop_accessory~prop_prepollen_all,SummaryInd, log="", pch=16,col=col.spp(species),ylab="proportional accessory costs",cex=0.9)
points(prop_accessory~prop_prepollen_all,SummarySpp[["mean"]], pch=16,col=col.spp(species),cex=1.8)
words.bottom.right(mod)
```

SummarySpp[["mean"]]$accessory_per_seed

###To discuss: Which measure of "accessory costs" should one use? 
* "accessory costs per successful seed" vs "proportion of reproductive investment to accessory tissues"  
* "per seed" is the formulation characteristically used and best relates the results to fecundity
* the "per seed" variant greatly reduces the accessory costs of species that produce lots of seeds - and since species vary considerably in seed production it stratifies species more  
* the "proportional" calculation essentially assumes total propagule investment (by weight) is equivalent, whether it is many small seeds or fewer big ones

##4. Does increasing RA, reproductive investment or relative height yield compounding returns in the form of higher seedset or increased propagule allocation?

###for RA:
* NO - seedset and relative propagule investment do not increase with increasing RA 
* Furthermore, looking within species  
    - Not a single species shows an increase in prop_propagule with increasing RA  
    - And for seedset vs RA, only HATE shows an increase, and HEPU a decrease  
    - So no, there are no compounding benefits to increasing RA

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14,warning=FALSE}
par(mfrow=c(2,2), cex=1, omi=c(0.2,0.3,0.3,0.3), mai=c(1.1,1,0.3,0.3)) 
mod <- lm(seedset~RA,data=subset(SummaryInd,seedset>0&seedset<1))
plot(RA~seedset,data=subset(SummaryInd,seedset>0&seedset<1), pch=16,col=col.spp(species))
words.bottom.right(mod)

mod <- lm(prop_propagule~RA,data=subset(SummaryInd,prop_propagule>0))
plot(RA~prop_propagule,data=subset(SummaryInd,seedset>0&seedset<1), pch=16,col=col.spp(species))
words.bottom.right(mod)
legend("topright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")

mod <- lm(propagule_per_seed~RA,data=subset(SummaryInd,seedset>0&seedset<1))
plot(RA~propagule_per_seed,data=subset(SummaryInd,seedset>0&seedset<1), log="x",pch=16,col=col.spp(species))
words.bottom.right.logx(mod)
```

###for total reproductive investment:  
* Not within species - as total reproductive investment increases there are no within species shifts with respect to seed set, proportional propagule allocation, or propagule allocation per seed
* Plotting all data shows the between species effect - that species with larger adult plants have lower seedset and proportional propagule investment  
* the third plot shows scatter for species where some (or all) individuals have actual seeds weighed, but not for species that really just have a mean seed weight plugged in for all individuals

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14,warning=FALSE}
par(mfrow=c(2,2), cex=1, omi=c(0.2,0.3,0.3,0.3), mai=c(1.1,1,0.3,0.3)) 
options(scipen=20)
mod <- lm(seedset~log(repro_inv),data=subset(SummaryInd,seedset>0&seedset<1))
plot(repro_inv~seedset,data=subset(SummaryInd,seedset>0&seedset<1), pch=16,col=col.spp(species),log="y")
words.bottom.right.logy(mod)

mod <- lm(prop_propagule~log(repro_inv),data=subset(SummaryInd,prop_propagule>0))
plot(repro_inv~prop_propagule,data=subset(SummaryInd,seedset>0&seedset<1), pch=16,col=col.spp(species),log="y")
words.bottom.right.logy(mod)
legend("topright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")

mod <- lm(log(propagule_per_seed)~log(repro_inv),data=subset(SummaryInd,seedset>0&seedset<1))
plot(repro_inv~propagule_per_seed,data=subset(SummaryInd,seedset>0&seedset<1), log="xy",pch=16,col=col.spp(species))
words.bottom.right.logxy(mod)
```

###for proportional height:
* Not within species - looked at within species regressions for prop_maxH against seedset, prop_propagule, propagule_per_seed and no consistent patterns  
* However there is a strong across species effect, such that species where more of the reproducing individuals have lower "prop_maxH" have higher seed set vs. the opposites  
* This indicates that species that begin reproducing at smaller sizes have higher seedset  
    * We already knew that individuals from species with shorter Hmax and individuals that are overall shorter have higher seedset  
    * In part this is suggetsing that the species with high seedset have, on average, mostly individuals that are much 
      shorter than the maxH achieved and that reproduction is curtailing their growth even more than maxH indicates

```{r, echo=FALSE,message=FALSE,fig.width=14, fig.height=14,warning=FALSE}
par(mfrow=c(2,2), cex=1, omi=c(0.2,0.3,0.3,0.3), mai=c(1.1,1,0.3,0.3)) 
options(scipen=20)
mod <- lm(prop_maxH~seedset,data=subset(SummaryInd,seedset>0&seedset<1))
plot(seedset~prop_maxH,data=subset(SummaryInd,seedset>0&seedset<1), pch=16,col=col.spp(species))
words.bottom.right(mod)

mod <- lm(log(height)~seedset,data=subset(SummaryInd,seedset>0&seedset<1))
plot(seedset~height,data=subset(SummaryInd,seedset>0&seedset<1), pch=16,col=col.spp(species),log="x")
words.bottom.right.logx(mod)

mod <- lm(prop_maxH~prop_propagule,data=subset(SummaryInd,prop_propagule>0))
plot(prop_propagule~prop_maxH,data=subset(SummaryInd,seedset>0&seedset<1), pch=16,col=col.spp(species))
words.bottom.right(mod)
legend("topright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")

mod <- lm(prop_maxH~log(propagule_per_seed),data=subset(SummaryInd,seedset>0&seedset<1))
plot(propagule_per_seed~prop_maxH,data=subset(SummaryInd,seedset>0&seedset<1), log="y",pch=16,col=col.spp(species))
words.bottom.right.logy(mod)
```

###5. What influences the correlation between RA and plant size (or total reproductive investment)?
  ***Don't trust anything until I know RA and repro_inv are correct, because PHPH and PUTU have huge influence on patterns***
```{r, echo=FALSE,message=FALSE,fig.width=12, fig.height=30,warning=FALSE}
par(mfrow=c(5,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2))

mod <- lm(RA~log(propagule_inv),data=subset(SummaryInd,RA>0&propagule_inv>0))
plot(propagule_inv~RA,data=subset(SummaryInd,RA>0&propagule_inv>0), pch=16,col=col.spp(species),log="y", ylab="energy to propagules (mg)")
words.top.left.logxy(mod)

mod <- lm(RA~prop_propagule,data=subset(SummaryInd,RA>0&prop_propagule>0))
plot(prop_propagule~RA,data=subset(SummaryInd,RA>0&prop_propagule>0), log="", pch=16,col=col.spp(species),ylab="propagules (as proportion repro inv)")
words.bottom.left.logx(mod)

mod <- lm(RA~log(prepollen_aborted_inv),data=subset(SummaryInd,RA>0&prepollen_aborted_inv>0))
plot(prepollen_aborted_inv~RA,data=subset(SummaryInd,RA<1&RA>0), pch=16,col=col.spp(species),log="y",ylab="pre-pollination aborted tissues (mg)")
words.top.left.logxy(mod)

mod <- lm(RA~prop_prepollen_aborted,data=subset(SummaryInd,RA>0&prop_prepollen_aborted>0))
plot(prop_prepollen_aborted~RA,data=subset(SummaryInd,RA>0&prop_prepollen_aborted>0), pch=16,col=col.spp(species), log="",ylab="pre-pollination aborted tissues (as proportion repro inv)")
words.bottom.right.logx(mod)

mod <- lm(RA~log(prepollen_success_inv),data=subset(SummaryInd,RA>0&prepollen_success_inv>0))
plot(prepollen_success_inv~RA,data=subset(SummaryInd,RA<1&RA>0), pch=16,col=col.spp(species),log="y",ylab="pre-pollination successful tissues (mg)")
words.top.left.logxy(mod)

mod <- lm(RA~prop_prepollen_success,data=subset(SummaryInd,RA>0&prop_prepollen_success>0))
plot(prop_prepollen_success~RA,data=subset(SummaryInd,RA>0&prop_prepollen_success>0), pch=16,col=col.spp(species), log="",ylab="pre-pollination successful tissues (as proportion repro inv)")
words.bottom.right.logx(mod)

mod <- lm(RA~log(packaging_dispersal_inv),data=subset(SummaryInd,RA>0&packaging_dispersal_inv>0))
plot(packaging_dispersal_inv~RA,data=subset(SummaryInd,RA>0&packaging_dispersal_inv>0), pch=16,col=col.spp(species),log="y", ylab="energy to packaging and dispersal (mg)")
words.bottom.right.logxy(mod)
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")

mod <- lm(RA~prop_packaging_dispersal,data=subset(SummaryInd,RA>0&prop_packaging_dispersal>0))
plot(prop_packaging_dispersal~RA,data=subset(SummaryInd,RA>0&prop_packaging_dispersal>0), log="", pch=16,col=col.spp(species),ylab="packaging and dispersal (as proportion repro inv)")
words.bottom.left.logx(mod)

mod <- lm(RA~log(postpollen_aborted_inv),data=subset(SummaryInd,RA>0&postpollen_aborted_inv>0))
plot(postpollen_aborted_inv~RA,data=subset(SummaryInd,RA>0&postpollen_aborted_inv>0), pch=16,col=col.spp(species),log="y", ylab="energy to tissues aborted postpollination (mg)")
words.bottom.right.logxy(mod)

mod <- lm(RA~prop_postpollen_aborted,data=subset(SummaryInd,RA>0&prop_postpollen_aborted>0))
plot(prop_postpollen_aborted~RA,data=subset(SummaryInd,RA>0&prop_postpollen_aborted>0), log="", pch=16,col=col.spp(species),ylab="tissues aborted postpollination (as proportion repro inv)")
words.bottom.left.logx(mod)
```

##6. Shifts in proportion to tissue types across species  
- Reasonably idiosyncratic patterns  
- Species very different in which accessory tissues they invest the most in - which reflects the differences in basic structure (i.e. Banksia cones), but also more subtlely that some species abort lots of buds (pre-pollination aborted), while other species abort more fully formed flowers   
- COER, EPMI, LEES, PILI invest the most energy in propagules  
- HATE invests by far the least into propagules  
```{r, echo=FALSE, message=FALSE, fig.width=14, fig.height=28,warning=FALSE}
par(mfrow=c(4,2), cex=1, omi=c(1,.1,.7,.1), mai=c(.5,1.1,.2,0.2))  
mean1 <- SummarySpp[["mean"]]$prop_propagule
se1 <- SummarySpp[["sd"]]$prop_propagule
boxplot(mean1~species,SummarySpp[["mean"]],cex.axis=.5,pch=16,log="y",ylab="proportion propagules",xlab="")
se_function(mean1,se1)

mean1 <- SummarySpp[["mean"]]$propagule_per_seed
se1 <- SummarySpp[["sd"]]$propagule_per_seed
boxplot(mean1~species,SummarySpp[["mean"]],cex.axis=.5,pch=16,log="y",ylab="propagules",xlab="")
se_function(mean1,se1)

mean1 <- SummarySpp[["mean"]]$prop_prepollen_all
se1 <- SummarySpp[["sd"]]$prop_prepollen_all
boxplot(mean1~species,data=subset(SummarySpp[["mean"]]),cex.axis=.5,pch=16,ylim=c(.1,1),ylab="pre-pollination tissues",xlab="")
se_function(mean1,se1)

mean1 <- SummarySpp[["mean"]]$prepollen_all_per_seed
se1 <- SummarySpp[["sd"]]$prepollen_all_per_seed
boxplot((.01+prepollen_all_per_seed)~species,data=subset(SummarySpp[["mean"]]),cex.axis=.5,pch=16,ylab="pre-pollination tissues",xlab="",log="y")
se_function(mean1,se1)

mean1 <- SummarySpp[["mean"]]$prop_packaging_dispersal
se1 <- SummarySpp[["sd"]]$prop_packaging_dispersal
boxplot(mean1~species,SummarySpp[["mean"]],cex.axis=.5,pch=16,log="y",ylim=c(.02,1),ylab="packaging and dispersal")
se_function(mean1,se1)

mean1 <- SummarySpp[["mean"]]$packaging_dispersal_per_seed
se1 <- SummarySpp[["sd"]]$packaging_dispersal_per_seed
boxplot((packaging_dispersal_per_seed+.1)~species,SummarySpp[["mean"]],cex.axis=.5,pch=16,ylab="packaging and dispersal",log="y")
se_function(mean1,se1)

mean1 <- SummarySpp[["mean"]]$prop_postpollen_aborted
se1 <- SummarySpp[["sd"]]$prop_postpollen_aborted
boxplot(mean1~species,SummarySpp[["mean"]],cex.axis=.5,pch=16,log="y",ylim=c(.02,1),ylab="aborted post-pollination tissues")
se_function(mean1,se1)

mean1 <- SummarySpp[["mean"]]$postpollen_aborted_per_seed
se1 <- SummarySpp[["sd"]]$postpollen_aborted_per_seed
boxplot((postpollen_aborted_per_seed+.1)~species,SummarySpp[["mean"]],cex.axis=.5,pch=16,ylab="aborted post-pollination tissues",log="y")
se_function(mean1,se1)

mtext("proportion of reproductive investment                                                                       mg investment per seed",3,outer=TRUE,cex=1.1)
mtext("species",1, outer=TRUE,cex=1.1)
```

- just looking among plants >3 years old  
```{r, echo=FALSE, message=FALSE, fig.width=14, fig.height=7}
par(mfrow=c(1,2), cex=1, omi=c(1,.1,.7,.1), mai=c(.5,1.1,.2,0.2)) 
tmp <- SummaryInd %>%
  filter(repro_inv!=0&age>3) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se), prop_prepollen_all)

plot(mean~as.factor(species),
     tmp,
     cex.axis=.5,pch=16,log="y",ylim=c(.02,1),
     ylab="Prop reproductive investment put into pre-pollination tissues")

mean1 <- tmp$mean
se1 <- tmp$se
se_function(mean1,se1)

tmp2 <- SummaryInd %>%
  filter(repro_inv!=0&age>3) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se), prop_packaging_dispersal)

plot(mean~as.factor(species),
     tmp2,
     cex.axis=.5,pch=16,log="y",ylim=c(0.08,.8),
     ylab="Prop reproductive investment put into packaging & dispersal tissues")

mean1 <- tmp2$mean
se1 <- tmp2$se

se_function(mean1,se1)
```