### 2015-09-08 leaf_weight_issue.Rmd
file looking at how leaf area by diameter plots

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=14, fig.height=7.5}
remake::make("export")

SummaryInd <- readRDS("export/SummaryInd.rds")
SummarySpp <- readRDS("export/SummarySpp.rds")
SummarySppAge <- readRDS("export/SummarySppAge.rds")
HarvestData <- readRDS("export/HarvestData.rds")
InvestmentByPart <- readRDS("export/Investment_FD_all.rds")

source("R/figures.R")
```

* Currently a slope isn't being calculated for "leaf vs diameter" only for whole plant and stem - leaves are then determined by subtraction. I think this is probably a mistake.  
* It would be better to fit a line through "leaf weight vs diameter" for the subset of individuals/segments that are linear and then apply a correction to the older plants  
* if helpful, I could create a lookup table that indicates which ageXsegment combinations to use for the linear regression for each species  
* then, for leaves, the "offset" would be calculated as the offset based on the single-point value of (total weight of leave, basal diameter), instead of applying a slope correction  
* In these plots  
    * points are colored by age  
    * basal segments (i.e. whole plant) values are larger and surrounded by a black ring
    * the first plot for each species is all individuals, while the second plot it individuals from the two oldest sites

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=15, fig.height=15}
par(mfcol=c(2,2), cex=1, omi=c(.3,.2,.1,.1), mai=c(1.1,1.1,.4,0.2)) 
data2 <- subset(HarvestData,start_end=="end")
data2 <- split(data2, data2$species)

data1 <- subset(HarvestData,start_end=="end"&age>8)
data1 <- split(data1, data1$species)

for(spp in names(data)) {

plot(leaf_weight~diameter,data2[[spp]],log="xy",pch=16,col=col.age(data[[spp]]$age),main=spp,cex=1.2)
points(leaf_weight~diameter,subset(data2[[spp]],segment==1),cex=1.5)

plot(leaf_weight~diameter,data1[[spp]],log="xy",pch=16,col=col.age(data[[spp]]$age),main=spp,cex=1.2)
points(leaf_weight~diameter,subset(data1[[spp]],segment==1),cex=1.5)

}
```