### 2015-09-10 tests.Rmd

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
yvar <- "shoot_leaf_area"
xvar <- "initial_leaf_area"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log10(x[[yvar]]) ~  (log10(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.3,
                    main=spp,col=col.age(data[[spp]]$age))
}

mtext("Shoot leaf area", 2, outer=TRUE,cex=1.2)
mtext("Leaf area", 1, outer=TRUE,cex=1.2)
legend("topleft",legend=labels.age(),col=col.age(),pch=16, cex=1.1,bty="n")
```

plot(repro_inv~growth_leaf,data=subset(SummaryInd,RA>0),log="xy",col=col.spp(species), pch=16)
mod <- lm(log(growth_stem_diameter)~RA_leaf_area,data=subset(SummaryInd,RA_leaf_area>0&RA_leaf_area<1.5))
words.top.right.logx(mod)

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
savpar(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(1.1,.7,.6,0.2)) 
yvar <- "total_inv"
xvar <- "leaf_area_0"

data <- subset(SummaryInd,age>2)

data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log10(x[[yvar]]) ~  (log10(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results
saved <- results[,1:2]

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.3, main=spp,col=col.age(data[[spp]]$age))
 }

mtext("Height growth (log-log)", 2, outer=TRUE,cex=1.2)
mtext("Reproductive investment", 1, outer=TRUE,cex=1.2)
legend("topleft",legend=labels.age(),col=col.age(),pch=16, cex=1.1,bty="n")
```

SummaryInd$leaf_inv_ratio = SummaryInd$growth_stem / SummaryInd$growth_leaf

```{r}
par(mfrow=c(2,2), cex=1, omi=c(.7,.7,.1,.1), mai=c(.3,.3,.6,0.2)) 
data <- subset(SummaryInd)
data <- split(data, data$species)

for(spp in names(data)) {
plot(leaf_weight~diameter,data[[spp]],log="xy",cex=1.3, main=spp,col=col.age(data[[spp]]$age),pch=16)
points(leaf_weight_0~diameter_0,data[[spp]],cex=1.3,pch=1,col=col.age(data[[spp]]$age))
arrows(data[[spp]]$diameter, data[[spp]]$leaf_weight, data[[spp]]$diameter_0,data[[spp]]$leaf_weight_0, length = 0,angle = 30, code = 2, col = col.age(data[[spp]]$age),lwd=.3)
}
```

```{r}
par(mfrow=c(2,2), cex=1, omi=c(.7,.7,.1,.1), mai=c(1.1,1.1,.6,0.2)) 
data <- subset(SummaryInd)
data <- split(data, data$species)

for(spp in names(data)) {
plot(stem_weight~diameter,data[[spp]],log="xy",cex=1.3, main=spp,col=col.age(data[[spp]]$age),pch=16)
points(stem_weight_0~diameter_0,data[[spp]],cex=1.3,pch=1,col=col.age(data[[spp]]$age))
arrows(data[[spp]]$diameter, data[[spp]]$stem_weight, data[[spp]]$diameter_0,data[[spp]]$stem_weight_0, length = 0,angle = 30, code = 2, col = col.age(data[[spp]]$age),lwd=.3)
points(leaf_weight~diameter,data[[spp]],cex=1.3,pch=17,col=col.age(data[[spp]]$age))
points(leaf_weight_0~diameter_0,data[[spp]],cex=1.3,pch=2,col=col.age(data[[spp]]$age))
arrows(data[[spp]]$diameter, data[[spp]]$leaf_weight, data[[spp]]$diameter_0,data[[spp]]$leaf_weight_0, length = 0,angle = 30, code = 1, col = col.age(data[[spp]]$age),lwd=.3,lty=3)
}
```

```{r}
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(log(growth_inv)~log(repro_inv),data=subset(SummaryInd,repro_inv>0))
plot((1+repro_inv)~growth_inv, SummaryInd, log="xy", col=col.spp(species),pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
words.top.left.logxy(mod)
legend("bottomright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5")
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75")
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9")
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95")
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1")
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01")

#mod <- lm(log(growth_leaf)~log(repro_inv),data=subset(SummaryInd,repro_inv>0))
plot((1+repro_inv)~growth_leaf, SummaryInd, log="xy", col=col.spp(species),pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
#words.top.left.logxy(mod)
legend("bottomright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5")
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75")
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9")
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95")
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1")
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01")
```

plot(growth_leaf~growth_stem,SummaryInd, log="xy", col=col.age(age),pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
abline(0,1)