### 2015-09-11 Mark_20150911.Rmd
small changes

SummaryInd$leaf_turnover <- SummaryInd$leaf_weight_0/SummaryInd$LL_death
SummaryInd$leaf_net <- SummaryInd$growth_leaf + SummaryInd$leaf_turnover
SummaryInd$RA_leaf_turnover <- SummaryInd$repro_inv /(SummaryInd$repro_inv + SummaryInd$leaf_net)

pdf("figs2/lab_meeting_followup.pdf", height=11.3, width=8)
plot
```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfrow=c(4,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1,1,.1,0.1)) 
yvar <- "LMA"
xvar <- "age"

data <- subset(SummaryInd)
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",cex=1, xlab="age",ylab="LMA",main=spp,col=col.age(data[[spp]]$age))
}
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}

par(mfcol=c(4,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1,1,.1,0.1)) 
yvar <- "RA_leaf_turnover"
xvar <- "age"

data <- subset(SummaryInd)
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",cex=1, xlab="age",ylab="",main=spp,col=col.age(data[[spp]]$age))
}
mtext("RA, including leaf turnover", 2, outer=TRUE,cex=1.2,line=-.6)

```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
par(mfcol=c(4,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1,1,.1,0.1)) 
yvar <- "leaf_turnover"
xvar <- "leaf_net"

data <- subset(SummaryInd)
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",cex=1, xlab="",ylab="",main=spp,col=col.age(data[[spp]]$age))
  abline(0,1)
  abline(log10(.111),1)
  abline(log10(.010101),1)
  mtext("net leaf production (mg)", 1, outer=TRUE,cex=1.2,line=-.6)   
  mtext("leaf tunover (mg)", 2, outer=TRUE,cex=1.2,line=-.6)   
}
```
dev.off()