### 2015-12-03 Flowering_parts_charts.Rmd
small changes

Always run
Opening files

``` {r}
accessoryList <- read.csv("~/RA schedules - Kuringgai research/Reproductive_Allocation_Kuringgai/data/accessoryList.csv")
accessoryListShort <- read.csv("~/RA schedules - Kuringgai research/Reproductive_Allocation_Kuringgai/data/accessoryListShort.csv")
accessoryListShortest <- read.csv("~/RA schedules - Kuringgai research/Reproductive_Allocation_Kuringgai/data/accessoryListShortest.csv")
tissue_energy <- read.csv("~/RA schedules - Kuringgai research/Reproductive_Allocation_Kuringgai/data/tissue_energy.csv")
nectar <- read.csv("~/RA schedules - Kuringgai research/Reproductive_Allocation_Kuringgai/data/nectar.csv")

col.barplot <- function(x=NULL){
  ret <- c("black","cyan4","cyan1","darkslategray3","firebrick","firebrick1","firebrick2","firebrick3","gold","bisque","chartreuse2","chartreuse3","chartreuse4","chartreuse2","deepskyblue",
           "dodgerblue2","deepskyblue","deepskyblue4","darkorange","darkorange1","darkorange2","darkorange3","darkorange4","darkorange3")
  names(ret) <- c("nectar","seed","seed_pod","fruit_mature","bud","petals","stigma","style","finished_flower","calyx","fruit_immature","seed_immature","seed_aborted","fruit_seed_immature","bract","pedicel","green_parts","inflorescence_stalk","cone_aborted","cone_base_brown","cone_base_green","cone_brown","cone_green","cone")

  if(!is.null(x)) {
    ret <- ret[as.character(x)]
  }
  ret
}

labels.barplot <- function(x=NULL){
  names(col.barplot(x))
}

Invest2 <- merge(InvestmentByPart,select(accessoryList,-sort), by=c("species","part"))

for(v in c("weight")) {
  i <- is.na(Invest2[[v]])
  Invest2[[v]][i] <- 0
}
```
View(Invest2)

Always run 
process nectar
```{r}
Invest_nectar <-subset(Invest2,(species=="BAER"|species=="GRBU"|species=="GRSP"|species=="HATE")&part=="flower_stigma")
Invest_nectar$part ="nectar"
Invest_nectar$parts_fewer ="nectar"
Invest_nectar$parts_fewer_yet ="nectar"
Invest_nectar$weight = 0

SummaryNectar <- 
  Invest_nectar %>%
  group_by(parts_fewer_yet,species) %>%
  summarise_each(funs(sum),count)

SummaryNectar<- merge(SummaryNectar,nectar, by=c("species"),all.x=TRUE)
SummaryNectar$tissue_energy <- SummaryNectar$count*SummaryNectar$days*SummaryNectar$energy_daily/1000
SummaryNectar <- select(SummaryNectar,-count,-days,-energy_daily, -part)
SummaryNectar$accessory_part = "nectar"
```

by "parts fewer"
```{r}
SummaryInvest <- 
  Invest2 %>%
  group_by(parts_fewer,species) %>%
  summarise_each(funs(sum),weight)

SummaryInvest <- merge(SummaryInvest,select(accessoryListShort,-sort), by=c("species","parts_fewer"),all.x=TRUE)

SummaryInvestSpp <- 
  Invest2 %>%
  group_by(species) %>%
  summarise_each(funs(sum),weight)

names(SummaryInvestSpp) <- c("species","weight_spp")

SummaryInvest <- merge(SummaryInvest,SummaryInvestSpp,by=c("species"),all.x=TRUE)
SummaryInvest$prop_weight <- round(SummaryInvest$weight / SummaryInvest$weight_spp,digits=3)
SummaryInvest$labels <- paste(SummaryInvest$parts_fewer," (",SummaryInvest$prop_weight,")")
SummaryInvest <- SummaryInvest[order(factor(SummaryInvest$parts_fewer,levels=c(c("seed","seed_pod","fruit_mature","bud","petals","stigma","style","finished_flower","calyx","fruit_immature","seed_immature","seed_aborted",
                                                                                 "bract","pedicel","inflorescence_stalk","cone_aborted","cone_base_brown","cone_base_green","cone_brown","cone_green")))),]
```

by "part"
```{r}
SummaryInvest <- 
  Invest2 %>%
  group_by(part,species) %>%
  summarise_each(funs(sum),weight)

#SummaryInvest <- merge(SummaryInvest,select(accessoryListShort,-sort), by=c("species","parts_fewer"),all.x=TRUE)

SummaryInvestSpp <- 
  Invest2 %>%
  group_by(species) %>%
  summarise_each(funs(sum),weight)

names(SummaryInvestSpp) <- c("species","weight_spp")

SummaryInvest <- merge(SummaryInvest,SummaryInvestSpp,by=c("species"),all.x=TRUE)
SummaryInvest$prop_weight <- round(SummaryInvest$weight / SummaryInvest$weight_spp,digits=3)
#SummaryInvest$labels <- paste(SummaryInvest$parts_fewer," (",SummaryInvest$prop_weight,")")
#SummaryInvest <- SummaryInvest[order(factor(SummaryInvest$parts_fewer,levels=c(c("seed","seed_pod","fruit_mature","bud","petals","stigma","style","finished_flower","calyx","fruit_immature","seed_immature","seed_aborted","bract","pedicel","inflorescence_stalk","cone_aborted","cone_base_brown","cone_base_green","cone_brown","cone_green")))),]
```

View(SummaryInvest)

by "parts fewer yet"
```{r}
SummaryInvest <- 
  Invest2 %>%
  group_by(parts_fewer_yet,species) %>%
  summarise_each(funs(sum),weight)

SummaryInvest <- merge(SummaryInvest,select(accessoryListShortest,-sort), by=c("species","parts_fewer_yet"),all.x=TRUE)

SummaryInvestSpp <- 
  Invest2 %>%
  group_by(species) %>%
  summarise_each(funs(sum),weight)

names(SummaryInvestSpp) <- c("species","weight_spp")

SummaryInvest <- merge(SummaryInvest,SummaryInvestSpp,by=c("species"),all.x=TRUE)
SummaryInvest$prop_weight <- round(SummaryInvest$weight / SummaryInvest$weight_spp,digits=3)

SummaryInvest <- merge(SummaryInvest,tissue_energy,by=c("parts_fewer_yet"),all.x=TRUE)
SummaryInvest$tissue_energy <- SummaryInvest$weight*SummaryInvest$energy/1000

SummaryInvest <- bind_rows(SummaryInvest,SummaryNectar)

SummaryEnergySpp <- 
  SummaryInvest %>%
  group_by(species) %>%
  summarise_each(funs(sum),tissue_energy)

names(SummaryEnergySpp) <- c("species","energy_spp")
SummaryInvest <- merge(SummaryInvest,SummaryEnergySpp,by=c("species"),all.x=TRUE)
SummaryInvest$prop_energy <- round(SummaryInvest$tissue_energy / SummaryInvest$energy_spp,digits=3)

SummaryInvest$labels <- paste(SummaryInvest$parts_fewer_yet," (",SummaryInvest$prop_weight,")")
SummaryInvest <- SummaryInvest[order(factor(SummaryInvest$parts_fewer_yet,levels=c(c("nectar","bud","petals","calyx","finished_flower","fruit_seed_immature","seed","seed_pod","fruit_mature","bract","pedicel","inflorescence_stalk","green_parts","cone")))),]
```

View(SummaryInvest)
View(Invest2)

by "parts"
```{r}
SummaryInvestParts <- 
  Invest2 %>%
  group_by(part,species) %>%
  summarise_each(funs(sum),weight)

SummaryInvestParts <- merge(SummaryInvestParts,select(accessoryList,-sort), by=c("species","part"),all.x=TRUE)

SummaryInvestSppParts <- 
  Invest2 %>%
  group_by(species) %>%
  summarise_each(funs(sum),weight)

names(SummaryInvestSppParts) <- c("species","weight_spp")

SummaryInvestParts <- merge(SummaryInvestParts,SummaryInvestSppParts,by=c("species"),all.x=TRUE)

names(SummaryInvestParts) <- c("species","part","weight_by_part","accessory_part","parts_fewer","parts_fewer_yet","weight_spp")

SummaryInvestParts <- merge(SummaryInvestParts,select(SummaryInvest,-accessory_part,-weight_spp),by=c("species","parts_fewer_yet"),all.x=TRUE)
SummaryInvestParts$prop_weight_few_parts = round(SummaryInvestParts$weight_by_part / SummaryInvestParts$weight,digits=3)
SummaryInvestParts$labels <- paste(SummaryInvestParts$part," (",SummaryInvestParts$prop_weight_few_parts,")")
SummaryInvestParts$splitting <-paste(SummaryInvestParts$species,"_",SummaryInvestParts$parts_fewer_yet)
```

unique(SummaryInvestParts$parts_fewer)

``` {r}
pdf("figs2/Dividing_flower_parts2.pdf", height=11, width=8)
plot
par(mfcol=c(4,3), cex=1, omi=c(.01,.01,.01,.01), mai=c(.5,.9,.5,.9)) 
data <- subset(SummaryInvestParts,weight>0)
data <- split(data, data$splitting)

for(spp in names(data)) {
  pie(data[[spp]]$weight_by_part,labels=data[[spp]]$labels,main=spp,cex=0.3,cex.main=0.5,col=col.age())
}
dev.off()
```

#barplot by dry weight
```{r}
barplot_data <- select(SummaryInvest,species, parts_fewer_yet, prop_weight)
barplot_data <- spread(barplot_data,species,prop_weight,fill=0,drop=FALSE)
barplot_data <- barplot_data[order(factor(barplot_data$parts_fewer_yet,levels=c(c(
  "bud","petals","stigma","style","finished_flower","calyx","fruit_immature","seed_immature","seed_aborted","fruit_seed_immature","seed","seed_pod","fruit_mature",
  "bract","pedicel","inflorescence_stalk","green_parts","cone_aborted","cone_base_brown","cone_base_green","cone_brown","cone_green","cone")))),]

barplot_data_matrix <- as.matrix(barplot_data[2:15])

#pdf("figs2/Flower_parts_barchart.pdf", height=8, width=11)
#plot
#win.metafile("figs2/RA_talk_barplot.wmf", height=4, width=10)
#plot
par(mfcol=c(1,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
barplot(barplot_data_matrix,beside=FALSE,width=0.8,cex.axis=.7,cex.names=.7,xlim=c(.5,17.2),col=col.barplot(barplot_data$parts_fewer_yet),ylab="proportion repro investment")
legend(13.5,1,legend=barplot_data$parts_fewer_yet,pch=16, cex=1,bty="n",col=col.barplot(barplot_data$parts_fewer_yet))

#dev.off()
```

#barchart by kJ
```{r}
barplot_data <- select(SummaryInvest,species, parts_fewer_yet, prop_energy)
barplot_data <- spread(barplot_data,species,prop_energy,fill=0,drop=FALSE)
barplot_data <- barplot_data[order(factor(barplot_data$parts_fewer_yet,levels=c(c("nectar",
  "bud","petals","stigma","style","finished_flower","calyx","fruit_immature","seed_immature","seed_aborted","fruit_seed_immature","seed","seed_pod","fruit_mature",
  "bract","pedicel","inflorescence_stalk","green_parts","cone_aborted","cone_base_brown","cone_base_green","cone_brown","cone_green","cone")))),]

barplot_data_matrix <- as.matrix(barplot_data[2:15])

pdf("figs2/Flower_parts_barchart_energy.pdf", height=8, width=11)

#win.metafile("figs2/RA_talk_barplot_energy.wmf", height=4, width=10)
plot
par(mfcol=c(1,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
barplot(barplot_data_matrix,beside=FALSE,width=0.8,cex.axis=.7,cex.names=.7,xlim=c(.5,17.2),col=col.barplot(barplot_data$parts_fewer_yet),ylab="proportion repro investment")
legend(13.5,1,legend=barplot_data$parts_fewer_yet,pch=16, cex=1,bty="n",col=col.barplot(barplot_data$parts_fewer_yet))

dev.off()
```

##For pie charts for overall species allocation
```{r}
data <- subset(SummaryInvest,weight>0)
data <- split(data, data$species)

pdf("figs2/Flower_parts_species_all.pdf", height=11, width=8)
plot
par(mfrow=c(3,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
for(spp in names(data)) {
  pie(data[[spp]]$weight,labels=data[[spp]]$part,main=spp,col=col.barplot(data[[spp]]$parts_fewer_yet),cex=0.8)
}
dev.off()
```

##For pie charts
```{r}
data <- subset(SummaryInvest,weight>0)
data <- split(data, data$species)

pdf("figs2/Flower_parts_fewer.pdf", height=11, width=8)
plot
par(mfrow=c(2,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
for(spp in names(data)) {
  pie(data[[spp]]$weight,labels=data[[spp]]$labels,main=spp,col=col.barplot(data[[spp]]$parts_fewer_yet),cex=0.8)
}
#dev.off()
```