### 2016-03-08 Accessory_figs_20160218.Rmd
work on accessory figures to show how cost of buds vs dispersal shifts with seeds size

#seed count, scaled to plant size, decreases 1:1 with seed size, while bud count decreases less steeply; indication that seed count is what is actually "programmed in"
```{r}

win.metafile("figs2/Accessory_Costs/Figure_costs_of_flowers_vs_seed_size.wmf", height=10, width=10)
plot

# species with bigger seeds abort a disprortionate # of buds
# plot(seed_size~scaled_failure_count,SummarySpp[["mean"]],pch=16,log="xy",col=col.spp(species),ylim=c(.01,1000),xlim=c(.00001,100))
# mod <- sma(seed_size~scaled_failure_count,data=subset(SummarySpp[["mean"]]),log="xy",slope=-1,method="OLS")
# sma_abline(mod)
# sma_sum(mod)
# 
# species with bigger seeds produce disproportionately few seeds
# plot(seed_size~scaled_seed_count,SummarySpp[["mean"]],pch=16,log="xy",col=col.spp(species),ylim=c(.01,1000),xlim=c(.00001,100))
# mod <- sma(seed_size~scaled_seed_count,data=subset(SummarySpp[["mean"]]),log="xy",slope=1,method="OLS")
# sma_abline(mod)
# sma_sum(mod)

par(mfcol=c(2,2), cex=1, omi=c(.1,.1,.1,.05), mai=c(1,1,.1,.05))
options(scipen = 4)

#put together, big-seeded plants produce lots more buds than predicted from 1:1 relationship, but the "correct" number of seeds
plot(scaled_seed_count~seed_size,SummaryInd,pch=16,log="xy",col=col.spp(species),xlim=c(.01,1000),ylim=c(1e-6,10),yaxt="n")
points(scaled_bud_count~seed_size,SummaryInd,pch=21,col=col.spp(species))
points(scaled_failure_count~seed_size,SummaryInd,pch=2,col=col.spp(species))
axis(2,yaxp=c(0.00001,0.0001,0.01,1))
mod1 <- sma(scaled_seed_count~seed_size,data=subset(SummaryInd,scaled_seed_count!=0),log="xy",slope=-1,method="OLS")
mod2 <- sma(scaled_bud_count~seed_size,data=subset(SummaryInd,scaled_bud_count!=0),log="xy",slope=-1,method="OLS")
mod3 <- sma(scaled_failure_count~seed_size,data=subset(SummaryInd,scaled_failure_count!=0),log="xy",slope=-1,method="OLS")
sma_abline(mod1)
abline(mod2$coef[[1]][1,1],mod2$coef[[1]][2,1],col="blue")
abline(mod3$coef[[1]][1,1],mod3$coef[[1]][2,1],col="red")

plot(0,0,type="n", axes=FALSE, ann=FALSE)

#species that abort a disprortionate # of buds can do so because the relative cost of buds (relative to seeds, dispersal) is small
plot(scaled_failure_count~prop_prepollencosts,SummarySpp[["mean"]],pch=2,log="y",col=col.spp(species),xlim=c(-.2,1),ylim=c(1e-6,10),cex=1.5)
mod <- sma(scaled_failure_count~prop_prepollencosts,data=subset(SummarySpp[["mean"]]),log="y",slope=1,method="OLS")
sma_abline(mod)
sma_sum(mod)
par("yaxp")

#species that produce a disproportionte number of seeds can do so because they have proportionately lower dispersal costs
plot(scaled_seed_count~prop_dispersalcosts,SummarySpp[["mean"]],pch=16,log="y",col=col.spp(species),xlim=c(-.2,1),ylim=c(1e-6,10),cex=1.5)
mod <- sma(scaled_seed_count~prop_dispersalcosts,data=subset(SummarySpp[["mean"]]),log="y",slope=1,method="OLS")
sma_abline(mod)
sma_sum(mod)

?par

plot(x,y, xaxt="n")
axis(1, xaxp=c(10, 200, 19), las=2

sma_sum(mod1)
sma_sum(mod2)
sma_sum(mod3)
dev.off()
```

#Neither within nor across species, does the proportion of reproductive energy spent on "success" correlate with total reproductive investment
```{r}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.05))

xvar <- "repro_inv"
yvar <- "prop_success"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], log="x",method="OLS")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),xlab="",ylab="",lty=results$linetype,pch=16,log="x")
```

#Table 1. correlations with total repro inv
```{r}
results <- as.data.frame(matrix(vector(), 0, 8,
                dimnames=list(c(), c("variable","r2","pval","Slope","Slope_lowCI","Slope_highCI","Int","Slope_test_p"))),
                stringsAsFactors=F)

for (i in c("total_weight","failure_inv","success_inv","propagule_inv","packaging_dispersal_inv","prepollen_success_inv","fruit_inv","flower_inv")) {
yvar <- i
xvar <- "repro_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)
fit <- sma(data[[yvar]] ~ data[[xvar]], method="OLS", log="xy",slope=1)
temp <- as.data.frame(c(yvar,select(fit$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Slope_test_p)))
colnames(temp)[1]<- "variable"
results <- bind_rows(results, temp)
}

write.csv(results, file="figs2/Accessory_Costs/Table02correlations_with_repro_inv.csv",row.names=FALSE)
```

#Figure 1. Correlations of various reproductive investment categories against plant weight
```{r}
win.metafile("figs2/Accessory_Costs/Figure01_correlations_with_plant_weight.wmf", height=10, width=10)
plot

par(mfrow=c(2,2), cex=1, omi=c(.1,.6,.1,.05), mai=c(.8,.1,.1,.05))
options(scipen = 4)

xvar <- "total_weight_0"
yvar <- "repro_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),log="xy",ylim=c(1,300000),xlim=c(1,10000000),xlab="",ylab="",lty=results$linetype,pch=16)
mod <- lm(log10(total_weight_0)~log10(repro_inv),data=subset(SummaryInd,repro_inv>0))
abline(0,1)
words.top.left.logxy(mod)
mtext(text="total reproductive investment (mg)",side=2,line=2,outer=TRUE)
mtext(text="total plant weight (mg)",side=1,line=2,outer=FALSE)

yvar <- "repro_inv"
xvar <- "propagule_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),log="xy",ylim=c(1,300000),xlim=c(1,300000), xlab="",ylab="",lty=results$linetype,pch=16)
abline(0,1)
mod <- lm(log10(propagule_inv)~log10(repro_inv),data=subset(SummaryInd,propagule_inv>0))
words.top.left.logxy(mod)
mtext(text="investment in propagules (mg)",side=1,line=2,outer=FALSE)

xvar <- "repro_all_count"
yvar <- "repro_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy",slope.test=1)
results <- lines(fit$groupsummary)
plot(fit,xlab="",ylab="",pch=16, col=col.spp(fit$groupsummary$group),lty=results$linetype,ylim=c(1,300000),xlim=c(1,30000))
mtext(text="buds initiated",side=1,line=2,outer=FALSE)

xvar <- "flower_inv"
yvar <- "repro_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy",slope.test=1)
results <- lines(fit$groupsummary)
plot(fit,xlab="",ylab="",pch=16, col=col.spp(fit$groupsummary$group),lty=results$linetype,ylim=c(1,300000),xlim=c(1,300000))
abline(0,1)
mtext(text="flower weight * bud count",side=1,line=2,outer=FALSE)

dev.off()
```

#correlations between repro inv and propagule inv by species
```{r}
yvar <- "repro_inv"
xvar <- "propagule_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- fit$groupsummary

write.csv(results, file="figs2/Accessory_Costs/Table02extra_propagule_invest_vs_repro_inv.csv",row.names=FALSE)
```

#when reproductive investment is scaled by leaf area there is no significant difference by seed size - so increase in repro inv with seed size simply because species with bigger seeds are bigger plants  
```{r}
plot(invest_per_leaf_area~seed_size,SummaryInd, col=col.spp(species),pch=16,log="xy")
points(invest_per_leaf_area~seed_size,SummarySpp[["mean"]], cex=2, col=col.spp(species),pch=16,log="xy")
mod <-sma(invest_per_leaf_area~seed_size,SummarySpp[["mean"]],log="xy",method="OLS")
summary(mod)
```

#Consider how accessory cost components shift in terms of costs per seed, and cost as fraction total reproductive investment

```{r}
win.metafile("figs2/Accessory_Costs/Figure02_seedsize_vs_reproinv.wmf", height=5, width=10)
plot

par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.1))

plot(repro_inv~seed_size,SummarySpp[["mean"]],log="xy",col=col.spp(species),pch=16)
mod <- lm(log10(repro_inv)~log10(seed_size),SummarySpp[["mean"]])
abline(mod$coeff[1],mod$coeff[2])
words.top.left.logxy(mod)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$repro_inv,SummarySpp[["se"]]$repro_inv)

plot(seedset~seed_size,SummarySpp[["mean"]],log="xy",col=col.spp(species),pch=16)
mod <- lm(log10(seedset)~log10(seed_size),SummarySpp[["mean"]])
abline(mod$coeff[1],mod$coeff[2])
words.bottom.left.logxy(mod)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$seedset,SummarySpp[["se"]]$seedset)

dev.off()
```

#Table showing accessory cost values for the species
```{r}
acc_costs_spp <- dplyr::select(SummarySpp[["mean"]],species,seed_size,seed_to_ovule_ratio,accessory_per_seed,total_accessory_Lord,prepollencosts,b_div_p,c_div_p,dispersalcosts_increment,dispersalcosts_increment2,prop_accessory_nonzero)
acc_costs_spp$species <- labels.spp.full()
acc_costs_spp <- format(acc_costs_spp, digit=2)

write.csv(acc_costs_spp, file="figs2/Accessory_Costs/Table02_AccCosts_by_Species.csv",row.names=FALSE)
```

#Figure 2 and Table 3. Scaling of accessory costs versus seed size, by species
```{r}
win.metafile("figs2/Accessory_Costs/Figure02_seed_size_correlations.wmf", height=5, width=20)
plot

par(mfrow=c(1,4), cex=1, omi=c(.6,.1,.05,.05), mai=c(.1,.8,.05,.05)) 
options(scipen = 5)

yvar <- "seedcosts"
xvar <- "seed_size"
data <- subset(SummarySpp[["mean"]])
plot(seedcosts~seed_size,data=data,col=col.spp(species),pch=16,log="xy",cex=1.5,xlab="",ylab="all costs to make 1 seed (mg) = seedcosts",ylim=c(.02,20000), xlim=c(.02,20000))
slopes <- sma(data[[yvar]] ~ data[[xvar]], method="OLS", slope.test=1,log="xy")
mod <- lm(log10(seedcosts)~log10(seed_size),SummarySpp[["mean"]])
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
#abline(slopes$groupsummary["Int"][[1]],1,col="red")
abline(0,1,col="red")
words.top.left.logxy(mod)

plot(accessory_per_seed ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,20000), xlim=c(.02,20000), ylab="total per seed accessory costs (mg/seed)",xlab="",cex=1.5)
mod <- lm(log10(accessory_per_seed) ~ log10(seed_size),data=subset(SummarySpp[["mean"]],accessory_per_seed>0))
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$accessory_per_seed,SummarySpp[["se"]]$accessory_per_seed)
slopes <- sma(accessory_per_seed ~ seed_size,data=subset(SummarySpp[["mean"]]),slope.test=1,log="xy")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
#abline(slopes$groupsummary["Int"][[1]],1,col="red")
abline(0,1,col="red")
words.bottom.left.logxy(mod)

plot(repro_inv~seed_size,SummarySpp[["mean"]],log="xy",ylab="reproductive investment (mg)",col=col.spp(species),pch=16, ylim=c(.02,20000), xlim=c(.02,20000),cex=1.5)
mod <- lm(log10(repro_inv)~log10(seed_size),data=subset(SummarySpp[["mean"]],species!="EPMI"))
abline(mod$coeff[1],mod$coeff[2])
words.bottom.left.logxy(mod)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$repro_inv,SummarySpp[["se"]]$repro_inv)
abline(0,1,col="red")

plot(seedset~seed_size,SummarySpp[["mean"]],log="xy",col=col.spp(species),pch=16, xlim=c(.02,20000),cex=1.5)
mod <- lm(log10(seedset)~log10(seed_size),SummarySpp[["mean"]])
abline(mod$coeff[1],mod$coeff[2])
words.bottom.left.logxy(mod)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$seedset,SummarySpp[["se"]]$seedset)

mtext(text="seed size (mg)",side=1,line=2,outer=TRUE)

dev.off()
```

```{r}
plot(accessory_per_seed ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), xlim=c(.02,10000), ylab="total per seed accessory costs (mg/seed)",xlab="",cex=1.5)
mod <- lm(log10(accessory_per_seed) ~ log10(seed_size),data=subset(SummarySpp[["mean"]],accessory_per_seed>0))
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$accessory_per_seed,SummarySpp[["se"]]$accessory_per_seed)
slopes <- sma(log10(accessory_per_seed) ~ log10(seed_size),data=subset(SummarySpp[["mean"]]),slope.test=1)
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
#abline(slopes$groupsummary["Int"][[1]],1,col="red")
abline(0,1,col="red")
words.bottom.right.logxy(mod)

acc_by_species <- select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p)

plot(prepollencosts ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="prepollen success costs (mg/seed)",xlab="",cex=1.5)
mod <- lm(log10(prepollencosts) ~ log10(seed_size),data=subset(SummarySpp[["mean"]]))
words.bottom.right.logxy(mod)
extra.topleft.logxy("slope <1, so bigger seeded species have lower fixed costs of making pre-pollination parts")
slopes <- sma(log10(prepollencosts) ~ log10(seed_size),data=subset(SummarySpp[["mean"]]),slope.test=1)
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,col="red")
acc_by_species <- dplyr::bind_rows(acc_by_species, select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(prepollen_failure_per_seed ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="prepollen failure costs (mg/seed)",xlab="",cex=1.5)
mod <- lm(log10(prepollen_failure_per_seed) ~ log10(seed_size),data=subset(SummarySpp[["mean"]]))
words.bottom.right.logxy(mod)
extra.topleft.logxy("the per seed costs of failure scale isometrically with seed size")
slopes <- sma(log10(prepollen_failure_per_seed) ~ log10(seed_size),data=subset(SummarySpp[["mean"]]),slope.test=1)
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,col="red")
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$prepollen_failure_per_seed,SummarySpp[["se"]]$prepollen_failure_per_seed)
acc_by_species <- dplyr::bind_rows(acc_by_species, select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(c_div_p ~ seed_size,data=subset(SummarySpp[["mean"]],species!="COER"),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="aborted fruit accessory costs (mg/seed)",xlab="",cex=1.5,xaxt="n")
mod <- lm(log10(c_div_p) ~ log10(seed_size),data=subset(SummarySpp[["mean"]],accessory_per_seed>0&species!="COER"&species!="BAER"))
words.top.left.logxy(mod)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$c_div_p,SummarySpp[["se"]]$c_div_p)
slopes <- sma(log10(c_div_p) ~ log10(seed_size),data=subset(SummarySpp[["mean"]],species!="COER"&species!="BAER"),slope.test=1)
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,col="red")
acc_by_species <- dplyr::bind_rows(acc_by_species, select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(dispersalcosts_increment ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="dispersal & packaging accessory costs (mg/seed)",xlab="",cex=1.5)
mod <- lm(log10(dispersalcosts_increment) ~ log10(seed_size),data=subset(SummarySpp[["mean"]],species!="LEES"))
words.top.left.logxy(mod)
slopes <- sma(log10(dispersalcosts_increment) ~ log10(seed_size),data=subset(SummarySpp[["mean"]]),slope.test=1)
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,col="red")
acc_by_species <- dplyr::bind_rows(acc_by_species, select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))
mtext(text="seed size (mg)",side=1,line=1.5,outer=TRUE)

acc_by_species$variable = c("total accessory costs","prepollen_success","prepollen_failure","c div p","dispersal")

write.csv(acc_by_species, file="figs2/Accessory_Costs/Table03_acc_by_species.csv",row.names=FALSE)
```

#species with bigger seeds, lower seedset spend less per flower. Here taking the prop of the cost to make a single flower that goes into pre-pollination parts
```{r}
win.metafile("figs2/Accessory_Costs/Figure_seedcosts_vs_seedsize.wmf", height=5, width=5)
plot

par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.1))

yvar <- "seedcosts"
xvar <- "seed_size"
data <- subset(SummarySpp[["mean"]])
plot(seedcosts~seed_size,data=data,col=col.spp(species),pch=16,log="xy",cex=1.5,xlab="seed size (mg)",ylab="all costs to make 1 seed (mg) = seedcosts",ylim=c(.02,10000), xlim=c(.02,10000))
slopes <- sma(data[[yvar]] ~ data[[xvar]], method="OLS", slope.test=1,log="xy")
mod <- lm(log10(seedcosts)~log10(seed_size),SummarySpp[["mean"]])
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
#abline(slopes$groupsummary["Int"][[1]],1,col="red")
abline(0,1,col="red")
extra.topleft.logxy("total accessory costs required to produce 1 seed are lower for smaller seeded species")
words.bottom.right.logxy(mod)
seedcosts_by_species <- select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p)
dev.off()
```

```{r}
xvar <- "seed_size"
yvar <- "prop_prepollencosts"
data <- subset(SummarySpp[["mean"]])
plot(prop_prepollencosts~seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy",cex=1.5,xlab="",ylab="prop seedcosts for prepollination")
extra.topleft.logxy("species with big seeds have low costs of producing a single flower")
slopes <- sma(data[[yvar]] ~ data[[xvar]], method="OLS", slope.test=1,log="xy")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
# abline(slopes$groupsummary["Int"][[1]],-1,col="red")
seedcosts_by_species <- dplyr::bind_rows(seedcosts_by_species, select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

yvar <- "prop_dispersalcosts"
plot(prop_dispersalcosts~seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy",cex=1.5,xlab="",ylab="prop seedcosts for dispersal")
extra.topleft.logxy("species with big seeds have high packaging and dispersal costs")
slopes <- sma(data[[yvar]] ~ data[[xvar]], method="OLS", slope.test=1,log="xy")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
# abline(slopes$groupsummary["Int"][[1]],1,col="red")
seedcosts_by_species <- dplyr::bind_rows(seedcosts_by_species, select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))
mtext(text="seed size (mg)",side=1,line=2,outer=TRUE)

seedcosts_by_species$variable = c("total seedcosts","prepollen costs (% of total)","dispersal costs (% of total)")
write.csv(seedcosts_by_species, file="figs2/Accessory_Costs/Table04_seedcosts_by_species.csv",row.names=FALSE)
```

```{r}
win.metafile("figs2/Accessory_Costs/Figure_seedset_vs_seedsize.wmf", height=5, width=5)
plot

par(mfrow=c(1,1), cex=1, omi=c(1,1,.1,.1), mai=c(1,1,.5,.5)) 
plot(seedset~seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy",cex=2)
mod <- lm(log10(seed_size)~log10(seedset),data=subset(SummarySpp[["mean"]]))
#abline(mod$coefficients[1],mod$coefficients[2])
words.top.left.logxy(mod)

dev.off()
```

#Table 4 and Figure 3.Per seed accessory costs are lower in individuals with higher seed count. 
###This effect is true across species, since larger-seeded species produce fewer seeds and have higher per seed accessory costs and true within species, since absolute accessory investment is a function of plant size not seedset
```{r}
win.metafile("figs2/Accessory_Costs/Figure_EXTRA_acc_cost_by_seed_count.wmf", height=5, width=5)
plot

par(mfrow=c(1,1), cex=1, omi=c(1,1,.1,.1), mai=c(1,1,.5,.5))

xvar <- "seed_count"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab=xvar,ylab=yvar,lty=results$linetype)

points(accessory_per_seed~seed_count,data=subset(SummarySpp[["mean"]],accessory_per_seed>0&seed_count>0),col="black",pch=16,cex=2.5)
points(accessory_per_seed~seed_count,data=subset(SummarySpp[["mean"]],accessory_per_seed>0&seed_count>0),col=col.spp(species),pch=16,cex=2)
mod <- lm(log10(accessory_per_seed)~log10(seed_count),data=subset(SummarySpp[["mean"]]))
words.top.right.logxy(mod)
legend_with_r2(results,"bottomleft")

results
write.csv(results, file="figs2/Accessory_Costs/Table04a_seedcount_vs_accessory.csv",row.names=FALSE)

dev.off()
```

```{r}
win.metafile("figs2/Accessory_Costs/Figure_acc_cost_by_seedset.wmf", height=5, width=5)
plot

par(mfrow=c(1,1), cex=1, omi=c(.5,.1,.1,.1), mai=c(.1,1,.3,.1))

xvar <- "seed_size"
yvar <- "seedcosts"

plot(seedcosts~seed_size,data=subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0),log="xy",pch=16,col=col.spp(species),cex=2)
mod <-lm(log10(seedcosts)~log10(seed_size),SummaryInd)
abline(mod$coeff[1],mod$coeff[2])
words.top.left.logxy(mod)

#text((10^(par()$usr[1])*1.4),(10^(par()$usr[4])*.8),summary(mod)$coefficients[1,1])
dev.off()
```

#Table 4 extra - showing propagule investment instead of seed_count. Could also replace with seedset
```{r}
win.metafile("figs2/Accessory_Costs/Figure_acc_cost_by_seedset.wmf", height=9, width=9)
plot

par(mfrow=c(2,2), cex=1, omi=c(.5,.1,.1,.1), mai=c(.1,1,.3,.1))

xvar <- "seedset"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab="total accessory costs (mg/seed)",lty=results$linetype,xlim=c(.0005,.5),ylim=c(.01,200000),xaxt="n")
legend_with_r2(results,"bottomleft")

write.csv(results, file="figs2/Accessory_Costs/Table04b_seedset_vs_accessory.csv",row.names=FALSE)

xvar <- "seedset"
yvar <- "b_div_p"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(results$group),pch=16,xlab="",ylab="b div p = pre-pollination costs (mg/seed)",lty=results$linetype,xlim=c(.0005,.5),ylim=c(.01,200000),xaxt="n")
legend_with_r2(results,"bottomleft")

write.csv(results, file="figs2/Accessory_Costs/Table06a_b_div_p_vs_seedset.csv",row.names=FALSE)

xvar <- "seedset"
yvar <- "c_div_p"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy") 
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(results$group),pch=16,xlab="",ylab="c div p = aborted post pollination (mg/seed)",lty=results$linetype,xlim=c(.0005,.5),ylim=c(.01,200000))
legend_with_r2(results,"bottomleft")

write.csv(results, file="figs2/Accessory_Costs/Table06b_c_div_p_vs_seedset.csv",row.names=FALSE)

xvar <- "seedset"
yvar <- "packaging_dispersal_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(results$group),pch=16,xlab="",ylab=yvar,lty=results$linetype,xlim=c(.0005,.5),ylim=c(.01,200000))
legend_with_r2(results,"bottomleft")
mtext(text="seedset",side=1,line=1.5,outer=TRUE)

dev.off()
```

```{r}
xvar <- "seedset"
yvar <- "packaging_dispersal_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)

results <- select(results,group,r2,Slope)
names(results) <- c("species","r2","slope")

results$ovules_per_fruit <- c(60,4,1,1,2,2,2,4,1,40,2,20,2)
max_seedset <- select(SummarySpp[["max"]],species,seedset) 
min_seedset <- select(SummarySpp[["min"]],species,seedset) 
results <- merge(results, max_seedset, by=c("species"), all=TRUE)
results <- merge(results, min_seedset, by=c("species"), all=TRUE)
names(results)=c("species","r2","slope","ovules_per_fruit","max_seedset","min_seedset")

win.metafile("figs2/Accessory_Costs/Table07_dispcosts_vs_ovules.wmf", height=6, width=6)
plot

par(mfrow=c(1,1), cex=1, omi=c(.4,.4,.1,.1), mai=c(.8,.8,.2,.2)) 
plot(ovules_per_fruit~r2,data=subset(results),log="y",col=col.spp(species),pch=16,cex=2,ylab="ovules per fruit",xlab="slope between dispersal costs and seedset")
mod <- lm(log10(ovules_per_fruit)~r2,data=subset(results,species!="PELA"))
abline(mod$coeff[1],mod$coeff[2])
words.top.left.logy(mod)
dev.off()

write.csv(results, file="figs2/Accessory_Costs/Table06c_dispersal_vs_seedset.csv",row.names=FALSE)
```

#Higher RE, RA, plant size do not lead to lower accessory costs; age shows mixed results
##Not a single species shows a reduction in per seed accessory costs with increased plant size.Indeed, EPMI has higher pre-pollination accessory costs    (slope = 0.38, r2=0.19, p=0.019) and accessory costs due to aborted fruit and seeds (slope = 0.83, r2=0.268, p=0.019) in larger individuals, as larger individuals of this species have lower seedset (slope = -0.31, r2=0.198, p=0.018). GRSP also has a nearly significant decline in seedset in larger individuals (slope = -0.78, r2=0.298, p=0.067).
```{r}
win.metafile("figs2/Accessory_Costs/Figure_RE_RA_size_age_patterns.wmf", height=10, width=10)
plot

par(mfrow=c(2,2), cex=1, omi=c(.1,.4,.1,.05), mai=c(.6,.3,.1,.1))
options(scipen = 2)

xvar <- "repro_inv"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab=yvar,lty=results$linetype)
mtext(text="reproductive investment (mg) (RE)",side=1,line=2,outer=FALSE)
legend_with_r2(results,"bottomright")

write.csv(results, file="figs2/Accessory_Costs/Table05a_accessory_vs_RE.csv",row.names=FALSE)

xvar <- "RA"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0&SummaryInd$age>2)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab="",lty=results$linetype)
mtext(text="RA",side=1,line=2,outer=FALSE)
legend_with_r2(results,"topleft")

write.csv(results, file="figs2/Accessory_Costs/Table05b_accessory_vs_RA.csv",row.names=FALSE)

xvar <- "total_weight_0"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab=yvar,lty=results$linetype)
mtext(text="total plant size (mg)",side=1,line=2,outer=FALSE)
legend_with_r2(results,"bottomright")

write.csv(results, file="figs2/Accessory_Costs/Table05c_accessory_vs_total_weight.csv",row.names=FALSE)

xvar <- "age"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab="",lty=results$linetype)
mtext(text="age (years)",side=1,line=2,outer=FALSE)
mtext(text="per seed accessory costs (mg/seed)",side=2,line=1,outer=TRUE)
legend_with_r2(results,"topleft")

write.csv(results, file="figs2/Accessory_Costs/Table05d_accessory_vs_age.csv",row.names=FALSE)
dev.off()
```

```{r}
win.metafile("figs2/Accessory_Costs/Figure_seedcount_reprocount.wmf", height=5, width=5)
plot

par(mfrow=c(1,1), cex=1, omi=c(.1,0.1.1,0.05), mai=c(.8,.8,.1,0.1))
options(scipen = 2)

xvar <- "prepollen_all_count"
yvar <- "seed_count"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="buds initiated (count)",ylab="seed count",lty=results$linetype,xlim=c(0.9,20000),ylim=c(0.9,20000))

abline(0,1,lwd=2)
abline(-2,1,lwd=2)

legend_with_r2(results,"topleft")

summary(results)
summary(fit)

fit2 <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], log="xy",slope.text=1)
summary(fit2)

dev.off()
```
#Considering the individual components of accessory costs vs. seedset - since seedset best predictor of accessory costs

#For each species the accessory cost components that cause shifts in total accessory costs is different and reflects both which components are largest and whichvary most with some factor - seedset, age, etc.
```{r}
xvar1 <- "packaging_dispersal_per_seed"
xvar2 <- "b_div_p"
xvar3 <- "c_div_p"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar1]]>0&SummaryInd[[xvar2]]>0&SummaryInd[[xvar3]]>0&species!="BAER"&species!="COER"&species!="LEES")
data <- split(data, data$species)

mod <- lapply(data, function(x) lm(log10(x[[yvar]]) ~  (log10(x[[xvar1]]))+(log10(x[[xvar2]]))+(log10(x[[xvar3]]))))
mod_LEES <- lm(log10(accessory_per_seed) ~ (log10(b_div_p)+(log10(c_div_p))),data=subset(SummaryInd,SummaryInd[[xvar2]]>0&SummaryInd[[xvar3]]>0&species=="LEES"))
mod_BAER <- lm(log10(accessory_per_seed) ~ (log10(b_div_p)+(log10(packaging_dispersal_per_seed))),data=subset(SummaryInd,SummaryInd[[xvar2]]>0&SummaryInd[[xvar1]]>0&species=="BAER"))
mod_COER <- lm(log10(accessory_per_seed) ~ (log10(b_div_p)+(log10(packaging_dispersal_per_seed))),data=subset(SummaryInd,SummaryInd[[xvar2]]>0&SummaryInd[[xvar1]]>0&species=="COER"))

species_coef <- as.data.frame(c("intercept","packaging&dispersal","b_div_p","c_div_p"))

for(i in 1:11) {
     species_coef <- dplyr::bind_cols(species_coef, as.data.frame(mod[[i]]$coefficients))
     }

names(species_coef) <- c("effect","BOLE","EPMI","GRBU","GRSP","HATE","HEPU","PELA","PEPU","PHPH","PILI","PUTU")

BAER_c <- as.data.frame(mod_BAER$coefficients)
BAER_c$effect <- c("intercept","b_div_p","packaging&dispersal")
names(BAER_c) <- c("BAER","effect")
species_coef <- dplyr::left_join(species_coef,BAER_c,by="effect")

COER_c <- as.data.frame(mod_COER$coefficients)
COER_c$effect <- c("intercept","b_div_p","packaging&dispersal")
names(COER_c) <- c("COER","effect")
species_coef <- dplyr::left_join(species_coef,COER_c,by="effect")

LEES_c <- as.data.frame(mod_LEES$coefficients)
LEES_c$effect <- c("intercept","b_div_p","c_div_p")
names(LEES_c) <- c("LEES","effect")
species_coef <- dplyr::left_join(species_coef,LEES_c,by="effect")

# Reorder by column name
species_coef <- species_coef[c("effect", "BAER", "BOLE", "COER", "EPMI", "GRBU", "GRSP", "HATE", "HEPU", "LEES", "PELA", "PEPU", "PHPH", "PILI", "PUTU")]

write.csv(species_coef, file="figs2/Accessory_Costs/Table08_accessory_components_coeff.csv",row.names=FALSE)

species_t <- as.data.frame(c("intercept","packaging&dispersal","b_div_p","c_div_p"))

for(i in 1:11) {
     species_t <- dplyr::bind_cols(species_t, as.data.frame(summary(mod[[i]])$coefficient[,4]))
     }

names(species_t) <- c("effect","BOLE","EPMI","GRBU","GRSP","HATE","HEPU","PELA","PEPU","PHPH","PILI","PUTU")

BAER_t <- as.data.frame(summary(mod_BAER)$coefficient[,4])
BAER_t$effect <- c("intercept","b_div_p","packaging&dispersal")
names(BAER_t) <- c("BAER","effect")
species_t <- dplyr::left_join(species_t,BAER_t,by="effect")

COER_t <- as.data.frame(summary(mod_COER)$coefficient[,4])
COER_t$effect <- c("intercept","b_div_p","packaging&dispersal")
names(COER_t) <- c("COER","effect")
species_t <- dplyr::left_join(species_t,COER_t,by="effect")

LEES_t <- as.data.frame(summary(mod_LEES)$coefficient[,4])
LEES_t$effect <- c("intercept","b_div_p","c_div_p")
names(LEES_t) <- c("LEES","effect")
species_t <- dplyr::left_join(species_t,LEES_t,by="effect")

# Reorder by column name
species_t <- species_t[c("effect", "BAER", "BOLE", "COER", "EPMI", "GRBU", "GRSP", "HATE", "HEPU", "LEES", "PELA", "PEPU", "PHPH", "PILI", "PUTU")]

write.csv(species_t, file="figs2/Accessory_Costs/Table08_accessory_components_t-value.csv",row.names=FALSE)

```

```{r}
plot.new()
par(mfrow=c(2,3), cex=1, omi=c(.1,0.1,0.1,0.05), mai=c(.8,.8,.1,0.1))

results <- as.data.frame(matrix(vector(), 0, 8,
                dimnames=list(c(), c("variable","r2","pval","Slope","Slope_lowCI","Slope_highCI","Int","Slope_test_p"))),
                stringsAsFactors=F)

for (i in c("prop_propagule","prop_success","prop_dispersalcosts","prop_prepollencosts","prop_seedcosts")) {
yvar <- i
xvar <- "seed_size"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)
data2 <- subset(SummarySpp[["mean"]])
fit <- sma(data[[yvar]] ~ data[[xvar]], method="OLS", log="x",slope=1)
temp <- as.data.frame(c(yvar,select(fit$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Slope_test_p)))
colnames(temp)[1]<- "variable"
results <- bind_rows(results, temp)
plot(data[[yvar]] ~ data[[xvar]],log="x",col=col.spp(data$species),pch=16,xlim=c(.01,500),ylim=c(.001,1.1))
points(data2[[yvar]] ~ data2[[xvar]],col=col.spp(data2$species),pch=16,cex=2)
}

write.csv(results, file="figs2/Accessory_Costs/Table_prop_tissues_inv_by_seed_size.csv",row.names=FALSE) 
```

#looking at how the slopes of the relationship between accessory investment and propagule investment shift across species 
##slopes only statistically different from 1 in 2 species
##intercepts quite variable - means that the "baseline" accessory investment, before you get any propagules varies
```{r}
par(mfrow=c(1,1), cex=1, omi=c(.2,.2,.1,.1), mai=c(1,1,.1,.1)) 
plot(propagule_inv~accessory_inv,data=subset(SummaryInd,accessory_inv>0&propagule_inv>0),col=col.spp(species),pch=16,log="xy",xlab="total accessory investment (mg)",ylab="total propagule investment (mg)")
slope_check <- sma(log10(propagule_inv)~ log10(accessory_inv),data=subset(SummaryInd,propagule_inv>0))
summary(slope_check)

data <- select(SummaryInd,species,propagule_inv,accessory_inv)
data <- subset(data,propagule_inv>0)
data <- split(data, data$species)

for(i in 1:14) {
slope_check <- sma(log10(propagule_inv)~ log10(accessory_inv),data[[i]],slope.test=1)
summary(slope_check)
}

xvar <- "accessory_inv"
yvar <- "propagule_inv"

data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab=xvar,ylab=yvar,lty=results$linetype)
```

```{r}
win.metafile("figs2/Accessory_Costs/Figure_success_vs_failure.wmf", height=6, width=8)
plot

par(mfrow=c(1,2), cex=1, omi=c(.8,.1,.1,.1), mai=c(.1,.8,.1,.1))

xvar <- "repro_inv"
yvar <- "prop_propagule"

data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,lty=results$linetype,ylim=c(.001,1),xlab="",ylab="prop repro energy to propagules")
abline(-0.30103,0,lwd=2)
abline(-1,0,lwd=2)

xvar <- "repro_inv"
yvar <- "prop_success"

data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,lty=results$linetype,ylim=c(.001,1),xlab="",ylab="prop repro energy to success")
abline(-0.30103,0,lwd=2)
abline(-1,0,lwd=2)
mtext(text="reproductive investment (mg)",side=1,line=2,outer=TRUE)

dev.off()
```

```{r}
ternary <- select(SummarySpp[["mean"]],species,seed_size,prepollencosts,dispersalcosts)
ternary$seedcosts = ternary$seed_size + ternary$prepollencosts + ternary$dispersalcosts
ternary$per_seed <-100*ternary$seed_size/ternary$seedcosts
ternary$per_poll <-100*ternary$prepollencosts/ternary$seedcosts
ternary$per_disp <-100*ternary$dispersalcosts/ternary$seedcosts
mod_poll <- sma(per_poll~seedcosts,ternary,slope.test=0,method="OLS",log="x")
mod_seed <- sma(per_seed~seedcosts,ternary,slope.test=0,method="OLS",log="x")
mod_disp <- sma(per_disp~seedcosts,ternary,slope.test=0,method="OLS",log="x")

costs_per_seed <- as.data.frame("costs for seed (%)")
costs_per_seed <- cbind(costs_per_seed,select(mod_seed$groupsummary,r2,pval,Int,Slope,Slope_lowCI,Slope_highCI,Slope_test,Slope_test_p))
names(costs_per_seed) <- c("investment category (mg)","r squared","p value","intercept","slope","slope_lowCI","slope_highCI","slope test","slope test p")

costs_per_seed2 <- as.data.frame("costs for pre-pollination (%)")
costs_per_seed2 <- cbind(costs_per_seed2,select(mod_poll$groupsummary,r2,pval,Int,Slope,Slope_lowCI,Slope_highCI,Slope_test,Slope_test_p))
names(costs_per_seed2) <- c("investment category (mg)","r squared","p value","intercept","slope","slope_lowCI","slope_highCI","slope test","slope test p")
costs_per_seed <- rbind(costs_per_seed,costs_per_seed2)

costs_per_seed2 <- as.data.frame("costs for dispersal (%)")
costs_per_seed2 <- cbind(costs_per_seed2,select(mod_disp$groupsummary,r2,pval,Int,Slope,Slope_lowCI,Slope_highCI,Slope_test,Slope_test_p))
names(costs_per_seed2) <- c("investment category (mg)","r squared","p value","intercept","slope","slope_lowCI","slope_highCI","slope test","slope test p")
costs_per_seed <- rbind(costs_per_seed,costs_per_seed2)

write.csv(costs_per_seed, file="figs2/Accessory_Costs/Table08_costs_per_seed.csv",row.names=FALSE)

win.metafile("figs2/Accessory_Costs/seedcost_components.wmf", height=6, width=6)
plot

plot(per_seed~seedcosts,ternary,col="black",log="x",pch=21,cex=2,ylim=c(0,75),ylab="prop of seedcosts")
points(per_poll~seedcosts,ternary,col="red",pch=16,cex=2)
points(per_disp~seedcosts,ternary,col="blue",pch=16,cex=2)

abline(mod_poll$groupsummary["Int"][[1]],mod_poll$groupsummary["Slope"][[1]],col="red",lwd=2)
abline(mod_seed$groupsummary["Int"][[1]],mod_seed$groupsummary["Slope"][[1]],col="black",lty=2,lwd=2)
abline(mod_disp$groupsummary["Int"][[1]],mod_disp$groupsummary["Slope"][[1]],col="blue",lwd=2)
legend("topleft",legend=c("seed","prepollen","pack&disp"),col=c("black","red","blue"),pch=16, cex=1.1,bty="n")
dev.off()
```

```{r}
win.metafile("figs2/Accessory_Costs/Ternary.wmf", height=5, width=5)
plot

ggtern(data=ternary,aes(y=per_seed,x=per_poll,z=per_disp)) +
  labs(y = "seed weight (%)",x = "prepoll (%)",z = "pack&disp (%)",title = "Seed cost components",size = .5) +
  geom_point(size=sqrt(ternary$seedcosts/2),shape=20,color=col.spp()) +
   theme_bw() + 
  geom_text(label=ternary$species, color="black", hjust=-.3, vjust=sqrt(ternary$seedcosts/500),size=2)

dev.off()
```