### 2016-05-17 scripts/Biology_talk.rmd
Move scripts into subfolder, remove from remake
Remove old plotting functions

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=8, fig.height=8}
win.metafile("figs2/RA_all_spp1.wmf", height=5, width=5)
plot
par(mfrow=c(1,1), cex=1, omi=c(.7,.7,.1,.1), mai=c(.1,.1,.1,0.1)) 
mod <- lm(log(growth_inv)~log(repro_inv),data=subset(SummaryInd,repro_inv>0))
plot((1+repro_inv)~growth_inv, SummaryInd, log="xy", col=col.spp(species),pch=16,xlim=c(1,1000000),ylim=c(1,1000000),xlab="",ylab="")
words.top.left.logxy(mod)
legend("bottomright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5",cex=.7)
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75",cex=.7)
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9",cex=.7)
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95",cex=.7)
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1",cex=.7)
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01",cex=.7)
mtext("reproductive investment (mg)", 2, outer=TRUE,cex=1,line=2)
mtext("growth investment (mg)", 1, outer=TRUE,cex=1,line=1.7)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=8, fig.height=8}
win.metafile("figs2/RA_all_spp2.wmf", height=5, width=5)
plot
par(mfrow=c(1,1), cex=1, omi=c(.7,.7,.1,.1), mai=c(.1,.1,.1,0.1))  
plot((1+repro_inv)~growth_inv, SummaryInd, log="xy", col=col.age(age),pch=16,xlim=c(1,1000000),ylim=c(1,1000000))
legend("bottomright",legend=labels.age(),col=col.age(),pch=16, cex=.8,bty="n")
abline(0,1)
text(2.1,3,srt=45,"RA = 0.5",cex=.7)
abline(log10(3),1)
text(2,9,srt=45,"RA = 0.75",cex=.7)
abline(log10(9),1)
text(2,25,srt=45,"RA = 0.9",cex=.7)
abline(log10(19),1)
text(2,60,srt=45,"RA = 0.95",cex=.7)
abline(log10(.111),1)
text(19,3,srt=45,"RA = 0.1",cex=.7)
abline(log10(.010101),1)
text(170,2.4,srt=45,"RA = 0.01",cex=.7)
mtext("reproductive investment (mg)", 2, outer=TRUE,cex=1,line=2)
mtext("growth investment (mg)", 1, outer=TRUE,cex=1,line=1.7)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_reproinv_by_age.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
xvar <- "age"
yvar <- "repro_inv"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(repro_inv~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("reproductive investment (mg)", 2, outer=TRUE,cex=1,line=2)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_growthinv_by_age.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
xvar <- "age"
yvar <- "growth_inv"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(growth_inv~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("vegetative investment (mg)", 2, outer=TRUE,cex=1,line=2.2)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_leafinv_by_age.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
xvar <- "age"
yvar <- "growth_leaf"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(growth_leaf~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("leaf investment (mg)", 2, outer=TRUE,cex=1,line=2.2)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_H_by_age.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
xvar <- "age"
yvar <- "height"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(height~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("height (mm)", 2, outer=TRUE,cex=1,line=2)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_weight_by_age.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
yvar <- "total_weight"
xvar <- "age"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
}
mtext("plant weight (mg)", 2, outer=TRUE,cex=1,line=2.2)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_plot1.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
yvar <- "RA"
xvar <- "age"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",xlab="",ylab="",xlim=c(1,33),ylim=c(0,1),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
}
mtext("RA", 2, outer=TRUE,cex=1,line=1.8)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_plot2.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.7,.01,.01), mai=c(.4,.15,.3,0.1)) 
yvar <- "total_inv"
xvar <- "leaf_area_midyear"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  arrows(data[[spp]]$leaf_area_midyear,data[[spp]]$growth_inv,data[[spp]]$leaf_area_midyear,data[[spp]]$total_inv,length = 0, angle = 30, code = 2, col = "black",lwd=1)
}
mtext("repro + growth investment (mg)", 2, outer=TRUE,cex=1,line=2.5)
mtext("leaf area (mm^2)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_plot3.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)) 
yvar <- "RA_leaf_area"
xvar <- "age"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",xlab="",ylab="",xlim=c(1,33),ylim=c(0,1.2),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
}
mtext("RA (based on leaf weight only)", 2, outer=TRUE,cex=1,line=1.8)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_leafweight_x_age.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.7,.01,.01), mai=c(.4,.15,.3,0.1)) 
yvar <- "leaf_weight"
xvar <- "age"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age))
}
mtext("leaf weight (mg)", 2, outer=TRUE,cex=1,line=2.5)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_plot5.wmf", height=7, width=4)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.7,.01,.01), mai=c(.4,.15,.3,0.1))
yvar <- "stem_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="BOLE"|species=="GRBU"|species=="EPMI"|species=="HATE")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar2,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age))
}
mtext("stem weight (mg)", 2, outer=TRUE,cex=1,line=2.5)
mtext("age (years)", 1, outer=TRUE,cex=1,line=0.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_RA_canopy.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="BAER"|species=="HATE"|species=="PELA"|species=="PEPU")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar2,log="x",xlab="",ylab="",xlim=c(1,33),ylim=c(0,1.2),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(RA_leaf_area~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("canopy species", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_RA_early.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="BOLE"|species=="PILI"|species=="HEPU"|species=="COER")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar2,log="x",xlab="",ylab="",xlim=c(1,33),ylim=c(0,2),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(RA_leaf_area~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("early reproducers", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_RA_mid.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="GRBU"|species=="GRSP"|species=="PUTU"|species=="PHPH")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar2,log="x",xlab="",ylab="",xlim=c(1,33),ylim=c(0,1.2),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(RA_leaf_area~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("mid-age reproducers", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_RA_mid-late.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="EPMI"|species=="LEES")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar2,log="x",xlab="",ylab="",xlim=c(1,33),ylim=c(0,1.2),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(RA_leaf_area~age,data2[[spp]],pch=16,cex=0.8)
}
mtext("epacrids", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
?win.metafile("figs2/RA_talk_LA_canopy.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="BAER"|species=="HATE"|species=="PELA"|species=="PEPU")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], xvar, xvar2,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,cex.axis=0.7,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(leaf_area_midyear~age,data2[[spp]],pch=16)
}
mtext("canopy species", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_LA_early.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="BOLE"|species=="PILI"|species=="HEPU"|species=="COER")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], xvar, xvar2,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(leaf_area_midyear~age,data2[[spp]],pch=16)
}
mtext("early reproducers", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_LA_mid.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="GRBU"|species=="GRSP"|species=="PUTU"|species=="PHPH")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], xvar, xvar2,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(leaf_area_midyear~age,data2[[spp]],pch=16)
}
mtext("mid-age reproducers", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_LA_mid-late.wmf", height=7, width=3)
plot
par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd, species=="EPMI"|species=="LEES")
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], xvar, xvar2,log="xy",xlab="",ylab="",xlim=c(1,33),cex=1,
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  points(leaf_area_midyear~age,data2[[spp]],pch=16)
}
mtext("epacrids", 3, outer=TRUE,cex=1.2,line=.5)
dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
#win.metafile("figs2/RA_talk_LA_decline2.wmf", height=7, width=3)
#plot
par(mfcol=c(4,4), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1)) 
SummaryInd$ratio_leaf_growth = log10(SummaryInd$leaf_area) - log10(SummaryInd$leaf_area_0)

yvar <- "ratio_leaf_growth"
xvar <- "age"

data <- subset(SummaryInd,individual!="GRBU_903"&individual!="PELA_905"&individual!="COER_804"&individual!="PEPU_905"&individual!="GRSP_907"&individual!="LEES_907")
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,xlab="",ylab="",cex=1,log="x",xlim=c(1,33),ylim=c(-0.35,1.75),
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  abline(0,0)
  }
#text("epacrids", 3, outer=TRUE,cex=1.2,line=.5)
#dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
#win.metafile("figs2/RA_talk_LA_decline2.wmf", height=7, width=3)
#plot

  for(v in c("growth_leaf_min")) {
    i <- (SummaryInd[[v]]) > 0
    SummaryInd[[v]][i] <- 0
    }

SummaryInd$growth_leaf_nonzero <- SummaryInd$growth_leaf - SummaryInd$growth_leaf_min +1

par(mfcol=c(4,4), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1))

yvar <- "growth_leaf_nonzero"
xvar <- "repro_inv"

data <- subset(SummaryInd,individual!="GRBU_903"&individual!="PELA_905"&individual!="COER_804"&individual!="PEPU_905"&individual!="GRSP_907"&individual!="LEES_907"&repro_inv>0)
data <- split(data, data$species)

lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,xlab="",ylab="",cex=1,log="xy",
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
    }
#text("epacrids", 3, outer=TRUE,cex=1.2,line=.5)
#dev.off()
```

  for(v in c("growth_leaf_min")) {
    i <- (SummaryInd[[v]]) > 0
    SummaryInd[[v]][i] <- 0
    }

SummaryInd$growth_leaf_min <- 1 + abs(SummaryInd$growth_leaf_min)
SummaryInd$growth_leaf_nonzero <- SummaryInd$growth_leaf + SummaryInd$growth_leaf_min

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
#win.metafile("figs2/RA_talk_LA_decline2.wmf", height=7, width=3)
#plot

par(mfcol=c(4,4), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1))

yvar <- "growth_leaf_nonzero"
xvar <- "age"

data <- subset(SummaryInd,individual!="GRBU_903"&individual!="PELA_905"&individual!="COER_804"&individual!="PEPU_905"&individual!="GRSP_907"&individual!="LEES_907")
data <- split(data, data$species)

lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,xlab="",ylab="",cex=1,log="xy", xlim=c(1,33),yaxt="n", 
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
  abline(log10(data[[spp]]$growth_leaf_min),0,lwd=3)
    }
#text("epacrids", 3, outer=TRUE,cex=1.2,line=.5)
#dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
#win.metafile("figs2/RA_talk_LA_decline2.wmf", height=7, width=3)
#plot

  for(v in c("growth_leaf_min")) {
    i <- (SummaryInd[[v]]) > 0
    SummaryInd[[v]][i] <- 0
    }

SummaryInd$growth_leaf_nonzero <- SummaryInd$growth_leaf - SummaryInd$growth_leaf_min

growth_by_site <- SummaryInd %>%
  group_by(species,age) %>%
  summarise_each(funs(mean, se), growth_leaf_nonzero,repro_inv)

growth_by_site$repro_inv_mean <- growth_by_site$repro_inv_mean +1
growth_by_site$growth_leaf_nonzero_mean <- growth_by_site$growth_leaf_nonzero_mean +1

par(mfcol=c(4,4), cex=1, omi=c(.3,.6,.3,.01), mai=c(.4,.4,.3,0.1))

yvar <- "growth_leaf_nonzero_mean"
xvar <- "repro_inv_mean"

data <- subset(growth_by_site)
data <- split(data, data$species)

lmfits <- lapply(data, function(x) lm(log10(x[[yvar]]) ~  (log10(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,xlab="",ylab="",cex=1,log="xy",
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
    }
#text("epacrids", 3, outer=TRUE,cex=1.2,line=.5)
#dev.off()
```

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
win.metafile("figs2/RA_talk_LA_tradeoff.wmf", height=7, width=4)
plot

  for(v in c("growth_leaf_min")) {
    i <- (SummaryInd[[v]]) > 0
    SummaryInd[[v]][i] <- 0
    }

SummaryInd$growth_leaf_nonzero <- SummaryInd$growth_leaf - SummaryInd$growth_leaf_min

par(mfcol=c(4,1), cex=1, omi=c(.3,.6,.01,.01), mai=c(.4,.15,.3,0.1)

yvar <- "growth_leaf_nonzero"
xvar <- "repro_inv"

data <- subset(SummaryInd,individual!="GRBU_903"&individual!="PELA_905"&individual!="COER_804"&individual!="PEPU_905"&individual!="GRSP_907"&individual!="LEES_907"&(species=="BOLE"|species=="EPMI"|species=="GRBU"|species=="HATE"))
data <- split(data, data$species)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,xlab="",ylab="",cex=1,log="xy",yaxt="n",
                    col=col.age(data[[spp]]$age),main=labels.spp.full(spp),cex.main=0.9)
    }

mtext("leaf investment (mg)", 2, outer=TRUE,cex=1,line=2)
mtext("reproductive investment (mg)", 1, outer=TRUE,cex=1,line=1.7)
dev.off()
```