### 2016-05-17 scripts/RA_combined_plots.Rmd

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=14, fig.height=7.5}
remake::make("export")

SummaryInd <- readRDS("export/SummaryInd.rds")
SummarySpp <- readRDS("export/SummarySpp.rds")
SummarySppAge <- readRDS("export/SummarySppAge.rds")

source("R/figures.R")
```

###Current best plots  
* the plot of investment by leaf area, indicates that allocation to reproduction, leads to less vegetative investment than predicted  
* plotting by age, shows that leaf investment and height growth are curtailed as plants age  
* I think there still needs to be more of a link to connect this decreased "growth" with age directly to reproduction - maybe a "new" version of the plot showing individuals with more reproduction have less growth?
```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
SummaryInd <- SummaryInd[order(factor(SummaryInd$species,levels=c(c("HATE","PELA","BAER","PEPU","LEES","EPMI","COER","PUTU","PHPH","GRBU","GRSP","HEPU","PILI","BOLE")))),]

#pdf("figs2/RA_schedule_plots_3.pdf", height=11.3, width=8)
#plot
par(mfcol=c(4,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1,1,.1,0.1)) 
yvar <- "leaf_weight"
yvar2 <- "height"
yvar3 <- "total_inv"
yvar4 <- "RA_leaf_area"
xvar <- "leaf_area_midyear"
xvar2 <- "age"
xvar3 <- "height"

data <- subset(SummaryInd)
data <- split(data, data$species)

data2 <- subset(SummarySppAge[["mean"]])
data2 <- split(data2, data2$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar3, xvar,log="xy",xlab="leaf area (midyear)",ylab="total investment (mg)",cex=1.1,
                    col=col.age(data[[spp]]$age),main=spp)
  arrows(data[[spp]]$leaf_area_midyear,data[[spp]]$growth_inv,data[[spp]]$leaf_area_midyear,data[[spp]]$total_inv,length = 0, angle = 30, code = 2, col = "black",lwd=1)
  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar2,log="x",xlab="age",ylab="RA",cex=1.1,ylim=c(0,1.4),
                    col=col.age(data[[spp]]$age))
  points(RA_leaf_area~age,data2[[spp]],pch=16)
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar2,log="xy",xlab="age",ylab="final leaf area",cex=1.1,
                    col=col.age(data[[spp]]$age))
  points(leaf_area~age,data2[[spp]],pch=16)
  plot_yvar_vs_xvar(data[[spp]], yvar2, xvar2,log="xy",xlab="age",ylab="height",cex=1.1,
                    col=col.age(data[[spp]]$age))
  points(height~age,data2[[spp]],pch=16)
  }
#dev.off()
```
SummaryInd$growth_ratio <- SummaryInd$growth_inv /SummaryInd$repro_inv

#How to depict a trade-off?
- only GRBU has an obvious trade-off, where absolute growth declines
```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=20, fig.height=20}
  par(mfcol=c(3,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.5,.7,.6,0.2)) 
xvar <- "growth_leaf"
xvar2 <- "growth_inv"
xvar3 <- "age"
yvar <- "repro_inv"
yvar2 <- "growth_ratio"

data <- subset(SummaryInd,repro_inv>0)
data <- split(data, data$species)

lmfits <- lapply(data, function(x) lm(log10(x[[yvar]]) ~  (log10(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1.3, main=spp,col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar2,log="xy",xlab="",ylab="",cex=1.3, main=spp,col=col.age(data[[spp]]$age))
    plot_yvar_vs_xvar(data[[spp]], yvar2, xvar3,log="xy",xlab="",ylab="",cex=1.3, main=spp,col=col.age(data[[spp]]$age))
 }

mtext("repro inv", 2, outer=TRUE,cex=1.2)
mtext("growth inv", 1, outer=TRUE,cex=1.2)
legend("topleft",legend=labels.age(),col=col.age(),pch=16, cex=1.1,bty="n")
```

###Trends by age
```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
par(mfcol=c(2,2), cex=1, omi=c(.1,.3,.3,.1), mai=c(1.1,1,.1,0.1)) 
xvar <- "age"

yvar1 <- "height"
yvar2 <- "diameter"
yvar3 <- "leaf_area"
yvar4 <- "total_weight_0"

data <- subset(SummaryInd)
data <- split(data, data$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar1, xvar,log="xy",ylab="height",xlab="age",cex=1.3,
                    col=col.age(data[[spp]]$age),main=spp)
  plot_yvar_vs_xvar(data[[spp]], yvar2, xvar,log="xy",ylab="diameter",xlab="age",cex=1.3,
                    col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar3, xvar,log="xy",ylab="leaf area",xlab="age",cex=1.3,
                    col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar,log="xy",ylab="weight",xlab="age",cex=1.3,
                    col=col.age(data[[spp]]$age))
  }
```

###Figures from lab meeting on Sept 3, 2015
```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.width=18, fig.height=15}
par(mfcol=c(4,3), cex=1, omi=c(.8,.3,.3,.1), mai=c(.5,1,.1,0.1)) 
SummaryInd$repro_inv_log <- SummaryInd$repro_inv + 10
yvar <- "growth_height_log"
yvar2 <- "growth_leaf_area_log"
yvar3 <- "shoot_leaf_area"
yvar4 <- "RA"
xvar <- "repro_inv_log"
xvar2 <- "total_weight_0"
xvar3 <- "prop_maxH"

data <- subset(SummaryInd)
data <- split(data, data$species)

for(spp in names(data)) {

  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",xlab="",ylab="height increase (log-log)",cex=1.3, col=col.age(data[[spp]]$age))
  legend("topright",legend=labels.age(),col=col.age(),pch=16, cex=1,bty="n")
  plot_yvar_vs_xvar(data[[spp]], yvar2, xvar,log="x",xlab="",ylab="total leaf area increase (log-log)",cex=1.3,col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar3, xvar,log="xy",xlab="",ylab="total shoot leaf area",cex=1.3, col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar,log="x",xlab="reproductive investment",ylab="RA",cex=1.3,col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar2,log="x",xlab=spp,ylab="",cex=1.3, col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar2, xvar2,log="x",xlab="",ylab="",cex=1.3, col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar3, xvar2,log="xy",xlab="",ylab="",cex=1.3, col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar2,log="x",xlab="initial plant weight",ylab="",cex=1.3,col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar3,log="",xlab="",ylab="",cex=1.3, col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar2, xvar3,log="",xlab="",ylab="",cex=1.3,col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar3, xvar3,log="y",xlab="",ylab="",cex=1.3,col=col.age(data[[spp]]$age))
  plot_yvar_vs_xvar(data[[spp]], yvar4, xvar3,log="",xlab="prop maxH",ylab="",cex=1.3,col=col.age(data[[spp]]$age))
  mtext("Reproductive investment (mg)                                                                            initial plant weight                                                                                    prop maxH", 1, line=1,outer=TRUE,cex=1.2)
  mtext(spp, 3, outer=TRUE,cex=1.2)   
}
```