### 2016-05-17 scripts/investment_in_parts.Rmd

InvestInd <- 
  InvestmentByPart %>%
  group_by(individual) %>%
  summarise_each(funs(sum),weight)
InvestInd$species <- substr(InvestInd$species,1,nchar(InvestInd$species)-4)

```{r}
InvestIndPart <- 
  InvestmentByPart %>%
  group_by(individual,part) %>%
  summarise_each(funs(sum),weight)
InvestIndPart$species <- InvestIndPart$individual
InvestIndPart$species <- substr(InvestIndPart$species,1,nchar(InvestIndPart$species)-4)
InvestIndPart <- merge(InvestIndPart,accessoryList,by=c("species","part"),all.x=TRUE)
```

```{r}
InvestIndAcc <-
  InvestIndPart %>%
  group_by(individual,accessory_part) %>%
  summarise_each(funs(sum),weight)

  for(v in c("weight")) {
    i <- is.na(InvestIndAcc[[v]])
    InvestIndAcc[[v]][i] <- 0
   }

InvestIndAcc1 <- tidyr::spread(InvestIndAcc,accessory_part,weight,fill=0,drop=FALSE)
View(InvestIndAcc)
```

InvestPetal <- subset(InvestmentByPart,part=="flower_petals")
InvestPetal <- 
  InvestPetal %>%
  group_by(individual) %>%
  summarise_each(funs(sum),weight)

names(InvestPetal) <- c("individual","weight_petal")

InvestInd <- dplyr::left_join(InvestInd,InvestPetal, by = "individual")
InvestInd$prop_petal <- InvestInd$weight_petal / InvestInd$weight
InvestInd$species <-InvestInd$individual
InvestInd$species <- substr(InvestInd$species,1,nchar(InvestInd$species)-4)

plot(prop_petal~weight,InvestInd,pch=16,log="x",col=col.spp(species))
mod <- lm(log(weight)~prop_petal,InvestInd)
summary(mod)

InvestIndPart <- tidyr::spread(InvestIndPart,part,weight)