### 2016-09-21 ms/Accessory/ms-supp.Rnw
Framework for supplementary materials document

\documentclass[10pt,twoside]{article}

\usepackage{suppmat}
\usepackage{listings}
\usepackage[T1]{fontenc} % Better underscores
\usepackage{longtable}

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}

\title{Accessories dominate the costs of producing seed}
\runninghead{C\lowercase{osts of producing seed}}

\author{E. H. Wenk\textsuperscript{1,\textasteriskcentered}, K. Abramowicz\textsuperscript{2}, M. Westoby\textsuperscript{1}, D.S. Falster\textsuperscript{1} }
\affiliation{\textsuperscript{1} Biological Sciences, Macquarie University NSW 2109, Australia \\
\textsuperscript{2} Department of Mathematics and Mathematical Statistics, Ume{\aa} University, 90187 Ume{\aa}, Sweden \\
\textsuperscript{\textasteriskcentered} Correspondence author. E-mail: \email{ehwenk@gmail.com}
}

\usepackage[authoryear,sectionbib,sort]{natbib}
\bibliographystyle{amnat2}

\pdfminorversion=4 % tell pdflatex to generate PDF in version 1.4
% Needed for problem-free conversion on manuscript central (arrgh)
% http://timotheepoisot.fr/2013/06/24/manuscript-central-pdf-fix/

\titleprefix{Supporting Materials for:}

\date{}

<<preload, echo=FALSE>>=
library(xtable)
library(knitr)
opts_chunk$set(echo=FALSE, fig.pos = "!h", fig.align="center", comment = NA)

flowering_parts <- read.csv("Sup_Table_flowering_parts.csv",
  stringsAsFactors=FALSE, check.names=FALSE)

print_table_species <- function(species, data) {
  data <- data[data$species == species, -c(1)]
  print(xtable(data,
            hline.after=c(1),
            align='p{0.1cm}p{4cm}p{6cm}p{3cm}p{1cm}p{1cm}',
            label=sprintf("tab:parts_%s", species),
            caption=sprintf("Table of parts for %s.", species)),
      sanitize.text.function=I,
      , table.placement = "H"
      , caption.placement = "top"
      , include.rownames = FALSE
      , include.colnames = TRUE
      , size = "small"
      , tabular.environment = 'longtable'
      , floating = FALSE)
}

@

\begin{document}

\maketitle

\tableofcontents

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{secnumdepth}{0}

\clearpage

\section{A -- Methods for estimating Reproductive allocation}

All computations implemented in \textsc{r} \citep{R-2015}.

\subsection{Banksia ericifolia}

Details on parts for \emph{Banksia ericifolia} found in Table \ref{tab:parts_Banksia ericifolia}. See Fig \ref{fig:map Banksia ericifolia} for a plot.

<<echo=FALSE,results='asis'>>=
print_table_species("Banksia ericifolia", flowering_parts)
@

\begin{figure}
\centering
<<echo=FALSE,results='asis'>>=
plot(BAER_GraphMaps$graph)
@
\caption{Map of parts for \emph{Banksia ericifolia}}
\label{fig:map Banksia ericifolia}
\end{figure}

\clearpage

\subsection{Conospermum ericifolium}

Details on parts for \emph{Conospermum ericifolium} found in Table \ref{tab:parts_Conospermum ericifolium}.  See Fig \ref{fig:map Conospermum ericifolium} for a plot.

<<echo=FALSE,results='asis'>>=
print_table_species("Conospermum ericifolium", flowering_parts)
@

\begin{figure}
\centering
<<echo=FALSE,results='asis'>>=
plot(BAER_GraphMaps$graph)
@
\caption{Map of parts for \emph{Conospermum ericifolium}}
\label{fig:map Conospermum ericifolium}
\end{figure}

\clearpage

\clearpage

\addcontentsline{toc}{section}{References}
\bibliography{references}

\end{document}