### 2016-11-24 scripts/Accessory_figs_20160905.Rmd
work on accessory costs

plot(seedset~scaled_provisioning_costs,SummarySpp[["mean"]],pch=16,log="xy",col=col.spp(species),yaxt="n",xaxt="n",ylab="",xlab="",cex=1)

mod_ovule <- sma(choosiness~prop_pollen_attract_vs_success,data=SummarySpp[["mean"]],log="xy",method="SMA")

axis(2, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=0.8,las=2)
axis(1, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=0.8)
mtext("costs to make a successful unit",1,line=1.8,cex=0.7)
mtext("count, scaled to leaf area",2,line=2.3,cex=0.7)

mod_x <- sma(accessory_inv~embryo_endo_inv,data=SummarySpp[["mean"]],log="xy",method="SMA")
mod_x <- sma(accessory_costs~embryo_endo_costs,data=SummarySpp[["mean"]],log="xy",method="SMA")
plot(zygote_set~embryo_endo_costs,SummaryInd,pch=16,log="x",col=col.spp(species),cex=1)

#Data 0. means, counts to include in manuscript
```{r}
#number of plants in study
count(subset(SummaryInd,age>1))

#number of reproducing individuals in study
count(subset(SummaryInd,repro_inv>0))

#number of individuals in study that produce seed
count(subset(SummaryInd,seed_count>0))

#average proportion energy going to failed tissues (among reproducing individuals)
1- mean(subset(SummaryInd$prop_success,SummaryInd$seed_count>0))
mean(subset(SummaryInd$seedset,SummaryInd$seed_count>0))

#average proportion of energy going to accessory costs (based on propagules, not seeds) (among individuals that produce seed)
mean(subset(SummaryInd$prop_accessory_vs_embryo_endo,SummaryInd$seed_count>0))
mean(subset(SummaryInd$prop_accessory_vs_propagule,SummaryInd$seed_count>0))
SummarySpp[["mean"]]$prop_accessory_vs_embryo_endo
SummarySpp[["mean"]]$species

#individual with lowest accessory costs
check <- subset(SummaryInd,prop_embryo_endo_vs_all_repro>.1)
check <- dplyr::select(check,individual,prop_embryo_endo_vs_all_repro)
max(check$prop_embryo_endo_vs_all_repro)

#r2 for relationship between propagule inv and repro inv across all individuals
mod <- sma(propagule_inv~repro_inv,subset(SummaryInd,SummaryInd$seedset>0),log="xy",method="SMA")

#individual with minimum accessory costs
min(SummaryInd$prop_accessory_vs_embryo_endo)

max(SummaryInd$prop_embryo_endo_vs_all_repro)

#relationship between plant weight and number of ovules initiated - bigger plants produce proportionally fewer buds
mod <- sma(ovule_count~total_weight,SummaryInd,method="SMA",log="xy",slope=1)
summary(mod)

#RE per seed unrelated to seedsize
mod <- sma(repro_costs ~ embryo_endo_costs, SummarySpp[["mean"]],method="SMA", log="xy",slope.test=1)
summary(mod)
plot(repro_costs ~ embryo_endo_costs, SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

#energy invested in accessory tissues unrelated to seedsize
mod <- sma(accessory_costs_using_ee ~ embryo_endo_costs, data=subset(SummarySpp[["mean"]]),method="SMA", log="xy",slope.test=1)
summary(mod)
plot(accessory_costs_using_ee ~ embryo_endo_costs, SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

#total seed costs correlated with seedsize
mod <- sma(success_costs ~ embryo_endo_costs, SummarySpp[["mean"]],method="SMA", log="xy")
summary(mod)
plot(success_costs ~ embryo_endo_costs, SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

#DISCUSSION: lower seedset = higher proportional investment in discarded tissues
plot(seedset~prop_discarded_vs_all_repro,data=subset(SummarySpp[["mean"]]),log="",pch=16,col=col.spp(species))
mod <- sma(seedset~prop_discarded_vs_all_repro,data=subset(SummarySpp[["mean"]],species!="BAER"),log="",method="SMA")
summary(mod)

#RESULTS - seed set vs seed size
mod <- sma(seedset ~ embryo_endo_costs, data=subset(SummarySpp[["mean"]]),method="SMA", log="xy",slope.test=1)
summary(mod)
plot(seedset ~ embryo_endo_costs, SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

#seed size by propagule investment - all species
mod <- sma(propagule_inv ~ embryo_endo_costs, SummarySpp[["mean"]],method="SMA", log="xy",slope.test=1)
summary(mod)
plot(propagule_inv ~ embryo_endo_costs, SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

#seed size by propagule investment - without EPMI
mod <- sma(propagule_inv ~ embryo_endo_costs, data=subset(SummarySpp[["mean"]],species!="EPMI"),method="SMA", log="xy",slope.test=1)
summary(mod)
plot(propagule_inv ~ embryo_endo_costs, data=subset(SummarySpp[["mean"]],species!="EPMI"),log="xy",pch=16,col=col.spp(species))

mod <- sma(repro_inv ~ embryo_endo_inv, SummaryInd,method="SMA", log="xy",slope.test=1)
summary(mod)
plot(repro_inv ~ embryo_endo_inv, SummaryInd,log="xy",pch=16,col=col.spp(species))

mod <- sma(repro_inv ~ flower_inv, SummaryInd,method="SMA", log="xy",slope.test=1)
summary(mod)
plot(repro_inv ~ flower_inv, SummaryInd,log="xy",pch=16,col=col.spp(species))

#DISCUSSION SECTION ON ESTIMATING REPRODUCTIVE EFFORT: seed size by reproductive investment - all species
mod <- sma(repro_inv ~ embryo_endo_costs, SummarySpp[["mean"]],method="SMA", log="xy",slope.test=1)
summary(mod)
plot(repro_inv ~ embryo_endo_costs, SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

#DISCUSSION SECTION ON ESTIMATING REPRODUCTIVE EFFORT: seed size by reproductive investment - without EPMI
mod <- sma(repro_inv ~ embryo_endo_costs, data=subset(SummarySpp[["mean"]],species!="EPMI"),method="SMA", log="xy",slope.test=1)
summary(mod)
plot(repro_inv ~ embryo_endo_costs, data=subset(SummarySpp[["mean"]],species!="EPMI"),log="xy",pch=16,col=col.spp(species))

#DISCUSSION SECTION ON ESTIMATING REPRODUCTIVE EFFORT: seed size by total weight - all species
mod <- sma(total_weight_0 ~ embryo_endo_costs, SummarySpp[["mean"]],method="SMA", log="xy",slope.test=1)
summary(mod)
plot(total_weight_0 ~ embryo_endo_costs, SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

yvar <- "ovule_count"
xvar <- "repro_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,xlab=xvar,ylab=yvar,pch=16, col=col.spp(fit$groupsummary$group),lty=results$linetype,ylim=c(1,300000),xlim=c(1,30000))
results

yvar <- "ovule_count"
xvar <- "leaf_area_0"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy",slope.test=1)
results <- lines(fit$groupsummary)
plot(fit,xlab=xvar,ylab=yvar,pch=16, col=col.spp(fit$groupsummary$group),lty=results$linetype,ylim=c(1,300000),xlim=c(1,30000))
results

yvar <- "propagule_inv"
xvar <- "leaf_area_0"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy",slope.test=1)
results <- lines(fit$groupsummary)
plot(fit,xlab=xvar,ylab=yvar,pch=16, col=col.spp(fit$groupsummary$group),lty=results$linetype,ylim=c(1,300000),xlim=c(1,30000))
results

xvar <- "propagule_inv"
yvar <- "repro_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy",slope.test=1)
results <- lines(fit$groupsummary)
plot(fit,xlab=xvar,ylab=yvar,pch=16, col=col.spp(fit$groupsummary$group),lty=results$linetype,ylim=c(1,300000),xlim=c(1,30000))
results
```

#Table 1 showing accessory cost values for the species
```{r}
acc_costs_spp <- dplyr::select(SummarySpp[["mean"]],species,embryo_endo_costs, seedset, repro_costs, prop_prepollen_discarded_vs_all_repro,
                               prop_postpollen_discarded_vs_all_repro, prop_pollen_attract_vs_all_repro, prop_pack_disp_vs_all_repro, 
                               prop_embryo_endo_vs_all_repro, success_costs, prop_pollen_attract_vs_all_repro, prop_pollen_attract_vs_success, 
                               prop_provisioning_vs_success, embryo_endo_costs, packaging_dispersal_costs, pollen_attract_costs, repro_inv,seed_count)
acc_costs_spp$species <- labels.spp.full()
acc_costs_spp_format <- acc_costs_spp
for (i in c("embryo_endo_costs","seedset","repro_costs","prop_prepollen_discarded_vs_all_repro","prop_postpollen_discarded_vs_all_repro",
            "prop_pollen_attract_vs_all_repro","prop_pack_disp_vs_all_repro","prop_embryo_endo_vs_all_repro","success_costs",
            "prop_pollen_attract_vs_success", "prop_provisioning_vs_success","packaging_dispersal_costs","pollen_attract_costs", "repro_inv")) {
  acc_costs_spp_format[[i]] <- round(acc_costs_spp_format[[i]],digits=3)
}

for (i in c("seed_count")) {
  acc_costs_spp_format[[i]] <- round(acc_costs_spp_format[[i]],digits=1)
}

for (i in c("prop_prepollen_discarded_vs_all_repro","prop_postpollen_discarded_vs_all_repro","prop_pack_disp_vs_all_repro",
            "prop_embryo_endo_vs_all_repro","prop_pollen_attract_vs_all_repro","prop_pollen_attract_vs_success", "prop_provisioning_vs_success")) {
  acc_costs_spp_format[[i]] <- round(100*acc_costs_spp_format[[i]],digits=2)
}

acc_costs_spp_format <- acc_costs_spp_format[order(acc_costs_spp_format$embryo_endo_costs),]

write.csv(acc_costs_spp_format, file="ms/Accessory/Table01_AccCosts_by_Species.csv",row.names=FALSE)
```

#Table 1. color dots
```{r}
win.metafile("ms/Accessory/Table01_dots.wmf", height=3, width=.2)
plot

par(mfrow=c(1,1), cex=1, omi=c(.01,.01,.01,.01), mai=c(.01,.01,.01,.01)) 
dots <- dplyr::select(SummarySpp[["mean"]],species,embryo_endo_costs)
dots <- dots[order(dots$embryo_endo_costs),] 
dots$x <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1)
dots$y <- c(28,26,24,22,20,18,16,14,12,10,8,6,4,2)

plot(y~x,dots,cex=1.5,pch=16,xlab="",ylab="",xaxt="n",yaxt="n",col=col.spp(dots$species),bty="n",xlim=c(0.8,1.2),ylim=c(1.8,28.2))

dev.off()
```
summary(slopes)

#Table 2. Slopes for various embryo size by acc cost regressions
```{r}
par(mfrow=c(2,3), cex=1, omi=c(.6,.1,.05,.05), mai=c(.1,.8,.05,.05)) 
options(scipen = 5)

library(smatr)

plot(repro_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), xlim=c(.02,10000), ylab="total per seed accessory costs (mg/seed)",xlab="",cex=1.5)
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$repro_costs,SummarySpp[["se"]]$accessory_costs)
slopes <- sma(repro_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1,method="SMA", log="xy")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
summary(slopes)

plot(prop_pollen_attract_vs_all_repro_all_repro~ seedset,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy",cex=1.5)

acc_by_species <- dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p)

plot(success_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="seed costs (mg/seed)",xlab="",cex=1.5)
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$success_costs,SummarySpp[["se"]]$success_costs)
slopes <- sma(success_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1,method="SMA", log="xy")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
acc_by_species <- dplyr::bind_rows(acc_by_species, dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(pollen_attract_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="prepollen success costs (mg/seed)",xlab="",cex=1.5)
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$pollen_attract_costs,SummarySpp[["se"]]$pollen_attract_costs)
slopes <- sma(pollen_attract_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1,method="SMA", log="xy")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
acc_by_species <- dplyr::bind_rows(acc_by_species, dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(provisioning_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="dispersal & packaging accessory costs (mg/seed)",xlab="",cex=1.5)
slopes <- sma(provisioning_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1, log="xy",method="SMA")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$provisioning_costs,SummarySpp[["se"]]$provisioning_costs)
acc_by_species <- dplyr::bind_rows(acc_by_species, dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))
mtext(text="seed size (mg)",side=1,line=1.5,outer=TRUE)

plot(discarded_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="all discarded costs (mg/seed)",xlab="",cex=1.5)
slopes <- sma(discarded_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1,method="SMA", log="xy")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$discarded_costs,SummarySpp[["se"]]$discarded_costs)
acc_by_species <- dplyr::bind_rows(acc_by_species, dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(prepollen_all_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="all pre-pollination costs (mg/seed)",xlab="",cex=1.5,xaxt="n")
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prepollen_all_costs,SummarySpp[["se"]]$prepollen_all_costs)
slopes <- sma(prepollen_all_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1, log="xy",method="SMA")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
acc_by_species <- dplyr::bind_rows(acc_by_species, dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(prepollen_discarded_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="aborted pre-pollination costs (mg/seed)",xlab="",cex=1.5,xaxt="n")
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prepollen_discarded_costs,SummarySpp[["se"]]$prepollen_discarded_costs)
slopes <- sma(prepollen_discarded_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1, log="xy",method="SMA")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
acc_by_species <- dplyr::bind_rows(acc_by_species, dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

plot(postpollen_aborted_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylim=c(.02,10000), ylab="aborted during provisioning costs (mg/seed)",xlab="",cex=1.5,xaxt="n")
y.error.bars(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$postpollen_aborted_costs,SummarySpp[["se"]]$postpollen_aborted_costs)
slopes <- sma(postpollen_aborted_costs ~ embryo_endo_costs,data=subset(SummarySpp[["mean"]]),slope.test=1, log="xy",method="SMA")
abline(slopes$groupsummary["Int"][[1]],slopes$groupsummary["Slope"][[1]])
abline(slopes$groupsummary["Int"][[1]],1,lty=2)
acc_by_species <- dplyr::bind_rows(acc_by_species, dplyr::select(slopes$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI,Slope_test_p))

acc_by_species$variable = c("total accessory costs","seed costs","pollen attraction costs","provisioning costs", "failed tissues costs","all pre-pollen costs per seed","failed pre-pollination costs","failed provisioning costs")

acc_by_species$Slope <- format(acc_by_species$Slope,nsmall=2,digits=2)
acc_by_species$Slope_lowCI <- format(acc_by_species$Slope_lowCI,nsmall=2,digits=2)
acc_by_species$Slope_highCI <- format(acc_by_species$Slope_highCI,nsmall=2,digits=2)

acc_by_species$slope2 <- paste(acc_by_species$Slope," (",acc_by_species$Slope_lowCI," - ",acc_by_species$Slope_highCI,")",sep="")

acc_by_species$r2 <- format(acc_by_species$r2,digits=2)
acc_by_species$pval <- format_p(acc_by_species$pval)
acc_by_species$Slope_test_p <- format(acc_by_species$Slope_test_p,digits=2)
acc_slopes <- dplyr::select(acc_by_species,variable,r2,slope2,Slope_test_p)

write.csv(acc_slopes, file="ms/Accessory/Table02_acc_cost_embryo_size_slopes.csv",row.names=FALSE)
```

#Table Supplementary.correlations between repro inv and propagule inv by species
```{r}
mod_propagule_inv_all <- sma(repro_inv~propagule_inv,data=SummaryInd,log="xy",method="SMA")
mod_propagule_inv_all

mod_flower_inv_all <- sma(repro_inv~flower_inv,data=SummaryInd,log="xy",method="SMA")
mod_flower_inv_all

yvar <- "repro_inv"
xvar <- "propagule_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy",slope.test=mod_propagule_inv_all$groupsummary["Slope"][[1]],elev.test=mod_propagule_inv_all$coef[[1]][1,1])
results <- fit$groupsummary
results$r2 <- format(results$r2,digits=2)
results$pval <- format(results$pval,digits=3)
acc_slopes <- dplyr::select(results,group,n,r2,pval)
names(acc_slopes) <- c("species","n_propagule","r2_propagule","p_propagule")

yvar <- "repro_inv"
xvar <- "flower_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA",  log="xy",slope.test=mod_flower_inv_all$groupsummary["Slope"][[1]],elev.test=mod_flower_inv_all$coef[[1]][1,1])
results <- fit$groupsummary
results$r2 <- format(results$r2,digits=2)
results$pval <- format(results$pval,digits=3)
acc_slopes2 <- dplyr::select(results,group,n,r2,pval)
names(acc_slopes2) <- c("species","n_flower","r2_flower","p_flower")
acc_slopes <- dplyr::left_join(acc_slopes, acc_slopes2, by ="species")
acc_slopes$species <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris microphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia", "Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile pulchella", "Phyllota phylicoides", "Pimelea linifolia", "Pultenaea tuberculata")

write.csv(acc_slopes, file="ms/Accessory/Table_SXX_propagule_invest_vs_all_repropro_inv.csv",row.names=FALSE)
```

#Table 3. correlations with total repro inv
```{r}
results <- as.data.frame(matrix(vector(), 0, 9,
                dimnames=list(c(), c("xvar","variable","r2","pval","Slope","Slope_lowCI","Slope_highCI","Int","Slope_test_p"))),
                stringsAsFactors=F)

for (i in c("total_weight","embryo_endo_inv","propagule_inv","fruit_inv","flower_inv","flower_inv2","success_inv","prepollen_success_inv","pack_disp_success_inv","discarded_inv")) {
  for (j in c("repro_inv")) {
yvar <- i
xvar <- j
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)
fit <- sma(data[[yvar]] ~ data[[xvar]], method="OLS", log="xy",slope.test=1)
temp <- as.data.frame(c(xvar, yvar,dplyr::select(fit$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Slope_test_p,n)))
colnames(temp)[2]<- "variable"
colnames(temp)[1]<- "xvar"
results <- dplyr::bind_rows(results, temp)
  }
}

results$slope2 <- paste(format(results$Slope,digits=3)," (",format(results$Slope_lowCI,digits=3)," - ",format(results$Slope_highCI,digits=3),")",sep="")

results$r2 <- format(results$r2,digits=3)
results$pval <- format(results$pval,digits=3)
results$Slope_test_p <- format(results$Slope_test_p,digits=3)
#results  <- dplyr::select(results,variable,n,r2,pval,slope2,Slope_test_p)
results  <- dplyr::select(results,variable,n,r2,pval)

write.csv(results, file="ms/Accessory/Table03_correlations_with_repro_inv.csv",row.names=FALSE)
```

#Figure 01. predictions
```{r}
win.metafile("ms/Accessory/Figure01b&c_predictions.wmf", height=3, width=6)
plot

par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.05,.2), mai=c(.8,.8,.1,.05)) 
plot(0,0,ylim=c(0,1),yaxt="n",xaxt="n",xlim=c(0,1),col="white",xlab="",ylab="")
segments(x0=.2,x1=.9,y0=.9,y1=.2,lty=2)
segments(x0=.1,x1=.8,y0=.8,y1=.1,lty=1)
axis(1, at=c(0,1),labels=c("lower","higher"),cex.axis=1)
axis(2, at=c(0,1),labels=c("lower","higher"),cex.axis=1,las=2)
mtext(text="count, scaled to plant size",side=2,line=1.1,outer=FALSE,cex=0.8)
mtext(text="costs (mg/seed)",side=1,line=1.0,outer=FALSE,cex=0.8)
legend("bottomleft",legend=c("seed count by reproductive costs","initiated ovules count by pollen-attraction costs"),col=c("black","black"),pch=c(16,21),bty="n",cex=0.5)
extra.top.left("b",2,1)

plot(0,0,ylim=c(0,1),yaxt="n",xaxt="n",xlim=c(0,1),col="white",xlab="",ylab="")
segments(x0=.1,x1=.9,y0=.9,y1=.1,lty=1)
axis(1, at=c(0,1),labels=c("lower","higher"),cex.axis=1)
axis(2, at=c(0,1),labels=c("lower","higher"),cex.axis=1,las=2)
mtext(text="choosiness (ovule:seed)",side=2,line=1.3,outer=FALSE,cex=0.8)
mtext(text="pollen attraction costs,",side=1,line=1.1,outer=FALSE,cex=0.8)
mtext(text="scaled to plant size",side=1,line=1.9,outer=FALSE,cex=0.8)
extra.top.left("c",2,1)
dev.off()

win.metafile("ms/Accessory/Figure01def_predictions.wmf", height=6, width=6)
plot

par(mfrow=c(2,2), cex=1, omi=c(.1,.1,.05,.2), mai=c(.8,.8,.1,.05))

plot(0,0,ylim=c(0,1),yaxt="n",xaxt="n",xlim=c(0,1),col="white",xlab="",ylab="")
segments(x0=.1,x1=.9,y0=.9,y1=.1,lty=1)
segments(x0=.15,x1=.9,y0=.15,y1=.9,lty=2)
axis(2, at=c(0,1),labels=c(0,1),cex.axis=1,las=2)
axis(1, at=c(0,1),labels=c("smaller","larger"),cex.axis=1)
mtext(text="proportion total success costs",side=2,line=1.8,outer=FALSE,cex=0.8)
mtext(text="seed size (mg)",side=1,line=1.3,outer=FALSE,cex=0.8)
legend("bottomleft",legend=c("pollen-attraction costs","provisioning costs"),col=c("black","black"),pch=c(16,21),bty="n",cex=0.5)
extra.top.left("d",2,1)

plot(0,0,ylim=c(0,1),yaxt="n",xaxt="n",xlim=c(0,1),col="white",xlab="",ylab="")
segments(x0=.1,x1=.9,y0=.9,y1=.1,lty=1)
segments(x0=.15,x1=.85,y0=.15,y1=.85,lty=2)
axis(2, at=c(0,1),labels=c(0,1),cex.axis=1,las=2)
axis(1, at=c(0,1),labels=c("smaller","larger"),cex.axis=1)
mtext(text="proportion all pollen-attraction investment",side=2,line=1.8,outer=FALSE,cex=0.8)
mtext(text="seed size (mg)",side=1,line=1.3,outer=FALSE,cex=0.8)
legend("bottomleft",legend=c("become seeds","are aborted before provisioning"),title="into ovules that:",col=c("black","black"),pch=c(16,21),bty="n",cex=0.5,xjust=0)
extra.top.left("e",2,1)

plot(0,0,ylim=c(0,1),yaxt="n",xaxt="n",xlim=c(0,1),col="white",xlab="",ylab="")
segments(x0=.1,x1=.9,y0=.9,y1=.1,lty=2)
segments(x0=.15,x1=.85,y0=.15,y1=.85,lty=1)
axis(2, at=c(0,1),labels=c(0,1),cex.axis=1,las=2)
axis(1, at=c(0,1),labels=c("smaller","larger"),cex.axis=1)
mtext(text="proportion all provisioning investment",side=2,line=1.8,outer=FALSE,cex=0.8)
mtext(text="seed size (mg)",side=1,line=1.3,outer=FALSE,cex=0.8)
legend("bottomleft",legend=c("become seeds","are aborted"),title="into zygotes that:",col=c("black","black"),pch=c(16,21),bty="n",cex=0.5,xjust=0)
extra.top.left("f",2,1)

dev.off()
```

#Figure 2
```{r}
win.metafile("ms/Accessory/Figure02.wmf", height=9, width=6)
plot

par(mfrow=c(3,2), cex=1, omi=c(.1,.1,.05,.05), mai=c(.7,.8,.05,.05)) 
options(scipen = 5)

#panel a
plot(scaled_seed_count~repro_costs,SummarySpp[["mean"]],pch=16,log="xy",col=venetian_red,xlim=c(.01,5000),ylim=c(5e-4,500),yaxt="n",xaxt="n",ylab="",xlab="",cex=1)
points(scaled_reach_flowering_count~pollen_attract_costs,SummarySpp[["mean"]],pch=16,col=cerulean,cex=1)

axis(2, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=0.8,las=2)
axis(1, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=0.8)
mtext("costs to make a successful unit",1,line=1.8,cex=0.7)
mtext("count, scaled to leaf area",2,line=2.3,cex=0.7)

mod_seed <- sma(scaled_seed_count~repro_costs,data=SummarySpp[["mean"]],log="xy",slope=-1,method="SMA")
mod_ovule2 <- sma(scaled_ovule_count~pollen_attract_costs,data=SummarySpp[["mean"]],log="xy",method="SMA",slope.test=mod_seed$groupsummary["Slope"][[1]],elev.test=mod_seed$coef[[1]][1,1])
mod_ovule <- sma(scaled_reach_flowering_count~pollen_attract_costs,data=SummarySpp[["mean"]],log="xy",method="SMA",slope.test=mod_seed$groupsummary["Slope"][[1]],elev.test=mod_seed$coef[[1]][1,1])

line_function_log_xy(SummarySpp[["mean"]]$repro_costs,mod_seed,1,col=venetian_red)
line_function_log_xy(SummarySpp[["mean"]]$pollen_attract_costs,mod_ovule,1,col=cerulean)
legend("bottomleft",legend=c("seed count by reproductive costs","mature ovules count by","pollen-attraction costs"),col=c(venetian_red,cerulean,"white"),pch=c(16,16,16),bty="n",cex=0.7)
extra.top.left.logxy("a",2,1)

#panel b
#plot(choosiness~prop_pollen_attract_vs_all_repro_all_repro,SummarySpp[["mean"]],log="xy",col=cerulean,pch=16,xlim=c(0.002,.5),ylim=c(3,800),xaxt="n",yaxt="n",cex=1,ylab="",xlab="")
plot(choosiness2~scaled_pollen_attract_costs,SummarySpp[["mean"]],log="xy",col=cerulean,pch=16,xlim=c(.002,.3),ylim=c(3,800),yaxt="n",cex=1,ylab="",xlab="",xaxt="n")
axis(2, at=c(10,100,1000), labels=c(10,100,1000),cex.axis=0.8,las=2)
axis(1, at=c(0.01,0.1), labels=c(0.01,0.1),cex.axis=0.8)
mtext("pollen-attractions costs, scaled to leaf area",1,line=1.7,cex=0.7)
mtext("choosiness (ovules:seed)",2,line=2.3,cex=0.7)
mod_choosy <- sma(choosiness2~scaled_pollen_attract_costs,data=subset(SummarySpp[["mean"]]),log="xy",slope=-1,method="SMA")
mod_choosy2 <- sma(choosiness2~scaled_pollen_attract_costs,data=subset(SummarySpp[["mean"]],species!="EPMI"),log="xy",slope=-1,method="SMA")
line_function_log_xy(SummarySpp[["mean"]]$scaled_pollen_attract_costs,mod_choosy2,1,col=cerulean)
extra.top.left.logxy("b",2,1)

#panel c - success proportions
plot(prop_pollen_attract_vs_success~embryo_endo_costs,SummarySpp[["mean"]],pch=16,log="x",col=cerulean,
     ylim=c(-0.1,1.1),xlim=c(.01,100),xlab="",ylab="",xaxt="n",yaxt="n")
points(prop_provisioning_vs_success~embryo_endo_costs,SummarySpp[["mean"]],pch=16,col=venetian_red)
#arrows(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prop_pollen_attract_vs_success-SummarySpp[["se"]]$prop_pollen_attract_vs_success,SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prop_pollen_attract_vs_success+SummarySpp[["se"]]$prop_pollen_attract_vs_success,code=3,angle=90, length=0.1,col=cerulean)
extra.top.left.logx("c",2,1)
axis(2, at=c(0,.2,.4,.6,.8,1), labels=c(0,0.2,0.4,0.6,0.8,1),cex.axis=0.8,las=2)
axis(1, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=0.8)
mtext("seed size (mg)",1,line=1.8,cex=0.7)
mtext("prop. successful investment",2,line=2.3,cex=0.7)
legend("bottomleft",legend=c("pollen_attraction costs","provisioning costs"),col=c(cerulean,venetian_red),pch=c(16,16),bty="n",cex=0.7)

mod_prop_pollen_attract_vs_success <- sma(prop_pollen_attract_vs_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_provisioning_vs_success <- sma(prop_provisioning_vs_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")

line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_pollen_attract_vs_success,1,col=cerulean)
line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_provisioning_vs_success,1,col=venetian_red)

#panel d - pollen-attraction proportions
plot(prop_prepollen_success~embryo_endo_costs,SummarySpp[["mean"]],pch=16,log="x",col=cerulean,ylim=c(-.1,1.1),xlim=c(.01,100),xlab="",ylab="",xaxt="n",yaxt="n")
points(prop_prepollen_discarded~embryo_endo_costs,SummarySpp[["mean"]],pch=16,col=venetian_red)
#arrows(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prop_prepollen_success-SummarySpp[["se"]]$prop_prepollen_success,SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prop_prepollen_success+SummarySpp[["se"]]$prop_prepollen_success,code=3,angle=90, length=0.1,col=cerulean)
#arrows(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prop_prepollen_discarded-SummarySpp[["se"]]$prop_prepollen_discarded,SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$prop_prepollen_discarded+SummarySpp[["se"]]$prop_prepollen_discarded,code=3,angle=90, length=0.1,col=venetian_red)
extra.top.left.logx("d",2,1)
axis(2, at=c(0,.2,.4,.6,.8,1), labels=c(0,0.2,0.4,0.6,0.8,1),cex.axis=0.8,las=2)
axis(1, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=0.8)
mtext("seed size (mg)",1,line=1.8,cex=0.7)
mtext("prop. total pollen-attraction investment",2,line=2.3,cex=0.7)
legend("bottomleft",legend=c("successful ovules","discarded ovules"),col=c(cerulean,venetian_red),pch=c(16,16),bty="n",cex=0.7)

mod_prop_prepollen_success <- sma(prop_prepollen_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_prepollen_discarded <- sma(prop_prepollen_discarded~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")

line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_prepollen_success,1,col=cerulean)
line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_prepollen_discarded,1,col=venetian_red)

#panel e - provisioning proportions
plot(prop_postpollen_success~embryo_endo_costs,SummarySpp[["mean"]],pch=16,log="x",col=cerulean,ylim=c(-0.1,1.1),xlim=c(.01,100),xlab="",ylab="",xaxt="n",yaxt="n")
points(prop_postpollen_discarded~embryo_endo_costs,SummarySpp[["mean"]],pch=16,col=venetian_red)
extra.top.left.logx("e",2,1)
axis(2, at=c(0,.2,.4,.6,.8,1), labels=c(0,0.2,0.4,0.6,0.8,1),cex.axis=0.8,las=2)
axis(1, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=0.8)
mtext("seed size (mg)",1,line=1.8,cex=0.7)
mtext("prop. total provisioning investment",2,line=2.3,cex=0.7)
legend("bottomleft",legend=c("successful zygotes","aborted zygotes"),col=c(cerulean,venetian_red),pch=c(16,16),bty="n",cex=0.7)

mod_prop_postpollen_success <- sma(prop_postpollen_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_postpollen_discarded <- sma(prop_postpollen_discarded~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")

line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_postpollen_success,1,col=cerulean)
line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_postpollen_discarded,1,col=venetian_red)

#panel f - change in slope of "success costs"
plot(pollen_attract_costs~embryo_endo_costs,SummarySpp[["mean"]],pch=16,log="xy",col=cerulean,xlim=c(.002,100),ylim=c(.01,20000),xlab="",ylab="",xaxt="n",yaxt="n")
points(provisioning_costs~embryo_endo_costs,SummarySpp[["mean"]],pch=16,col=venetian_red)
extra.top.left.logxy("f",2,1)
axis(2, at=c(0.01,1,100,10000), labels=c(0.01,1,100,10000),cex.axis=0.8,las=2)
axis(1, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=0.8)
mtext("seed size (mg)",1,line=1.8,cex=0.7)
mtext("success cost (mg/seed)",2,line=2.3,cex=0.7)
legend("bottomright",legend=c("pollen_attraction","provisioning"),col=c(cerulean,venetian_red),pch=c(16,16),bty="n",cex=0.7)

mod_repro_costs <- sma(repro_costs~embryo_endo_costs,SummarySpp[["mean"]],log="xy",method="SMA")
mod_pollen_attract_costs <- sma(pollen_attract_costs~embryo_endo_costs,SummarySpp[["mean"]],log="xy",method="SMA",slope.test = mod_repro_costs$groupsummary$Slope_lowCI)
mod_provisioning_costs <- sma(provisioning_costs~embryo_endo_costs,SummarySpp[["mean"]],log="xy",method="SMA",slope.test = mod_repro_costs$groupsummary$Slope)

line_function_log_xy(SummarySpp[["mean"]]$embryo_endo_costs,mod_pollen_attract_costs,1,col=cerulean)
line_function_log_xy(SummarySpp[["mean"]]$embryo_endo_costs,mod_provisioning_costs,1,col=venetian_red)

dev.off()
```

#Table with above model results
```{r}
mod_prop_pollen_attract_vs_all_repro <- sma(prop_pollen_attract_vs_all_repro~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_provisioning_vs_success <- sma(prop_provisioning_vs_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_prepollen_success <- sma(prop_prepollen_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_prepollen_success_SMA <- sma(prop_prepollen_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="SMA")
mod_prop_prepollen_discarded <- sma(prop_prepollen_discarded~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_postpollen_success <- sma(prop_postpollen_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")
mod_prop_postpollen_discarded <- sma(prop_postpollen_discarded~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="OLS")

energy_prop <- dplyr::select(mod_prop_pollen_attract_vs_all_repro$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI)
energy_prop$test[1] <- "Proportion sucess costs to prepollen"

temp <- dplyr::select(mod_prop_prepollen_success$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI)
temp$test[1] <- "Proportion prepollen costs to success"
energy_prop <- rbind(energy_prop,temp)

temp <- dplyr::select(mod_prop_postpollen_success$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI)
temp$test[1] <- "Proportion postpollen costs to success"
energy_prop <- rbind(energy_prop,temp)

#temp <- dplyr::select(mod_prop_provisioning_vs_success$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI)
#temp$test[1] <- "Proportion postpollen costs"
#energy_prop <- rbind(energy_prop,temp)

#temp <- dplyr::select(mod_prop_prepollen_discarded$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI)
#temp$test[1] <- "Proportion prepollen costs to discarded"
#energy_prop <- rbind(energy_prop,temp)

#temp <- dplyr::select(mod_prop_postpollen_discarded$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Int_lowCI,Int_highCI)
#temp$test[1] <- "Proportion postpollen costs to discarded"
#energy_prop <- rbind(energy_prop,temp)

```

#Figure 2 OLD. Scaling of accessory costs versus weight of embryo + endosperm, by species
```{r}
win.metafile("ms/Accessory_Costs/Figure02_embryo_size_acc_cost_correlations.wmf", height=9, width=10)
plot

par(mfrow=c(2,2), cex=1, omi=c(1,.1,.05,.05), mai=c(.2,.8,.1,.05)) 
options(scipen = 5)

plot((ovule_count/seed_count)~prop_pollen_attract_vs_all_repro_all_repro,SummarySpp[["mean"]],pch=16,log="xy",col=col.spp(species),ylab="",ylim=c(.8,500),yaxt="n",xlim=c(.0015,.5),xaxt="n")
axis(2, at=c(1,10,100), labels=c(1,10,100),cex.axis=1)
axis(1, at=c(0.001,0.005,0.01,0.05,0.1,0.5), labels=c(0.001,0.005,0.01,0.05,0.1,0.5),cex.axis=1)
mtext("choosiness (ovules initiated per seed matured)",2,line=2.3)
mtext("proportion reproductive investment for successful pollen_attraction costs",1,line=2.3)

mod_choosy <- sma((ovule_count/seed_count)~prop_pollen_attract_vs_all_repro_all_repro,data=SummarySpp[["mean"]],log="xy",slope=-1,method="SMA")

line_function_log_xy(SummarySpp[["mean"]]$prop_pollen_attract_vs_all_repro_all_repro,mod_choosy,1)
summary(mod_choosy)

plot(seedset~embryo_endo_costs,SummarySpp[["mean"]],log="xy",col=col.spp(species),pch=16, xlim=c(.002,500),ylim=c(.004,.6),cex=1.5,xaxt="n",xlab="",ylab="")
mod <- sma(seedset~embryo_endo_costs,SummarySpp[["mean"]],log="xy", method="SMA")
y.error.bars.2(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$seedset,SummarySpp[["se"]]$seedset)
#text((10^(0.99*(par("usr")[2]-par("usr")[1])+par("usr")[1])),(10^(0.97*(par("usr")[4]-par("usr")[3])+par("usr")[3])),paste("r2 = ",round(sma_sum(mod)[1],2)),adj=1)
line_function_log_xy(SummarySpp[["mean"]]$embryo_endo_costs,mod,1)
axis(1, at=c(0.01,1,100), labels=c("","",""),cex.axis=1)
mtext("seed set",2,line=2.3)
extra.top.left.logxy("b",2,1.2)
summary(mod)

plot(pollen_attract_costs~embryo_endo_costs,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="xy", yaxt="n",xaxt="n", ylim=c(.02,5000), xlim=c(.002,500), ylab="",xlab="",cex=1.5)
y.error.bars.2(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$pollen_attract_costs,SummarySpp[["se"]]$pollen_attract_costs)

points(provisioning_costs~embryo_endo_costs,SummarySpp[["mean"]],col=col.spp(species),pch=21,cex=1.5)
y.error.bars.2(SummarySpp[["mean"]]$embryo_endo_costs,SummarySpp[["mean"]]$provisioning_costs,SummarySpp[["se"]]$provisioning_costs)

axis(2, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)
axis(1, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=1)
mtext("costs (mg/ovule)",2,line=2.3)

mod_prepollen <- sma(pollen_attract_costs~embryo_endo_costs,SummarySpp[["mean"]],log="xy",method="SMA")
mod_provision <- sma(postpollen_all_per_seed~embryo_endo_costs,SummarySpp[["mean"]],log="xy",method="SMA",slope.test=mod_prepollen$groupsummary["Slope"][[1]])

line_function_log_xy(SummarySpp[["mean"]]$embryo_endo_costs,mod_prepollen,1)
line_function_log_xy(SummarySpp[["mean"]]$embryo_endo_costs,mod_provision,2)

legend("topright",legend=c("pollen_attraction","provisioning"),col=c("black","black"),pch=c(16,21),bty="n")
extra.top.left.logxy("c",2,1.2)

plot(prop_pollen_attract_vs_all_repro~embryo_endo_costs,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="x", yaxt="n",xaxt="n", ylim=c(0,1), xlim=c(.002,500), ylab="",xlab="",cex=1.5)
points(prop_prepollen_all_vs_all_repro~embryo_endo_costs,SummarySpp[["mean"]],col=col.spp(species),pch=1,cex=1.5)
axis(2, at=c(0,.5,1), labels=c(0,0.5,1.0),cex.axis=1)
axis(1, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=1)
extra.top.left.logx("d",2,1.2)

mod_prop_pollen_attract_vs_all_repro <- sma(prop_pollen_attract_vs_all_repro~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="SMA",slope.test=0)
mod_prop_prepollen_all_vs_all_repro <- sma(prop_prepollen_all_vs_all_repro~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="SMA",slope.test=0)

line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_pollen_attract_vs_all_repro,1)

summary(mod_prop_pollen_attract_vs_all_repro)
summary(mod_prop_prepollen_all_vs_all_repro)

plot(prop_postpollen_success~embryo_endo_costs,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="x", yaxt="n",xaxt="n", ylim=c(0,1), xlim=c(.002,500), ylab="",xlab="",cex=1.5)
points(prop_prepollen_success~embryo_endo_costs,SummarySpp[["mean"]],col=col.spp(species),pch=1,cex=1.5)
axis(2, at=c(0,.5,1), labels=c(0,0.5,1.0),cex.axis=1)
axis(1, at=c(0.01,1,100), labels=c(0.01,1,100),cex.axis=1)
extra.top.left.logx("d",2,1.2)

mod_prop_postpollen_success <- sma(prop_postpollen_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="SMA")
mod_prop_prepollen_success <- sma(prop_prepollen_success~embryo_endo_costs,SummarySpp[["mean"]],log="x",method="SMA")

line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_postpollen_success,1)
line_function_log_x(SummarySpp[["mean"]]$embryo_endo_costs,mod_prop_prepollen_success,3)

summary(mod_prop_postpollen_success)
summary(mod_prop_prepollen_success)

mtext(text="     seed size (mg)",side=1,line=1.5,outer=TRUE,adj=0.5)
mtext(text="                     parental pessimist",line=3,side=1,adj=0,outer=TRUE,font=3)
mtext(text="parental optimist    ",line=3,side=1,adj=1,outer=TRUE,font=3)

dev.off()
```

#Figure 3. Correlations of various reproductive investment categories against plant weight
```{r}
win.metafile("ms/Accessory/Figure03_correlations_with_plant_weight.wmf", height=5, width=10)
plot

par(mfrow=c(1,2), cex=1, omi=c(.1,.6,.1,.05), mai=c(.8,.1,.1,.05))
options(scipen = 4)

library(smatr)

yvar <- "repro_inv"
xvar <- "embryo_endo_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),log="xy",ylim=c(.3,300000),xlim=c(.3,300000), xlab="",ylab="",lty=results$linetype,pch=16)
abline(0,1)
mod <- sma(embryo_endo_inv~repro_inv,data=subset(SummaryInd,propagule_inv>0),log="xy",method="OLS")
mtext(text="total embryo and endosperm investment (mg)",side=1,line=2.2,outer=FALSE)
mtext(text="total reproductive investment (mg)",side=2,line=2,outer=TRUE)
extra.top.left.logxy("a",2,1.2)

xvar <- "flower_inv"
yvar <- "repro_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="OLS", log="xy",slope.test=1)
results <- lines(fit$groupsummary)
plot(fit,xlab="",ylab="",pch=16, col=col.spp(fit$groupsummary$group),lty=results$linetype,ylim=c(.3,300000),xlim=c(.3,300000),yaxt="n")
axis(2, at=c(1,100,10000),labels=c("","",""))
abline(0,1)
mtext(text="flower weight * bud count (mg)",side=1,line=2.2,outer=FALSE)
extra.top.left.logxy("b",2,1.2)

dev.off()
```

#Supplementary Table. Higher RE, RA, plant size do not lead to lower accessory costs; age shows mixed results
```{r}
win.metafile("ms/Accessory/Figure_SXX_RE_RA_size_age_patterns.wmf", height=10, width=10)
plot

par(mfrow=c(2,2), cex=1, omi=c(.1,.4,.1,.05), mai=c(.6,.3,.1,.1))
options(scipen = 2)

xvar <- "repro_inv"
yvar <- "accessory_costs"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab=yvar,lty=results$linetype)
mtext(text="reproductive investment (mg) (RE)",side=1,line=2,outer=FALSE)
legend_with_r2(results,"bottomright")

write.csv(results, file="ms/Accessory/Table05a_accessory_vs_all_repro.csv",row.names=FALSE)

xvar <- "total_weight_0"
yvar <- "accessory_costs"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab=yvar,lty=results$linetype)
mtext(text="total plant size (mg)",side=1,line=2,outer=FALSE)
legend_with_r2(results,"bottomright")

write.csv(results, file="ms/Accessory/Table05b_accessory_vs_total_weight.csv",row.names=FALSE)

xvar <- "age"
yvar <- "accessory_costs"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="",ylab="",lty=results$linetype)
mtext(text="age (years)",side=1,line=2,outer=FALSE)
mtext(text="per seed accessory costs (mg/seed)",side=2,line=1,outer=TRUE)
legend_with_r2(results,"topleft")

write.csv(results, file="ms/Accessory/Table05c_accessory_vs_age.csv",row.names=FALSE)
dev.off()
```

#Table S1.correlations between repro inv and flower inv by species (noted in Discussiong)
```{r}
yvar <- "repro_inv"
xvar <- "embryo_endo_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- fit$groupsummary
results$r2 <- format(results$r2,digits=2)
results$pval <- format(results$pval,digits=3)
acc_slopes <- dplyr::select(results,group,n,r2,pval)

yvar <- "repro_inv"
xvar <- "flower_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- fit$groupsummary
results$r2 <- format(results$r2,digits=2)
results$pval <- format(results$pval,digits=3)
acc_slopes2 <- dplyr::select(results,group,n,r2,pval)

acc_slopes <- merge(acc_slopes,acc_slopes2,by="group")

write.csv(acc_slopes, file="ms/Accessory/Table_S1_extra_flower_invest_vs_all_repro_inv.csv",row.names=FALSE)
```

#Table 7 extra. correlations with total repro inv
```{r}
results <- as.data.frame(matrix(vector(), 0, 9,
                dimnames=list(c(), c("xvar","variable","r2","pval","Slope","Slope_lowCI","Slope_highCI","Int","Slope_test_p"))),
                stringsAsFactors=F)

for (i in c("total_weight","repro_inv","discarded_inv","success_inv","propagule_inv","packaging_dispersal_inv","prepollen_success_inv","fruit_inv","flower_inv")) {
  for (j in c("total_weight","leaf_area_0","diameter_0","repro_inv","propagule_inv")) {
yvar <- i
xvar <- j
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)
fit <- sma(data[[yvar]] ~ data[[xvar]], method="SMA", log="xy",slope.test=1)
temp <- as.data.frame(c(xvar, yvar,dplyr::select(fit$groupsummary,r2,pval,Slope,Slope_lowCI,Slope_highCI,Int,Slope_test_p,n)))
colnames(temp)[2]<- "variable"
colnames(temp)[1]<- "xvar"
results <- bind_rows(results, temp)
  }
}

write.csv(results, file="ms/Accessory_Costs/Table07_EXTRA_correlations_with_repro_inv.csv",row.names=FALSE)
```

###GOOD VISUALS - not currently in manuscript
```{r}
xvar <- "leaf_area_0"
yvar <- "flower_inv"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),log="xy",ylim=c(1,300000),xlim=c(1,10000000),xlab="",ylab="",lty=results$linetype,pch=16)
results

plot(propagule_costs~total_weight_0,SummaryInd,log="xy",pch=16,col=col.spp(species))
plot(propagule_costs~total_weight_0,SummarySpp[["mean"]],log="xy",pch=16,col=col.spp(species))

```

```{r}
Cairo("ms/Accessory_Costs/FigureXX_seedcount_reprocount.svg", height=5, width=5,type="svg",pointsize=10,units="in")
#win.metafile("ms/Accessory_Costs/FigureXX_seedcount_reprocount.wmf", height=5, width=5)

plot

par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,0.05), mai=c(.8,.8,.1,0.1))
options(scipen = 2)

xvar <- "ovule_count"
yvar <- "seed_count"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy",slope=1)
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab="buds initiated (count)",ylab="seed count",lty=results$linetype,xlim=c(0.9,20000),ylim=c(0.9,20000))

dplyr::select(results,group,n,r2,pval,Slope,Slope_test_p)
abline(0,1,lwd=2)
abline(-2,1,lwd=2)

legend_with_r2(results,"topleft")

dev.off()
```

#Although very little energy goes into propagules (or embryos) a notable amount of energy is required for successful seeds to form
```{r}
win.metafile("ms/Accessory_Costs/FigureXX_success_vs_discarded.wmf", height=6, width=8)
plot

par(mfrow=c(1,2), cex=1, omi=c(.8,.1,.1,.1), mai=c(.1,.8,.1,.1))

xvar <- "repro_inv"
yvar <- "prop_propagule"

data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,lty=results$linetype,ylim=c(.001,1),xlab="",ylab="prop repro energy to propagules")
abline(-0.30103,0,lwd=2)
abline(-1,0,lwd=2)

xvar <- "repro_inv"
yvar <- "prop_success"

data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,lty=results$linetype,ylim=c(.001,1),xlab="",ylab="prop repro energy to success")
abline(-0.30103,0,lwd=2)
abline(-1,0,lwd=2)
mtext(text="reproductive investment (mg)",side=1,line=2,outer=TRUE)

dev.off()
```

#Neither within nor across species, does the proportion of reproductive energy spent on "success" correlate with total reproductive investment
```{r}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(1.2,1.2,.1,.05))

xvar <- "repro_inv"
yvar <- "prop_success"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], log="x",method="SMA")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),xlab=xvar,ylab=yvar,lty=results$linetype,pch=16,log="x")
```

#when reproductive investment is scaled by plant size there is no significant difference by seed size - so increase in repro inv with seed size simply because species with bigger seeds are bigger plants  
```{r}
plot(scaled_repro_inv~propagule_costs,SummaryInd, col=col.spp(species),pch=16,log="xy")
points(scaled_repro_inv~propagule_costs,SummarySpp[["mean"]], cex=2, col=col.spp(species),pch=16)
mod <-sma(scaled_repro_inv~propagule_costs,SummarySpp[["mean"]],log="xy",method="SMA")
summary(mod)
```

#This effect is true across species, since larger-seeded species produce fewer seeds and have higher per seed accessory costs and true within species, since absolute accessory investment is a function of plant size not seedset
```{r}
win.metafile("ms/Accessory_Costs/FigureXX_EXTRA_acc_cost_by_seed_count.wmf", height=5, width=5)
plot

par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1,1,.5,.5))

xvar <- "seed_count"
yvar <- "accessory_costs"
data <- subset(SummaryInd,SummaryInd[[xvar]]>0&SummaryInd[[yvar]]>0)

fit <- sma(data[[yvar]] ~ data[[xvar]]*data[["species"]], method="SMA", log="xy")
results <- lines(fit$groupsummary)
plot(fit,col=col.spp(fit$groupsummary$group),pch=16,xlab=xvar,ylab=yvar,lty=results$linetype)

points(accessory_costs~seed_count,data=subset(SummarySpp[["mean"]],accessory_costs>0&seed_count>0),col="black",pch=16,cex=2.5)
points(accessory_costs~seed_count,data=subset(SummarySpp[["mean"]],accessory_costs>0&seed_count>0),col=col.spp(species),pch=16,cex=2)
mod <- sma(accessory_costs~seed_count,data=subset(SummarySpp[["mean"]]),method="SMA", log="xy")
#words.top.right.logxy(mod)
legend_with_r2(results,"bottomleft")

write.csv(results, file="ms/Accessory_Costs/TableXX_seedcount_vs_accessory.csv",row.names=FALSE)

dev.off()
```