### 2016-12-13 scripts/Accessory_figs_2015_12.Rmd

```{r}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1,1,.5,.5)) 
plot(costs_for_seed ~ seed_size,data=subset(SummaryInd,repro_inv>0),col=col.spp(species),pch=16,log="xy")
points(costs_for_seed ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(seed_size) ~ log(costs_for_seed),data=subset(SummarySpp[["mean"]])) 
words.top.left.logxy(mod)

plot(costs_for_aborted ~ seed_size,data=subset(SummaryInd,repro_inv>0),col=col.spp(species),pch=16,log="xy")
points(costs_for_aborted ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(seed_size) ~ log(costs_for_aborted),data=subset(SummarySpp[["mean"]])) 
words.top.left.logxy(mod)

plot(accessory_inv ~ seed_size,data=subset(SummaryInd,repro_inv>0),col=col.spp(species),pch=16,log="xy")
points(accessory_inv ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(seed_size) ~ log(accessory_inv),data=subset(SummarySpp[["mean"]])) 
words.top.left.logxy(mod)
abline(mod)

plot(costs_for_seed ~ costs_for_aborted,data=subset(SummaryInd,repro_inv>0),col=col.spp(species),pch=16,log="xy")
points(costs_for_seed ~ costs_for_aborted,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(costs_for_seed) ~ log(costs_for_aborted),data=subset(SummarySpp[["mean"]]))
words.top.left.logxy(mod)

plot(costs_for_seed ~ repro_inv,data=subset(SummaryInd,repro_inv>0),col=col.spp(species),pch=16,log="xy")
abline(0,1)
```

SummaryInd$prop_costs_seed <- SummaryInd$costs_for_seed / SummaryInd$repro_inv
SummaryInd$prop_costs_aborted <- SummaryInd$costs_for_aborted / SummaryInd$repro_inv

SummaryInd$check <- SummaryInd$prepollen_all_inv / (SummaryInd$count_aborted + SummaryInd$seed_count)

plot(check ~ prepollencosts,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
abline(0,1)

```{r}
par(mfrow=c(2,2), cex=1, omi=c(.1,.1,.1,.1), mai=c(1,1,.5,.5)) 
options(scipen = 999)

plot(seedcosts ~ seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(seedcosts),data=subset(SummaryInd, seedcosts>0))
words.bottom.right.logxy(mod)

plot(prepollencosts ~ seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(prepollencosts),data=subset(SummaryInd,prepollencosts>0))
words.bottom.right.logxy(mod)

plot(abortedcosts ~ seed_size,data=subset(SummaryInd, abortedcosts>.001),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(abortedcosts),data=subset(SummaryInd, abortedcosts>0.001))
words.bottom.right.logxy(mod)

plot(b_div_p ~ seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy",ylab="per seed cost to procur pollen")
mod <- lm(log(seed_size) ~ log(b_div_p),data=subset(SummaryInd,b_div_p>0))
words.bottom.right.logxy(mod)

plot(c_div_p ~ seed_size,data=subset(SummaryInd,c_div_p>0&species!="COER"),col=col.spp(species),pch=16,log="xy",ylab="per seed cost of aborted fruits and seeds")
mod <- lm(log(seed_size) ~ log(c_div_p),data=subset(SummaryInd,c_div_p>0&species!="COER"))
words.bottom.right.logxy(mod)

plot(dispersalcosts_increment ~ seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy",ylab="cost of dispersal tissues")
mod <- lm(log(seed_size) ~ log(dispersalcosts_increment),data=subset(SummaryInd))
words.bottom.right.logxy(mod)

plot(total_accessory_Lord ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylab="total per seed accessory costs")
mod <- lm(log(total_accessory_Lord) ~ log(seed_size),data=subset(SummarySpp[["mean"]],total_accessory_Lord>0))
words.bottom.right.logxy(mod)
summary(mod)

plot(accessory_per_seed ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylab="total per seed accessory costs")
mod <- lm(log(total_accessory_Lord) ~ log(seed_size),data=subset(SummarySpp[["mean"]],total_accessory_Lord>0))
words.bottom.right.logxy(mod)
summary(mod)
```

```{r}
plot(b_div_p ~ seed_size,data=subset(SummaryInd),col=col.age(age),pch=16,log="xy", ylab="per seed cost to procur pollen")
mod <- lm(log(b_div_p) ~ log(seed_size)*age,data=subset(SummaryInd,b_div_p>0))
words.bottom.right.logxy(mod)
summary(mod)
slopes <- sma(log(b_div_p) ~ log(seed_size),data=subset(SummaryInd,b_div_p>0),slope.test=1)
summary(slopes)

plot(c_div_p ~ seed_size,data=subset(SummaryInd,c_div_p>0.00001),col=col.age(age),pch=16,log="xy", ylab="per seed cost of aborted fruits and seeds")
mod <- lm(log(c_div_p) ~ log(seed_size)*age,data=subset(SummaryInd,c_div_p>0.00001))
words.bottom.right.logxy(mod)
summary(mod)
slopes <- sma(log(c_div_p) ~ log(seed_size),data=subset(SummaryInd,c_div_p>0),slope.test=1)
summary(slopes)

plot(dispersalcosts_increment ~ seed_size,data=subset(SummaryInd),col=col.age(age),pch=16,log="xy", ylab="per seed cost of packaging and disp")
mod <- lm(log(dispersalcosts_increment) ~ log(seed_size)*age,data=subset(SummaryInd,dispersalcosts_increment>0))
words.bottom.right.logxy(mod)
slopes <- sma(log(dispersalcosts_increment) ~ log(seed_size),data=subset(SummaryInd,dispersalcosts_increment>0),slope.test=1)
summary(slopes)

plot(total_accessory_Lord  ~ seed_size,data=subset(SummaryInd),col=col.age(age),pch=16,log="xy")
mod <- lm(log(total_accessory_Lord) ~ log(seed_size)*age,data=subset(SummaryInd,total_accessory_Lord>0))
words.bottom.right.logxy(mod)
summary(mod)
slopes <- sma(log(total_accessory_Lord) ~ log(seed_size),data=subset(SummaryInd,total_accessory_Lord>0),slope.test=1)
summary(slopes)
slopes2 <- as.matrix(coef(slopes))
abline(log(slopes2[1]),slopes2[2])
#calculation of total accessory investment divided by seed count (accessory_per_seed) very closely correlated with accessory costs per seed using the equations in Lord
plot(total_accessory_Lord  ~ accessory_per_seed,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(total_accessory_Lord) ~ log(accessory_per_seed),data=subset(SummaryInd,total_accessory_Lord>0))
words.bottom.right.logxy(mod)
slopes <- sma(log(total_accessory_Lord) ~ log(accessory_per_seed),data=subset(SummaryInd,total_accessory_Lord>0),slope.test=1)
summary(slopes)
summary(mod)

plot(seed_to_ovule_ratio ~ seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")

plot(seedset ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(seed_to_ovule_ratio),data=subset(SummaryInd,seed_to_ovule_ratio>0))
words.bottom.left.logxy(mod)

#on an energy basis, propagule inv and accessory inv increase 1 to 1; implies seed size is sufficient information, but need to account of additional component
plot(accessory_inv ~ propagule_inv,data=subset(SummaryInd,propagule_inv>0),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(propagule_inv) ~ log(accessory_inv)*age,data=subset(SummaryInd,propagule_inv>0))
words.bottom.right.logxy(mod)
slopes <- sma(log(propagule_inv) ~ log(accessory_inv),data=subset(SummaryInd,propagule_inv>0),slope.test=1)
summary(slopes)
```

```{r}
win.metafile("figs2/acc_cost_seed_size.wmf", height=6, width=6)

plot
plot(prepollencosts~seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy",cex=1.3)
mod <- lm(log(seed_size)~log(accessory_costs),data=subset(SummaryInd))
words.top.left.logxy(mod)
abline((log(10)^coef(mod)[1]),coef(mod)[2])
abline((log(10)^coef(mod)[1]),1,col="red")

summary(mod)
dev.off()
```

plot(seedset~seedcosts,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy",cex=1)
points(seedset~accessory_costs,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(accessory_costs)~log(seedset),data=subset(SummaryInd,seedset>0))
words.bottom.left.logxy(mod)
abline((log(10)^coef(mod)[1]),(coef(mod)[2]))
abline((log(10)^coef(mod)[1]),-1,col="red")

summary(mod)

```{r}
par(mfrow=c(2,3), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
#total reproductive investment much stronger correlation with total weight 
win.metafile("figs2/best_correlate2.wmf", height=6, width=12)
plot

par(mfrow=c(1,3), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(propagule_inv~total_weight_0,data=subset(SummaryInd,propagule_inv>0),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(total_weight_0)~log(propagule_inv),data=subset(SummaryInd,propagule_inv>0))
words.top.left.logxy(mod)

plot(repro_inv~total_weight_0,data=subset(SummaryInd,repro_inv>0),col=col.spp(species),pch=16,log="xy",xlim=c(10,10000000),ylim=c(1,1000000))
mod <- lm(log(total_weight_0)~log(repro_inv),data=subset(SummaryInd,repro_inv>0))
words.top.left.logxy(mod)

plot(prepollen_all_inv~total_weight_0,data=subset(SummaryInd,prepollen_all_inv>0),col=col.spp(species),pch=16,log="xy",xlim=c(10,10000000),ylim=c(1,1000000))
mod <- lm(log(total_weight_0)~log(prepollen_all_inv),data=subset(SummaryInd,prepollen_all_inv>0))
words.top.left.logxy(mod)

dev.off()

plot(propagule_inv~total_inv,data=subset(SummaryInd,propagule_inv>0),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(total_inv)~log(propagule_inv),data=subset(SummaryInd,propagule_inv>0))
words.top.left.logxy(mod)

plot(repro_inv~total_inv,data=subset(SummaryInd,repro_inv>0),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(total_inv)~log(repro_inv),data=subset(SummaryInd,repro_inv>0))
words.top.left.logxy(mod)

plot(prepollen_all_inv~total_inv,data=subset(SummaryInd,prepollen_all_inv>0),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(total_inv)~log(prepollen_all_inv),data=subset(SummaryInd,prepollen_all_inv>0))
words.top.left.logxy(mod)

#shift in accessory costs per seed with seed size
par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
win.metafile("figs2/accessory_per_seed_vs_seed_size.wmf", height=6, width=12)
plot
plot(costs_per_seed~seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
points(costs_per_seed~seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(seed_size)~log(costs_per_seed),data=subset(SummarySpp[["mean"]]))
words.top.left.logxy(mod)
summary(mod)

plot(seedcosts~seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy",cex=1.3)
mod <- lm(log(seed_size)~log(seedcosts),data=subset(SummaryInd))
words.top.left.logxy(mod)
dev.off()

#each species follows its own linear trajectory, but species lines are offset from one another
#species with lower intercepts have greater benefit if seed count is metric that matters; species with higher intercept less disadvntaged is a few heavy seeds worth as much as many smaller seeds
win.metafile("figs2/accessory_per_seed_vs_prop_propagule.wmf", height=6, width=6)
plot
plot(accessory_per_seed~prop_propagule,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
dev.off()
```

plot(seed_count~accessory_per_seed,data=subset(SummarySpp[["mean"]],seed_count>1),col=col.spp(species),pch=16,log="xy")
```{r}
plot(accessory_per_seed~growth_leaf,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
points(accessory_per_seed~growth_leaf,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(growth_leaf)~log(accessory_per_seed),data=subset(SummarySpp[["mean"]]))
words.top.left.logxy(mod)

plot(prop_propagule~growth_leaf,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
points(prop_propagule~growth_leaf,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(growth_leaf)~log(prop_propagule),data=subset(SummarySpp[["mean"]]))
words.bottom.left.logxy(mod)

plot(accessory_per_seed~seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size)~log(accessory_per_seed),data=subset(SummarySpp[["mean"]]))
words.top.left.logxy(mod)
abline(log(5),1)

plot(prop_propagule~seed_size,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
points(prop_propagule~seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)
mod <- lm(log(seed_size)~log(prop_propagule),data=subset(SummarySpp[["mean"]]))
words.top.left.logxy(mod)

plot(costs_per_seed~accessory_inv,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
plot(seedset~accessory_inv,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
```

plot(accessory_per_seed~prop_propagule,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
points(accessory_per_seed~prop_propagule,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,cex=2)

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(seedset~prepollen_all_count,data=subset(SummaryInd,seedset>0),col=col.spp(species),pch=16,log="xy")

plot(total_weight~prepollen_all_count,data=subset(SummaryInd,prepollen_all_count>0),col=col.spp(species),pch=16,log="xy")
plot(seed_size~prepollen_all_count,data=subset(SummaryInd,prepollen_all_count>0),col=col.spp(species),pch=16,log="xy")
#points(seed_size~prepollen_all_count,data=subset(SummarySpp[["mean"]],prepollen_all_count>0),col=col.spp(species),pch=16,cex=2)

plot(seed_size~seed_count,data=subset(SummaryInd,seed_count>0),col=col.spp(species),pch=16,log="xy")
points(seed_size~seed_count,data=subset(SummarySpp[["mean"]],seed_count>0),col=col.spp(species),pch=16,cex=2)

plot(seed_size~seedset,data=subset(SummaryInd,seed_count>0),col=col.spp(species),pch=16,log="xy")
points(seed_size~seedset,data=subset(SummarySpp[["mean"]],seed_count>0),col=col.spp(species),pch=16,cex=2)
mod<-lm(log(seedset)~log(seed_size),data=subset(SummaryInd,seed_count>0))
words.bottom.left.logxy(mod)
```

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(seedset~prop_propagule,SummaryInd,col=col.spp(species),pch=16,log="xy")
abline(0,1)

plot(seedset~propagule_per_seed,SummaryInd,col=col.spp(species),pch=16,log="xy")
points(seedset~propagule_per_seed,SummarySpp[["mean"]],col=col.spp(species),pch=16,cex=2)
mod <- lm(log(propagule_per_seed)~log(seedset),data=subset(SummaryInd,seedset>0))
words.bottom.left.logxy(mod)

plot(seedset~prop_propagule,SummaryInd,col=col.spp(species),pch=16,log="xy")
points(seedset~prop_propagule,SummarySpp[["mean"]],col=col.spp(species),pch=16,cex=2)
mod <- lm(log(prop_propagule)~log(seedset),data=subset(SummaryInd,seedset>0))
words.bottom.left.logxy(mod)
#text(seedset~propagule_per_seed,data=subset(SummaryInd,species=="PELA"),labels=individual,offset=1)

par(mfrow=c(4,4), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
xvar <- "seedset"
yvar <- "prop_propagule"

data <- subset(SummaryInd,seedset>0&prop_propagule>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age),xlim=c(0.0005,.6),ylim=c(0.0005,.6))
}
```

```{r}
par(mfrow=c(4,4), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
xvar <- "age"
yvar <- "seedset"

data <- subset(SummaryInd,seedset>0&age>3)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

```{r}
data <- subset(SummaryInd)
data <- split(data, data$species)

#pdf("figs2/Flower_parts_fewer.pdf", height=11, width=8)
#plot
par(mfrow=c(4,4), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
for(spp in names(data)) {
  pie(data[[spp]]$weight,labels=data[[spp]]$labels,main=spp,col=col.barplot(data[[spp]]$parts_fewer_yet),cex=0.8)
}
#dev.off()
```

```{r}
#par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))

```

```{r}
plot(accessory_inv~seedcosts,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
plot(seedset~seedcosts,data=subset(SummaryInd,seedset>0),col=col.spp(species),pch=16,log="x")
plot(seedset~costs_per_seed,data=subset(SummaryInd),col=col.spp(species),pch=16,log="x")
plot(packaging_dispersal_inv~accessory_inv,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
plot(propagule_inv~repro_inv,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
plot(propagule_inv~total_weight,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
plot(repro_inv~total_weight,data=subset(SummaryInd),col=col.spp(species),pch=16,log="xy")
abline(0,1)
```

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
xvar <- "total_weight"
yvar <- "prepollen_all_count"

data <- subset(SummaryInd,prepollen_all_count>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=14}
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
xvar <- "growth_leaf"
yvar <- "seedset"

data <- subset(SummaryInd,seedset>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

mod <- lm(log(seedset)~log(prepollen_all_count)*log(growth_leaf)*log(total_weight),data=subset(SummaryInd,seedset>0&growth_leaf>0))
summary(mod)
plot(mod)

mod <- lm(log(prepollen_all_count)~log(growth_leaf)*log(total_weight)*log(maxH),data=subset(SummaryInd,seedset>0&growth_leaf>0))
summary(mod)
plot(mod)

plot(seedset~total_weight,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="xy")
mod <- lm(log(total_weight)~log(seedset),SummarySpp[["mean"]])
words.top.right.logxy(mod)

plot(seedset~total_weight,data=subset(SummaryInd,seedset>0),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seedset)~log(total_weight)*species,data=subset(SummaryInd,seedset>0))
words.top.right.logxy(mod)
summary(mod)

plot(seedset~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="x")
mod <- lm(log(seed_size)~seedset,SummarySpp[["mean"]])
words.top.right.logx(mod)

plot(seedset~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="x")
mod <- lm(log(seed_size)~seedset,SummarySpp[["mean"]])
words.top.right.logx(mod)

plot(seedset~total_weight,data=subset(SummaryInd,seedset>0),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(total_weight)~log(seedset),data=subset(SummaryInd,seedset>0))
words.top.right.logxy(mod)

par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))

plot(accessory_inv~seed_size,SummaryInd,col=col.spp(species),pch=16,log="xy")
points(accessory_inv~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,cex=2)

plot(accessory_per_seed~seed_size,SummaryInd,col=col.spp(species),pch=16,log="xy")
points(accessory_per_seed~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,cex=2)

plot(prop_propagule~seed_size,SummaryInd,col=col.spp(species),pch=16,log="xy")
points(prop_propagule~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,cex=2)
mod <- lm (log(seed_size)~log(prop_propagule), data=subset(SummaryInd,prop_propagule>0&species!="EPMI"))
words.bottom.right.logxy(mod)

plot(prop_propagule~total_inv,SummaryInd,col=col.spp(species),pch=16,log="xy")
points(prop_propagule~total_inv,SummarySpp[["mean"]],col=col.spp(species),pch=16,cex=2)

names(Table1) <- c("Variable","r2","intercept","slope")
Table1[1] <- c("Total reproductive investment",summary(mod)[8]

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=14}
options(scipen=10)
par(mfrow=c(2,2), cex=1, omi=c(.6,0.3,.3,0.3), mai=c(.8,1,.3,0.3))
mod <- lm (log(total_weight)~log(repro_inv), data=subset(SummaryInd,repro_inv>0))
plot(repro_inv~total_weight,data=subset(SummaryInd,repro_inv>0), log="xy",xlab="", col=col.spp(species), pch=16, ylab="Total reproductive investment (mg)",xlim=c(100,2000000))
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(fruit_weight), data=subset(SummaryInd,fruit_weight>0))
plot(fruit_weight~total_weight,data=subset(SummaryInd,fruit_weight>0), log="xy",col=col.spp(species), pch=16, ylab="Fruit investment (mg)",xlim=c(100,2000000),xlab="")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(propagule_inv), data=subset(SummaryInd,propagule_inv>0))
plot(propagule_inv~total_weight,data=subset(SummaryInd,propagule_inv>0), log="xy",col=col.spp(species), pch=16, xlab="", ylab="Propagule investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(prepollen_all_inv), data=subset(SummaryInd,prepollen_all_inv>0))
plot(prepollen_all_inv~total_weight,data=subset(SummaryInd,prepollen_all_inv>0), log="xy",col=col.spp(species), pch=16, xlab="", ylab="Prepollen investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mtext("Total plant size (mg)",1, outer=TRUE,cex=1.2)
```

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=14, fig.height=14}
options(scipen=10)
par(mfrow=c(2,2), cex=1, omi=c(.6,0.3,.3,0.3), mai=c(.8,1,.3,0.3))
mod <- lm (log(total_inv)~log(repro_inv), data=subset(SummaryInd,repro_inv>0))
plot(repro_inv~total_inv,data=subset(SummaryInd,repro_inv>0), log="xy",xlab="", col=col.spp(species), pch=16, ylab="Total reproductive investment (mg)",xlim=c(100,2000000))
words.bottom.right.logxy(mod)

mod <- lm (log(total_inv)~log(fruit_weight), data=subset(SummaryInd,fruit_weight>0))
plot(fruit_weight~total_inv,data=subset(SummaryInd,fruit_weight>0), log="xy",col=col.spp(species), pch=16, ylab="Fruit investment (mg)",xlim=c(100,2000000),xlab="")
words.bottom.right.logxy(mod)

mod <- lm (log(total_inv)~log(propagule_inv), data=subset(SummaryInd,propagule_inv>0))
plot(propagule_inv~total_inv,data=subset(SummaryInd,propagule_inv>0), log="xy",col=col.spp(species), pch=16, xlab="", ylab="Propagule investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm (log(total_inv)~log(prepollen_all_inv), data=subset(SummaryInd,prepollen_all_inv>0))
plot(prepollen_all_inv~total_inv,data=subset(SummaryInd,prepollen_all_inv>0), log="xy",col=col.spp(species), pch=16, xlab="", ylab="Prepollen investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mtext("Total invesment (mg)",1, outer=TRUE,cex=1.2)
```

mod <- lm (log(total_inv)~(prop_prepollen_all), data=subset(SummaryInd,prepollen_all_inv>0))
plot(prop_prepollen_all~total_inv,data=subset(SummaryInd,prepollen_all_inv>0), log="x",col=col.spp(species), pch=16, xlab="", ylab="Prepollen investment (mg)", xlim=c(100,2000000))
legend("topright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")
words.bottom.right.logx(mod)

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.width=7.5,fig.height=7.5}
par(mfrow=c(1,1), cex=1, omi=c(.6,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(prop_fruit~RA,data=subset(SummaryInd,prop_propagule>0),log="y",pch=16,col=col.age(age))
mod <- lm(log(prop_fruit)~(RA)*(age),data=subset(SummaryInd,prop_propagule>0&age>3))
  summary(mod)

abline(0,1)
abline(-0.30103,1)
abline(-1,1)
text(120,350,"all RE to accessory tissues",srt=45)
text(12000,720,"10% RE to accessory tissues",srt=45)
text(5,1.5,"50%",srt=45)
```

```
par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
xvar <- "RA"
yvar <- "prop_propagule"

data <- subset(SummaryInd,prop_propagule>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm((x[[yvar]]) ~  ((x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

InvestPetal <- subset(SummaryInvest,parts_fewer=="petals")
View(InvestPetal)
plot(prop_weight,SummaryInvest,color=col.spp(parts_fewer))

```{r}
SummaryInd$prop_fruit <- SummaryInd$fruit_weight / SummaryInd$repro_inv

SummaryInd$costs_for_seeds_produced <- SummaryInd$costs_per_seed*SummaryInd$seed_count
SummaryInd$costs_for_aborted <- SummaryInd$repro_inv - SummaryInd$costs_for_seeds_produced

plot(costs_for_seeds_produced~costs_for_aborted,data=subset(SummaryInd,seed_count>0),log="xy",col=col.spp(species), pch=16)

par(mfrow=c(4,4), cex=1, omi=c(.7,.7,.1,.1), mai=c(.7,.9,.4,0.2)) 
xvar <- "costs_for_aborted"
yvar <- "costs_for_seeds_produced"

data <- subset(SummaryInd,prop_propagule>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

plot(seedcosts~seed_size,data=subset(SummaryInd),log="xy",col=col.spp(species), pch=16)
plot(accessory_per_seed~seed_size,data=subset(SummaryInd),log="xy",col=col.spp(species), pch=16)

par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(total_accessory_Lord ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy", ylab="total per seed accessory costs")
abline(0,1)
mod <- lm(log(total_accessory_Lord) ~ log(seed_size),data=subset(SummarySpp[["mean"]],total_accessory_Lord>0))
words.bottom.right.logxy(mod)
summary(mod)

plot(c_div_p ~ seed_size,data=subset(SummarySpp[["mean"]],species!="COER"&species!="BAER"),col=col.spp(species),pch=16,log="xy")
abline(0,1)
mod <- lm(log(c_div_p) ~ log(seed_size),data=subset(SummarySpp[["mean"]],species!="COER"&species!="BAER",total_accessory_Lord>0))
words.bottom.right.logxy(mod)

plot(b_div_p ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy")
abline(1.1,1)
mod <- lm(log(b_div_p) ~ log(seed_size),data=subset(SummarySpp[["mean"]],total_accessory_Lord>0))
words.bottom.right.logxy(mod)

plot(dispersalcosts_increment ~ seed_size,data=subset(SummarySpp[["mean"]]),col=col.spp(species),pch=16,log="xy")
abline(0,1)
mod <- lm(log(dispersalcosts_increment) ~ log(seed_size),data=subset(SummarySpp[["mean"]],dispersalcosts_increment>0))
words.bottom.right.logxy(mod)

```{r}
par(mfrow=c(4,4), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
xvar <- "total_accessory_Lord"
yvar <- "RA"

data <- subset(SummaryInd,total_accessory_Lord>0&RA>0&age>2)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

```{r}
par(mfrow=c(4,4), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
xvar <- "total_accessory_Lord"
yvar <- "repro_inv"

data <- subset(SummaryInd,total_accessory_Lord>0&RA>0&age>2)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

```{r}
par(mfrow=c(4,4), cex=1, omi=c(.1,.1,.1,.1), mai=c(.5,.5,.5,.5)) 
xvar <- "b_div_p"
yvar <- "repro_inv"

data <- subset(SummaryInd,b_div_p>0&RA>0&age>2)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)
results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",xlab="",ylab="",cex=1, 
                    main=spp,col=col.age(data[[spp]]$age))
}
```

#seedset lower in plants with bigger seeds, except at oldest site, when large seeded species doing relatively better and relationship flat
```{r}
par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(seedset ~ seed_size,data=subset(SummarySppAge[["mean"]],age==32),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(seedset),data=subset(SummarySppAge[["mean"]],age==32))
words.bottom.left.logxy(mod)

plot(seedset ~ seed_size,data=subset(SummarySppAge[["mean"]],age==9),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(seedset),data=subset(SummarySppAge[["mean"]],age==9))
words.bottom.left.logxy(mod)

plot(seedset ~ seed_size,data=subset(SummarySppAge[["mean"]],age==7),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(seedset),data=subset(SummarySppAge[["mean"]],age==7))
words.bottom.left.logxy(mod)

plot(seedset ~ seed_size,data=subset(SummarySppAge[["mean"]],age==5),col=col.spp(species),pch=16,log="xy")
mod <- lm(log(seed_size) ~ log(seedset),data=subset(SummarySppAge[["mean"]],age==5&seedset>0))
words.bottom.left.logxy(mod)
```

```{r}
par(mfrow=c(2,2), cex=1, omi=c(.4,0.3,.3,0.3), mai=c(1.1,1,.3,0.3))
plot(b_div_p ~ seed_size,data=subset(SummaryInd),col=col.age(age),pch=16,log="xy")
abline(0,1)
plot(c_div_p ~ seed_size,data=subset(SummaryInd,species!="COER"),col=col.age(age),pch=16,log="xy")
abline(0,1)
plot(dispersalcosts_increment ~ seed_size,data=subset(SummaryInd),col=col.age(age),pch=16,log="xy")
abline(0,1)
plot(total_accessory_Lord ~ seed_size,data=subset(SummaryInd),col=col.age(age),pch=16,log="xy")
```